### 2013-03-13 R-working/knitr-example/example.Rmd
Fix IshiHara; knitr example

This post examines the features of [R Markdown](http://www.rstudio.org/docs/authoring/using_markdown) 
using [knitr](http://yihui.name/knitr/) in Rstudio 0.96.
This combination of tools provides an exciting improvement in usability for 
[reproducible analysis](http://stats.stackexchange.com/a/15006/183).
Specifically, this post 
(1) discusses getting started with R Markdown and `knitr` in  Rstudio 0.96;
(2) provides a basic example of producing console output and plots using R Markdown;
(3) highlights several code chunk options such as caching and controlling how input and output is displayed;
(4) demonstrates use of standard Markdown notation as well as the extended features of formulas and tables; and 
(5) discusses the implications of R Markdown.
This post was produced with R Markdown. The source code is available here as a gist. 
The post may be most useful if the source code and displayed post are viewed side by side. 
In some instances, I include a copy of the R Markdown in the displayed HTML, but most of the time I assume you are reading the source and post side by side.

<!-- more -->

  ## Getting started
  To work with R Markdown, if necessary:

  * Install [R](http://www.r-project.org/)
* Install the lastest version of [RStudio](http://rstudio.org/download/) (at time of posting, this is 0.96)
* Install the latest version of the `knitr` package: `install.packages("knitr")`

To run the basic working example that produced this blog post:

  * Open R Studio, and go to File - New - R Markdown
* If necessary install `ggplot2` and `lattice` packages: `install.packages("ggplot2"); install.packages("lattice") `
* Paste in the contents of this gist (which contains the R Markdown file used to produce this post) and save the file with an `.rmd` extension
* Click Knit HTML

## Prepare for analyses
```{r }
set.seed(1234)
library(ggplot2)
library(lattice)
```

## Basic console output
To insert an R code chunk, you can type it manually or just press `Chunks - Insert chunks` or use the shortcut key. This will produce the following code chunk:

  ```{r}

```

Pressing tab when inside the braces will bring up code chunk options.

The following R code chunk labelled `basicconsole` is as follows:

  ```{r }
x <- 1:10
y <- round(rnorm(10, x, 1), 2)
df <- data.frame(x, y)
df
```

The code chunk input and output is then displayed as follows:

  ```{r basicconsole}
x <- 1:10
y <- round(rnorm(10, x, 1), 2)
df <- data.frame(x, y)
df
```

## Plots
Images generated by `knitr` are saved in a figures folder. However, they also appear to be represented in the HTML output using a [data URI scheme]( http://en.wikipedia.org/wiki/Data_URI_scheme). This means that you can paste the HTML into a blog post or discussion forum and you don't have to worry about finding a place to store the images; they're embedded in the HTML.

### Simple plot
Here is a basic plot using base graphics:

  ```{r }
plot(x)
```

```{r simpleplot}
plot(x)
```

Note that unlike traditional Sweave, there is no need to write `fig=TRUE`.

### Multiple plots
Also, unlike traditional Sweave, you can include multiple plots in one code chunk:

  ```{r }
boxplot(1:10~rep(1:2,5))
plot(x, y)
```

```{r multipleplots}
boxplot(1:10~rep(1:2,5))
plot(x, y)
```

### `ggplot2` plot
Ggplot2 plots work well:

  ```{r ggplot2ex}
qplot(x, y, data=df)
```

### `lattice` plot
As do lattice plots:

  ```{r latticeex}
xyplot(y~x)
```

Note that unlike traditional Sweave, there is no need to print lattice plots directly.

## R Code chunk features
### Create Markdown code from R
The following code hides the command input (i.e., `echo=FALSE`), and outputs the content directly as code (i.e., `results=asis`, which is similar to `results=tex` in Sweave).

```{r , results='asis', echo=FALSE}
cat("Here are some dot points\n\n")
cat(paste("* The value of y[", 1:3, "] is ", y[1:3], sep="", collapse="\n"))
```

```{r dotpointprint, results='asis', echo=FALSE}
cat("Here are some dot points\n\n")
cat(paste("* The value of y[", 1:3, "] is ", y[1:3], sep="", collapse="\n"))
```

### Create Markdown table code from R
```{r , results='asis', echo=FALSE}
cat("x | y", "--- | ---", sep="\n")
cat(apply(df, 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

```{r createtable, results='asis', echo=FALSE}
cat("x | y", "--- | ---", sep="\n")
cat(apply(df, 1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

### Control output display
The folllowing code supresses display of R input commands (i.e., `echo=FALSE`)
and removes any preceding text from console output (`comment=""`; the default is `comment="##"`).

```{r echo=FALSE, comment="", echo=FALSE}
head(df)
```

```{r echo=FALSE, comment="", echo=FALSE}
head(df)
```

### Control figure size
The following is an example of a smaller figure using `fig.width` and `fig.height` options.

```{r , fig.width=3, fig.height=3}
plot(x)
```

```{r smallplot, fig.width=3, fig.height=3}
plot(x)
```

### Cache analysis
Caching analyses is straightforward.
Here's example code. 
On the first run on my computer, this took about 10 seconds.
On subsequent runs, this code was not run.

If you want to rerun cached code chunks, just [delete the contents of the `cache` folder](http://stackoverflow.com/a/10629121/180892)

```{r , cache=TRUE}
for (i in 1:5000) {
lm((i+1)~i)
}
```

## Basic markdown functionality
For those not familiar with standard [Markdown](http://daringfireball.net/projects/markdown/), the following may be useful.
See the source code for how to produce such points. However, RStudio does include a Markdown quick reference button that adequatly covers this material.

### Dot Points
Simple dot points:

* Point 1
* Point 2
* Point 3

and numeric dot points:

1. Number 1
2. Number 2
3. Number 3

and nested dot points:

* A
* A.1
* A.2
* B
* B.1
* B.2

### Equations
Equations are included by using LaTeX notation and including them either between single dollar signs (inline equations) or double dollar signs (displayed equations).
If you hang around the Q&A site [CrossValidated](http://stats.stackexchange.com) you'll be familiar with this idea.

There are inline equations such as $y_i = \alpha + \beta x_i + e_i$.

And displayed formulas:

  $$\frac{1}{1+\exp(-x)}$$

  knitr provides self-contained HTML code that calls a Mathjax script to display formulas.
However, in order to include the script in my blog posts I [took the script](https://gist.github.com/2716053) and incorporated it into my blogger template.
If you are viewing this post through syndication or an RSS reader, this may not work.
You may need to view this post on my website.

### Tables
Tables can be included using the following notation

A  | B | C
--- | --- | ---
  1  | Male | Blue
2  | Female | Pink

### Hyperlinks

* If you like this post, you may wish to subscribe to [my RSS feed](http://feeds.feedburner.com/jeromyanglim).

### Images
Here's an example image:

![image from redmond barry building unimelb](http://i.imgur.com/RVNmr.jpg)

### Code
Here is Markdown R code chunk displayed as code:

```{r}
x <- 1:10
x
```

And then there's inline code such as `x <- 1:10`.

### Quote
Let's quote some stuff:

> To be, or not to be, that is the question:
> Whether 'tis nobler in the mind to suffer
> The slings and arrows of outrageous fortune,

## Conclusion

* R Markdown is awesome. 
* The ratio of markup to content is excellent. 
* For exploratory analyses, blog posts, and the like R Markdown will be a powerful productivity booster. 
* For journal articles, LaTeX will presumably still be required.
* The RStudio team have made the whole process very user friendly.
* RStudio provides useful shortcut keys for compiling to HTML, and running code chunks. These shortcut keys are presented in a clear way.
* The incorporated extensions to Markdown, particularly formula and table support, are particularly useful.
* Jump-to-chunk feature facilitates navigation. It helps if your code chunks have informative names.
* Code completion on R code chunk options is really helpful. See also [chunk options documentation on the knitr website](http://yihui.name/knitr/options).
* Other recent posts on R markdown include those by :
  * [Christopher Gandrud](http://christophergandrud.blogspot.com.au/2012/05/dynamic-content-with-rstudio-markdown.html)
* [Markcus Gesmann](http://lamages.blogspot.com.au/2012/05/interactive-reports-in-r-with-knitr-and.html)
* [Rstudio on R Markdown](http://rstudio.org/docs/authoring/using_markdown)
* [Yihui Xie](http://yihui.name/knitr/): I really want to thank him for developing `knitr`. 
He has also posted [this example of R Markdown](https://github.com/yihui/knitr/blob/master/inst/examples/knitr-minimal.Rmd).

## Questions
The following are a few questions I encountered along the way that might interest others.
### Annoying `<br/>`'s
**Question:** I asked on the Rstudio discussion site:
  [Why does Markdown to HTML insert `<br/>` on new lines?](http://support.rstudio.org/help/discussions/problems/2329-why-does-r-markdown-to-html-insert-br-when-there-is-a-new-line-of-text)

**Answer:** I just do a find and delete on this text for now.

### Temporarily disable caching
**Question:** I asked on StackOverflow about 
[How to set cache=FALSE for a knitr markdown document and override code chunk settings?](http://stackoverflow.com/q/10628665/180892)

**Answer:**  Delete the cache folder. But there are other possible workflows.

### Equivalent of Sexpr
**Question:** I asked on Stack Overvlow about [whether there an R Markdown equivalent to Sexpr in Sweave?](http://stackoverflow.com/q/10629416/180892).

**Answer:** Include the code between brackets of "backick r space" and "backtick". 
E.g., in the source code I have calculated 2 + 2 = `r 2 + 2` .

### 2013-03-13 R/report.R

@@ -8,9 +8,9 @@ source('R/report-fun.R')
 studyNames     <-  getStudyNames()
 
 #import data
-s<-"Ishihara0000"
+s<-"Aiba2005"
 processStudy(s, verbose = TRUE)
-d<-addStudies(studyNames, reprocess = TRUE, verbose = FALSE)
+d<-addStudies(studyNames, reprocess = FALSE, verbose = FALSE)
 makePlotPanel(d$data, s, pdf=FALSE)
 
 #plot of all data
@@ -37,5 +37,3 @@ writeEmail(d)
 ##
 writeLines(sprintf('%s,', getContributors(d)$name))
 
-
-d$data$dataset[which(d$data$species =="Jc")]

### 2013-03-26 R/reportmd.Rmd
Added automated reports with markdown (first version)

Report for study: `r .study`
========================================================

Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.

Number of records for each variable
-----------------------------------

```{r echo=FALSE}
counts<-apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
counts<-counts[counts>0]  #only include no zeros
counts <- counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
```
```{r echo=FALSE}
cbind(counts)
```

More details
------------------------------------

### Sites

```{r echo=FALSE}
sitevarnames <- var.def$Variable[var.def$Group=="site"]

sitevarnames <- intersect(sitevarnames, names(.dat$data))
```

```{r echo=FALSE, comment=''}
if("location" %in% sitevarnames){
  sites <- sort(levels(.dat$data[,"location"]))
  cat(paste(sites, collapse="\n"))
}
```

```{r echo=FALSE, fig.width=8, fig.height=6}
if("latitude" %in% sitevarnames){
  latlong <- .dat$data[,c("location","latitude","longitude")]
  latlong <- latlong[!duplicated(latlong$location),]

  suppressPackageStartupMessages(library(maps))

  map("world")
  points(latlong$longitude, latlong$latitude, pch=19, col="red")
}
```

### Species

The dataset includes the following species (`r nlevels(.dat$data$species)` species in total):
`r cbind(sort(levels(.dat$data$species)))`

These species belong to the following families:
```{r echo=FALSE}
specvars <- var.def$Variable[var.def$Group=="species"]
specvars <- intersect(specvars, names(.dat$data))
```
```{r echo=FALSE, comment=''}
if("family" %in% specvars){
  families <- sort(unique(.dat$data[,"family"]))
  cat(paste(families, collapse="\n"))
}

```

And include the following plant functional types:
```{r echo=FALSE, comment=''}
if("pft" %in% specvars){
  pfts <- sort(unique(.dat$data[,"pft"]))
  cat(paste(pfts, collapse="\n"))
}

```

Plot of data
---------------------------------
```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
makePlotPanel(d$data, "Ishihara0000", pdf=FALSE)
```

### 2013-03-26 R/reportfun-markdown.R

@@ -1,0 +1,15 @@
+
+
+
+studyReportMd <- function(alldata, study=NULL){
+
+  .study <- study
+  .dat <- extractThisStduy(alldata, study)
+
+  
+  library(knitr)
+  suppressMessages(knit2html("R/reportmd.Rmd", output=paste0("reportmd/",.study,"-report.html")))
+
+}
+  
+  

### 2013-03-26 R/reportmd.Rmd
Update to markdown report (bug fixes), added instructions to report.R

@@
 if("location" %in% sitevarnames){
-  sites <- sort(levels(.dat$data[,"location"]))
+  sites <- sort(unique(as.character(.dat$data[,"location"])))
   cat(paste(sites, collapse="\n"))
@@
 
-```{r echo=FALSE, fig.width=8, fig.height=6}
+```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE}
 if("latitude" %in% sitevarnames){
@@
 
-  suppressPackageStartupMessages(library(maps))
+  library(maps)
 
@@
 
-The dataset includes the following species (`r nlevels(.dat$data$species)` species in total):
-`r cbind(sort(levels(.dat$data$species)))`
+The dataset includes the following species (`r length(unique(.dat$data$species))` species in total):
+`r cbind(sort(unique(as.character(.dat$data$species))))`
 
@@
 ```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
-makePlotPanel(d$data, "Ishihara0000", pdf=FALSE)
+makePlotPanel(d$data, .study, pdf=FALSE)
 ```

### 2013-03-26 R/report.R

@@ -3,6 +3,7 @@ rm(list=ls())
 source('R/Biomass-fun.R')
 source('R/import-fun.R')
 source('R/report-fun.R')
+source('R/reportfun-markdown.R')
 
 #names of all studies
 studyNames     <-  getStudyNames()
@@ -25,6 +26,12 @@ studyReport(d, "Ishihara0000")
 
 writeEmail(d)
 
+
+## NEW: markdown report.
+# Need package 'knitr' installed
+studyReportMd(d, "Ishihara0000")
+studyReportMd(d, "Aiba2007")
+studyReportMd(d, "Bond-Lamberty2002")
 
 ##
 writeLines(sprintf('%s,', getContributors(d)$name))

### 2013-03-26 R/reportfun-markdown.R

@@ -5,7 +5,6 @@ studyReportMd <- function(alldata, study=NULL){
 
   .study <- study
   .dat <- extractThisStduy(alldata, study)
-
   
   library(knitr)
   suppressMessages(knit2html("R/reportmd.Rmd", output=paste0("reportmd/",.study,"-report.html")))

### 2013-03-27 R/reportmd.Rmd
report improved with country maps and more information regarding species traits and what's missing

@@
 ```{r echo=FALSE}
-counts<-apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
-counts<-counts[counts>0]  #only include no zeros
+counts  <-  apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
+counts  <-  counts[counts>0]  #only include no zeros
 counts  <-  counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
@@
 
-### Sites
+### Please analyse if points on the map match locations and coordinates
 
-```{r echo=FALSE}
-sitevarnames <- var.def$Variable[var.def$Group=="site"]
-
-sitevarnames <- intersect(sitevarnames, names(.dat$data))
+```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
+prepdata  <-  prepMapInfo(alldata$data, study)
+drawWorldPlot (prepdata)
 ```
 
-```{r echo=FALSE, comment=''}
-if("location" %in% sitevarnames){
-  sites <- sort(unique(as.character(.dat$data[,"location"])))
-  cat(paste(sites, collapse="\n"))
-}
+### Please analyse if points on country map match locations and coordinates
+
+```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
+repMissingInfo(prepdata)
+drawCountryPlot(prepdata)
 ```
 
-```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE}
-if("latitude" %in% sitevarnames){
-  latlong <- .dat$data[,c("location","latitude","longitude")]
-  latlong <- latlong[!duplicated(latlong$location),]
+### Species - please provide any missing information
 
-  library(maps)
-
-  map("world")
-  points(latlong$longitude, latlong$latitude, pch=19, col="red")
-}
-```
-
-### Species
-
-The dataset includes the following species (`r length(unique(.dat$data$species))` species in total):
-`r cbind(sort(unique(as.character(.dat$data$species))))`
-
-These species belong to the following families:
-```{r echo=FALSE}
-specvars <- var.def$Variable[var.def$Group=="species"]
-specvars <- intersect(specvars, names(.dat$data))
-```
-```{r echo=FALSE, comment=''}
-if("family" %in% specvars){
-  families <- sort(unique(.dat$data[,"family"]))
-  cat(paste(families, collapse="\n"))
-}
-
-```
-
-And include the following plant functional types:
-```{r echo=FALSE, comment=''}
-if("pft" %in% specvars){
-  pfts <- sort(unique(.dat$data[,"pft"]))
-  cat(paste(pfts, collapse="\n"))
-}
-
+The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` families, presenting `r length(unique(.dat$data$pft))` functional types, growing in `r length(unique(.dat$data$growingCondition))` conditions within `r length(unique(.dat$data$vegetation))` major types of habitat:
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+fam.spec(.dat$data)
 ```

### 2013-03-27 R/packages.R

@@ -1,0 +1,2 @@
+library(maps)
+library(mapdata)

### 2013-03-27 R/report-fun.R

@@ -55,7 +55,7 @@ writeEmail<-function(d, fileName=paste("output/Email.txt", sep='')){
 
 
 makePlotPanel<-function(data, study, dir="report", col="grey", pdf=TRUE){
-  cat(study, " ")
+  cat("This is how the study", study, "fits in the entire dataset distribution")
   
   if(pdf){    
     path<-paste0("output/", dir)
@@ -85,81 +85,143 @@ makePlot<-function(data, study, xvar, yvar, xlab, ylab, main="", col="grey"){
   points(data[i,xvar], data[i,yvar], col = "red")  
 }
 
-makeMapPlot<-function(data, study, dir="report", pdf=TRUE){
+prepMapInfo<-function(data, study){
   
-  datum  <-  data[study==data$dataset,]
-  coord  <-  unique(paste0(datum$latitude,";", datum$longitude, ";", datum$location))  
-  catdat <-  data.frame(lat=as.numeric(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][1]}))),
+  data   <-  data[study==data$dataset,]
+  coord  <-  unique(paste0(data$latitude,";", data$longitude, ";", data$location))  
+  data   <-  data.frame(lat=as.numeric(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][1]}))),
                         lon=as.numeric(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][2]}))),
                         loc=as.character(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][3]}))),
                         stringsAsFactors=FALSE)
-  
-  catdat$colour  <-  as.character(sample(colors(),nrow(catdat)))
-  ytext          <-  c(seq(-100,-180,length.out=10), seq(100,180,length.out=10))
-  
-  #if the study has more than 20 coordinates, then we need to split it, otherwise the maps can't be read
-  if(nrow(catdat) <= 20){
-    catdat$ytxt    <-  ytext[1:nrow(catdat)]
-    catdat$xtxt    <-  rep(-175, nrow(catdat))
-    drawPlot(catdat, study=study, dir=dir, pdf=pdf)
-  } else {
-    catdat$org  <-  as.integer(cut(1:nrow(catdat), ceiling(nrow(catdat)/20)))
[…]

### 2013-03-27 R/report.R

@@ -4,6 +4,7 @@ source('R/Biomass-fun.R')
 source('R/import-fun.R')
 source('R/report-fun.R')
 source('R/reportfun-markdown.R')
+source('R/packages.R')
 
 #names of all studies
 studyNames     <-  getStudyNames()

### 2013-03-28 R/reportmd.Rmd
Minor changes to markdown reports. Also added quiet=TRUE option to makePlotPanel

@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, comment=''}
 counts  <-  apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
@@
 counts  <-  counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
-```
-```{r echo=FALSE}
 cbind(counts)
@@
 
-### Please analyse if points on the map match locations and coordinates
+### Please check if the points on the map match locations and coordinates
 
@@
 The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` families, presenting `r length(unique(.dat$data$pft))` functional types, growing in `r length(unique(.dat$data$growingCondition))` conditions within `r length(unique(.dat$data$vegetation))` major types of habitat:
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
 fam.spec(.dat$data)
@@
 ---------------------------------
+
+This is how the study `r .study` (red symbols) fits in the entire dataset.
+
 ```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
-makePlotPanel(d$data, .study, pdf=FALSE)
+makePlotPanel(d$data, .study, pdf=FALSE, quiet=TRUE)
 ```

### 2013-03-28 R/report-fun.R

@@ -54,8 +54,8 @@ writeEmail<-function(d, fileName=paste("output/Email.txt", sep='')){
 }
 
 
-makePlotPanel<-function(data, study, dir="report", col="grey", pdf=TRUE){
-  cat("This is how the study", study, "fits in the entire dataset distribution")
+makePlotPanel<-function(data, study, dir="report", col="grey", pdf=TRUE, quiet=FALSE){
+  if(!quiet)cat("This is how the study", study, "fits in the entire dataset distribution")
   
   if(pdf){    
     path<-paste0("output/", dir)

### 2013-03-28 R/report.R

@@ -29,7 +29,7 @@ writeEmail(d)
 
 
 ## NEW: markdown report.
-# Need package 'knitr' installed
+# Need package 'knitr' installed (added to packages.R; RAD)
 studyReportMd(d, "Ishihara0000")
 studyReportMd(d, "Aiba2007")
 studyReportMd(d, "Bond-Lamberty2002")

### 2013-04-03 R/reportmd.Rmd
new reports including contacts, references

@@
 
+Study reference: `r .dat$ref`
+
+Study contact: `r .dat$contact$name`.
+
+Contact email: `r .dat$contact$email`.
+
+Contact address: `r .dat$contact$address`.
+
+Data description
+-----------------------------------
+
 Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.
+
+The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` family(ies), presenting `r length(unique(.dat$data$pft))` functional type(s), growing in `r length(unique(.dat$data$growingCondition))` condition(s) within `r length(unique(.dat$data$vegetation))` major type(s) of habitat:
 
@@
 
-### Species - please provide any missing information
+### Location level - please provide any missing information
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
+locLevelInfo(.dat$data)
+```
 
-The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` families, presenting `r length(unique(.dat$data$pft))` functional types, growing in `r length(unique(.dat$data$growingCondition))` conditions within `r length(unique(.dat$data$vegetation))` major types of habitat:
+### Stand level - please provide any missing information
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-fam.spec(.dat$data)
+standLevelInfo(.dat$data)
+```
+
+### Species level - please provide any missing information
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
+spLevelInfo(.dat$data)
+```
+
+### Method - please provide any missing information/follow template
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
+printMeta(.dat$data)
 ```

### 2013-04-03 R/report-fun.R

@@ -92,12 +92,12 @@ prepMapInfo<-function(data, study){
   data   <-  data.frame(lat=as.numeric(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][1]}))),
                         lon=as.numeric(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][2]}))),
                         loc=as.character(unlist(lapply(coord, function(x){strsplit(x,";")[[1]][3]}))),
+                        country=NA,
                         stringsAsFactors=FALSE)
   i      <-  !is.na(data$lat) | !is.na(data$lon)
   if( any(i) ){
-    data$country[i]  <-  map.where(x=data$lon[i], y=data$lat[i])
-  }
-  data$country       <-  as.character(unlist(lapply(data$country, function(x){strsplit(x,":")[[1]][1]})))
+    data$country[i]  <-  as.character(strsplit(map.where(x=data$lon[i], y=data$lat[i]),":")[[1]][1])
+  }  
 
   #final object
   data
@@ -208,23 +208,60 @@ drawCountryPlot  <-  function(data){
   }
 }
 
-fam.spec  <-  function(data){
+locLevelInfo  <-  function(data){
+  loc  <-  unique(paste(data$location, data$map, data$mat, data$longitude, data$latitude, data$vegetation, sep=";"))
+  loc  <-  data.frame(Location=as.character(unlist(lapply(loc, function(x){strsplit(x,";")[[1]][1]}))),
+                      MAP=as.numeric(unlist(lapply(loc, function(x){strsplit(x,";")[[1]][2]}))),
+                      MAT=as.numeric(unlist(lapply(loc, function(x){strsplit(x,";")[[1]][3]}))),
+                      Longitude=as.numeric(unlist(lapply(loc, function(x){strsplit(x,";")[[1]][4]}))),
+                      Latitude=as.numeric(unlist(lapply(loc, function(x){strsplit(x,";")[[1]][5]}))),
+                      Vegetation_Type=as.character(unlist(lapply(loc, function(x){strsplit(x,";")[[1]][6]}))),
+                      stringsAsFactors=FALSE)
+  
+  loc[is.na(loc) | loc=="" | loc=="NA"]  <-  "????"  
+  loc
+}
+
+
+standLevelInfo  <-  function(data){
[…]

### 2013-04-03 R/report.R

@@ -3,7 +3,6 @@ rm(list=ls())
 source('R/Biomass-fun.R')
 source('R/import-fun.R')
 source('R/report-fun.R')
-source('R/reportfun-markdown.R')
 source('R/packages.R')
 
 #names of all studies
@@ -33,6 +32,8 @@ writeEmail(d)
 studyReportMd(d, "Ishihara0000")
 studyReportMd(d, "Aiba2007")
 studyReportMd(d, "Bond-Lamberty2002")
+
+l  <-  lapply(studyNames, function(x){studyReportMd(d, x)})
 
 ##
 writeLines(sprintf('%s,', getContributors(d)$name))

### 2013-04-03 R/reportmd.Rmd
Metadata information include; metadata files standardised

@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-printMeta(.dat$data)
+metaTable  <-  printMeta(.dat$data)
+for(j in 1:nrow(metaTable)){
+  cat(paste0(metaTable[j,1], ": ", metaTable[j,2], "\n"))
+}
+
 ```

### 2013-04-03 R/report-fun.R

@@ -256,7 +256,6 @@ spLevelInfo  <-  function(data){
 }
 
 printMeta  <-  function(data){
-  browser()
   dataset   <-  as.character(unique(data$dataset))
   openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
   openMeta

### 2013-04-09 R/reportmd.Rmd
Improved formatting of knitr report

--- 2013-04-03 R/reportmd.Rmd
@@
-Report for study: `r .study`
-========================================================
+# Report for study: `r .study`
 
-Study reference: `r .dat$ref`
+**Reference:** `r .dat$ref[2]`
 
-Study contact: `r .dat$contact$name`.
+**Data contributor:** `r .dat$contact$name`.
 
-Contact email: `r .dat$contact$email`.
+**Email:** `r .dat$contact$email`.
 
-Contact address: `r .dat$contact$address`.
+**Address:** `r .dat$contact$address`.
 
-Data description
------------------------------------
+## Overview of data provided
 
@@
 
-Number of records for each variable
------------------------------------
+### Number of records for each variable
 
@@
 
-More details
-------------------------------------
+## Study details
 
-### Please check if the points on the map match locations and coordinates
+### Location
+
+**Location dtaa**
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
+locLevelInfo(.dat$data)
+```
+
+**Global map**
 
@@
 
-### Please analyse if points on country map match locations and coordinates
+**Country map**
 
@@
 
-### Location level - please provide any missing information
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-locLevelInfo(.dat$data)
-```
+**Stand description**
 
-### Stand level - please provide any missing information
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
@@
 
-### Species level - please provide any missing information
+### Species data
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
@@
 
-### Method - please provide any missing information/follow template
+### Methods
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-printMeta(.dat$data)
+metaTable  <-  printMeta(.dat$data)
+for(j in 1:nrow(metaTable)){
+  cat(paste0(metaTable[j,1], ": ", metaTable[j,2], "\n\n"))
+}
+
 ```
 
-Plot of data
----------------------------------
+## Plots of data
 

### 2013-04-16 R/reportmd.Rmd
every possible combination of plots added, added a small change to Baraloto2006's latitude because the original coordinate provided was too close to the sea and the package maps was not able to recognize the country

@@
 
-**Location dtaa**
+**Location data**
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
@@
 
-This is how the study `r .study` (red symbols) fits in the entire dataset.
+This is how the study `r .study` (each color represents a species) fits in the entire dataset (grey).
 

### 2013-05-14 report/reportmd.Rmd
Moved report scripts to own folder

--- 2013-04-16 R/reportmd.Rmd
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
+options(width =160)
 locLevelInfo(.dat$data)
@@
 ```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
-makePlotPanel(d$data, .study, pdf=FALSE, quiet=TRUE)
+makePlotPanel(alldata$data, .study, pdf=FALSE, quiet=TRUE)
 ```

### 2013-05-14 report/report-byStudy.R

@@ -1,13 +1,19 @@
 rm(list=ls())
 source('R/packages.R')
 source('R/import-fun.R')
-source('R/report-fun.R')
+source('report/report-fun.R')
 
 # Load all data
-dat <- loadStudies(reprocess=TRUE)
+dat <- loadStudies(reprocess=FALSE)
 
-# List all study names
-studyNames <- getStudyNames()
+# Markdown report (HTML)
+studyReportMd(dat, "Ilomaki2003")
+studyReportMd(dat, "Parviainen1999")
+
+# All reports:
+allStudyReport(dat)
+
+
 
 #create dataNew.csv files for authors to fill in
 generateAllDataNew(dat, studyNames)
@@ -15,12 +21,3 @@ generateAllDataNew(dat, studyNames)
 #-- Not touched : will change with new email code??
 #for each study, put variable.definition, contactInfo and Reference in the email folder
 y  <-  lapply(studyNames, function(x)emailFiles(d$data, x))
-
-
-# Markdown report (HTML)
-studyReportMd(dat, "Ilomaki2003")
-studyReportMd(dat, "Parviainen1999")
-
-# All reports:
-allStudyReport(dat, studyNames)
-

### 2013-05-14 report/report-fun.R

@@ -9,7 +9,7 @@ extractStudy<-function(alldata, study){
 }
 
 
-allStudyReport <- function(data, studynames, progressbar=TRUE){
+allStudyReport <- function(data, studynames= getStudyNames(), progressbar=TRUE){
 
   message("Generating ", length(studynames), " markdown reports.")
   if(progressbar){
@@ -363,7 +363,7 @@ studyReportMd <- function(alldata, study=NULL, Dir="report-per-study", delete=TR
   path<-paste0("output/", Dir)
   if(!file.exists(path)) dir.create(path)
   
-  suppressMessages(knit2html("R/reportmd.Rmd", quiet=TRUE))
+  suppressMessages(knit2html("report/reportmd.Rmd", quiet=TRUE))
   
   file.copy("reportmd.html", paste0(path,"/",.study,"-report.html"))
   

### 2013-05-15 R/formatBib.R
Changed referencing to use .bib, added these to all folders

@@ -5,8 +5,8 @@ formatBib <- function(bibfile,
                       fieldSep=". "
                       ){
   
-  require(gdata)
-  require(bibtex)
+  library(gdata, quietly=TRUE)
+  library(bibtex, quietly=TRUE)
   
   bibs <- read.bib(bibfile)
   

### 2013-05-15 R/import.R

@@ -465,10 +465,8 @@ data.path <- function(studyName, ...){
 }
 
 readReference <- function(studyName) {
-  ref <- read.csv(data.path(studyName, "studyRef.csv"),
-                  header=TRUE, stringsAsFactors=FALSE,
-                  strip.white=TRUE)
-  data.frame(dataset=studyName, ref)
+  reference <-data.path(studyName, "studyRef.bib")
+  data.frame(dataset=studyName, reference)
 }
 
 readContact <- function(studyName) {

### 2013-05-15 report/report-byStudy.R

@@ -2,6 +2,7 @@ rm(list=ls())
 source('R/import.R')
 source('report/report-fun.R')
 source('R/plotting.R')
+source('R/formatBib.R')
 
 # Load all data
 dat <- loadStudies(reprocess=FALSE)

### 2013-05-15 report/report-fun.R

@@ -20,36 +20,6 @@ allStudyReport <- function(data, studynames= getStudyNames(), progressbar=TRUE){
   
 }
 
-studyReport<-function(alldata, study="all"){
-  if(study =="all")
-    d<-alldata
-  else
-    d<- extractStudy(alldata, study)
-  
-  cat("\nSTUDIES: ", as.character(unique(d$data$dataset)), "\n")
-  
-  cat("Data for",  length(d$data$dataset),"individuals from", length(unique(d$data$location)), "locations covering",  length(unique(d$data$species)), "species.\n\n")
-  
-  cat("\nNUMBER OF RECORDS FOR EACH VARIABLE: \n")  
-  counts<-apply(!is.na(d$data), MARGIN=2, FUN = sum)
-  counts<-counts[counts>0]  #only include no zeros
-  #only include tree variables
-  (counts<-counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]])
-  
-  cat("\nMORE DETAIL:\n")  
-  cat("\nSITES:\n")  
-  cat("\nprint information for each site, ie. ", var.def$Variable[var.def$Group=="site"])  
-  cat("\nmap of locations used")  
-  
-  cat("\nSPECIES: ", length(unique(d$data$species)), "\n", paste0(as.character(unique(sort(d$data$species))), "; "), "\n")
-  cat("\nprint information for each species, ie. ", var.def$Variable[var.def$Group=="species"])  
-  cat("\n also number of records for each species")  
-  
-  #Brief look at data
-  makePlotPanel(alldata$data, study, pdf=FALSE)
-  
-}
-
 
 locLevelInfo  <-  function(data){
   
@@ -108,7 +78,7 @@ studyReportMd <- function(alldata, study=NULL, Dir="report-per-study", delete=TR
   
   #Check output directory exists
   path<-paste0("output/", Dir)
-  if(!file.exists(path)) dir.create(path)
+  if(!file.exists(path)) dir.create(path, recursive=TRUE)
   
   suppressMessages(knit2html("report/reportmd.Rmd", quiet=TRUE))
   
@@ -212,8 +182,8 @@ generateDataNew  <-  function(data, studyName){
 
 emailFiles  <-  function(data, studyName){
[…]

### 2013-05-17 report/reportmd.Rmd
Changed referencing to use .bib, added these to all folders
improve creation of knitr scripts
Initial from baad1. Removed bad line endings on csv files. Moved pdfs and all accessory friles to xtra folder for each data folder.

@@
 
-**Reference:** `r .dat$ref[2]`
+**Reference:** `r formatBib(paste0("../", as.character(.dat$ref$reference)))[[1]]`
 
@@
 ```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
-prepdata  <-  prepMapInfo(alldata$data, study)
+prepdata  <-  prepMapInfo(alldata$data, .study)
 drawWorldPlot (prepdata)

### 2013-05-17 R/formatBib.R

@@ -1,0 +1,88 @@
+
+
+formatBib <- function(bibfile,
+                      initialSep=".",
+                      fieldSep=". "
+                      ){
+  
+  library(gdata, quietly=TRUE)
+  library(bibtex, quietly=TRUE)
+  
+  bibs <- read.bib(bibfile)
+  
+  
+  f <- lapply(bibs, function(b){
+    
+    b <- unclass(b)[[1]]
+    
+    # strip html tags in title (typically <i> etc.):
+    tit <- gsub("<.+?>","",b$title)
+    Encoding(tit) <- "UTF-8"  #allows for funny characters
+    
+    # author initials
+    a <- unclass(b$author)
+    
+    # family name
+    family <- unlist(sapply(a, "[[", "family"))
+    # extra unlist is necessary for NULL elements
+    
+    Encoding(family) <- "UTF-8"
+    
+    # Extract initials from given names
+    given <- lapply(a, "[[", "given")
+    initials <- lapply(given, function(x){
+      
+      # If initials only, they are in {}
+      if(any(grepl("}",x,fixed=TRUE))){
+        cl <- gsub("{","",x,fixed=TRUE)
+        cl <- gsub("}","",cl,fixed=TRUE)
+        cl <- gsub(".","",cl,fixed=TRUE)
+        
+        init <- strsplit(cl, "")[[1]]
+        return(init)
+      } else {  
+        substr(x,1,1)
+      }
+    })
+    
+    
+    initials <- unname(sapply(initials, function(x){
+      paste0(paste(x, collapse=initialSep),initialSep)
+      }))
+    
+    # author names
+    p <- paste(family, initials, collapse=", ")
+    
+    # field separator, unless it is identical to the initial separator
+    if(trim(fieldSep) != trim(initialSep))
+      p <- paste0(p, fieldSep)
+    else
+      p <- paste0(p, " ")
+    
+    # year
+    p <- paste0(p, b$year, fieldSep)
+    
+    # title
+    p <- paste0(p, tit, fieldSep)
+    
+    # journal name
+    jour <- b$journal
+    Encoding(jour) <- "UTF-8"
+    p <- paste0(p, jour, fieldSep)
+    
+    # not sure how often this happens ... double '-'
+    pag <- gsub("--","-",b$pages)
+    Encoding(pag) <- "UTF-8"  # allows for long -
+    
+    # volume and page numbers
+    p <- paste0(p, b$volume,":",pag)
+  
+  return(p)
+  })
+  
+  
+return(f)
+}
+
+
+

### 2013-05-17 R/import.R

@@ -1,0 +1,519 @@
+library(bibtex, quietly=TRUE)
+
+#set directories for study
+dir.rawData   <-  "data" 
+dir.cleanData <-  "output/data"
+var.def       <-  read.csv("config/variableDefinitions.csv", h=TRUE, stringsAsFactors=FALSE)#variable definitions
+met.def       <-  read.csv("config/methodsDefinitions.csv", h=TRUE, stringsAsFactors=FALSE)#definition of methods
+
+
+
+
+# Get list of studies included in database
+getStudyNames <-function(){
+  dir(dir.rawData)
+}
+
+
+
+#' Adds studies to the central dataset
+#'
+#' @param studyNames Character vector of study names to be added
+#' @param data If provided, will add studies to this dataframe
+#' @param reprocess If TRUE, will reprocess studies even if they already exist in the data directory
+#' @param verbose If TRUE, print messages to screen, good for isolating problems  
+#' @return merged list with three parts: data, reference, contact,
+#'    each is a dataframe with all data combined.
+#' @keywords misc
+#' @export
+loadStudies <- function(studyNames=getStudyNames(), data=NULL,
+                        reprocess=FALSE, verbose=FALSE) {
+  d <- lapply(studyNames, loadStudy, reprocess=reprocess,
+              verbose=verbose)
+  vars <- c("data", "ref", "contact")
+  f <- function(v)
+    do.call(rbind, lapply(d, "[[", v))
+  structure(lapply(vars, f), names=vars)
+}
+
+##' Load data from specified studyName
+##'
+##' @param studyName: name of folder where data stored
+##' @param reprocess: force data to be reprocessed
+##' @param verbose: print tsages to screen, good for isolating problems  
+##' @return list with three parts: data, reference, contact
+##' @export
+loadStudy <- function(studyName, reprocess=FALSE, verbose=FALSE) {
+  if (verbose)
+    cat(studyName, " ")
+
+  list(data=readDataProcessed(studyName, reprocess),
+       ref=readReference(studyName),
+       contact=readContact(studyName))
+}
+
+readDataProcessed <- function(studyName, reprocess=TRUE) {
+  filename <- studyDataFile(studyName)
+  if ( reprocess || !file.exists(filename) )
+    processStudy(studyName)
+  else
+    read.csv(filename, header=TRUE, stringsAsFactors=FALSE)
+}
+
+## TODO: Roxygenise(?)
+# Processes raw data from studyName, for incorporation into main dataset.
[…]

### 2013-05-17 R/load.R

@@ -1,0 +1,12 @@
+rm(list=ls())
+source('R/import.R')
+source('R/packages.R')
+
+#prepare new files for pre-import
+NewStudies  <-  getStudyNames()
+newS        <-  NewStudies[59]
+makeDataImport(newS)
+setUpFiles(newS)
+
+#process all studies
+d <- loadStudies(reprocess=TRUE)

### 2013-05-17 R/plotting.R

@@ -1,0 +1,285 @@
+library(maps, quietly=TRUE)
+library(mapdata, quietly=TRUE)
+library(gdata, quietly=TRUE)
+
+
+
+makePlotPanel<-function(data, study, dir="report-per-study", col="grey", pdf=TRUE, quiet=FALSE){
+  if(!quiet)cat("This is how the study", study, "fits in the entire dataset distribution")
+  
+  if(pdf){    
+    path<-paste0("output/", dir)
+    if(!file.exists(path))
+      dir.create(path)
+    pdf(file=paste0(path,"/", study,".pdf"))
+  }
+  
+  dat        <-  data[data$dataset==study,]
+  plot.vars  <-  var.def$Variable[var.def$Group=="tree"]
+  plot.vars  <-  plot.vars[plot.vars %in% c("growingCondition","status","light") == FALSE]
+  available  <-  dat[,names(dat) %in% plot.vars]   
+  available  <-  apply(available,2,as.numeric)
+  available  <-  available[,colSums(available, na.rm=TRUE) != 0]
+  
+  #set up a vector of colors, each species with different color
+  species  <-  as.numeric(as.factor(dat$species))
+  if(min(species)!=1){species  <-  species - (min(species)-1)}
+  colorvec <-  niceColors(length(unique(species)))
+  colorvec <-  colorvec[match(species, unique(species))]
+  
+  par(mfrow=c(2,2))
+  z  <- 0
+  for(j in 1:(dim(available)[2]-1)){
+    for(k in (j+1):dim(available)[2]){
+      testNa  <-  available[,j]+available[,k]
+      if(length(testNa[!is.na(testNa)]) != 0){
+        z  <-  z + 1
+        if(z==5){par(mfrow=c(2,2)); z  <- 1}
+        varjx  <-  paste0(var.def$label[var.def$Variable==colnames(available)[j]], " (", var.def$Units[var.def$Variable==colnames(available)[j]], ")")
+        varjy  <-  paste0(var.def$label[var.def$Variable==colnames(available)[k]], " (", var.def$Units[var.def$Variable==colnames(available)[k]], ")")
+        makePlot(data, available, studycol=colorvec, xvar = colnames(available)[j], yvar = colnames(available)[k],  xlab=varjx, ylab=varjy)
+      }
+    }
[…]

### 2013-05-17 data/Aiba2007/dataManipulate.r

@@ -1,0 +1,5 @@
+manipulate <- function(raw) {
+  raw[["species"]]              <-  paste(raw[["genus"]], raw[["species"]])
+  raw
+}
+

### 2013-05-17 data/Baraloto2006/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw$grouping  <-  makeGroups(raw, c("Soil","H2O", "P"))
+  
+  raw
+}
+

### 2013-05-17 data/Bond-Lamberty2002/dataManipulate.r

@@ -1,0 +1,9 @@
+manipulate <- function(raw) {
+  
+  raw$grouping  <-  makeGroups(raw, c("Site","HarvestYear","Edaphic"))
+  raw$age       <-  raw[["HarvestYear"]] - raw[["BurnYear"]] 
+  raw[["stemMass"]]  <-  raw[["TotBranch"]] + raw[["Stem"]]
+  
+  raw
+}
+

### 2013-05-17 data/Coll2008/dataManipulate.r

@@ -1,0 +1,29 @@
+manipulate <- function(raw) {
+  # leaf and biomass measurements come from different individuals, so I'll take 
+  # the average leaf measure from each variable and species and then paste this 
+  # into the individual trees that contain the biomass data
+  raw1       <-  raw[!is.na(raw[["tree"]]), 1:6]
+  raw1.mean  <-  
+    data.frame(individual.leaf.mass.g=
+                 tapply(raw1[["individual.leaf.mass.g"]], raw1[["species"]], mean),            
+               individual.leaf.area.cm2=
+                 tapply(raw1[["individual.leaf.area.cm2"]], raw1[["species"]], mean),
+               specific.leaf.area.cm2.g=
+                 tapply(raw1[["specific.leaf.area.cm2/g"]], raw1[["species"]], mean))
+  
+  raw       <-  raw[!is.na(raw[["biomass.samples"]]), c(1,4:16)]
+  
+  raw[["individual.leaf.mass.g"]] <- raw1.mean[match(raw[["species"]], rownames(raw1.mean)),
+                                                    "individual.leaf.mass.g"]
+  raw[["individual.leaf.area.cm2"]] <- raw1.mean[match(raw[["species"]], rownames(raw1.mean)), 
+                                                  "individual.leaf.area.cm2"]
+  raw[["specific.leaf.area.cm2/g"]] <- raw1.mean[match(raw[["species"]], rownames(raw1.mean)), 
+                                                    "specific.leaf.area.cm2.g"]
+  raw[["stem.biomass.g"]] <- raw[["branch.biomass.g"]] + raw[["trunk.biomass.g"]]
+  rm(raw1)
+  
+  raw                            
+  
+  raw
+}
+

### 2013-05-17 data/Delagrange0000a/dataManipulate.R

@@ -1,0 +1,5 @@
+manipulate <- function(raw) {
+  raw$grouping  <-  makeGroups(raw, c("plot_ID", "Subplot_ID"))
+  raw
+}
+

### 2013-05-17 data/Delagrange0000b/dataManipulate.R

@@ -1,0 +1,5 @@
+manipulate <- function(raw) {
+  raw$grouping  <-  makeGroups(raw, c("Site", "Seedbed"))
+  raw
+}
+

### 2013-05-17 data/Delagrange2004/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw$grouping             <- makeGroups(raw, c("Group", "Last perturbation"))
+  
+  
+  raw
+}
+

### 2013-05-17 data/Epron2011/dataManipulate.r

@@ -1,0 +1,12 @@
+manipulate <- function(raw) {
+  raw[["a.babh"]] <-  raw[["a.st"]] - raw[["a.st1"]]
+  raw[["group"]]  <-  paste(raw[["Variable"]], raw[["contributor"]], raw[["Nutrition"]], sep="; ")
+  raw[["a.cs"]]  <-  pi*(raw[["a.cp"]]/100)^2
+  
+  #Error correction - heights for plants age =0.5 at top of datasets in different units to rest of datasets (cm rather than m)
+  i <-1:8
+  raw[["h.t"]][i] <- raw[["h.t"]][i]/100
+  raw[["h.c"]][i] <- raw[["h.c"]][i]/100
+  raw
+}
+

### 2013-05-17 data/Gargaglione2010/dataManipulate.R

@@ -1,0 +1,13 @@
+manipulate <- function(raw) {
+  raw$stem_mass         <-  raw[["total sapwood"]] + raw[["Total bark mass"]] + raw[["total heartwood"]]
+  raw$branch_mass       <-  raw[["branches sapwood mass"]] + raw[["branches bark mass"]] + raw[["branches heartwood mass"]]
+  raw$coarse_root_mass  <-  raw[["medium roots mass"]] + raw[["coarse roots mass"]]
+  
+  raw[["total N sapwood"]]    <-  (raw[["total N sapwood"]]/1000)/raw[["total sapwood"]]
+  raw[["Total N bark"]]       <-  (raw[["Total N bark"]]/1000)/raw[["Total bark mass"]]
+  raw[["total N heartwood"]]  <-  (raw[["total N heartwood"]]/1000)/raw[["total heartwood"]]
+  raw[["N fine roots"]]       <-  (raw[["N fine roots"]]/1000)/raw[["fine roots mass"]]
+  raw$NcoarseRoots            <-  ((raw[["N medium roots"]] + raw[["N coarse roots"]])/1000)/raw$coarse_root_mass
+  raw
+}
+

### 2013-05-17 data/Ishihara0000/dataManipulate.R

@@ -1,0 +1,30 @@
+manipulate <- function(raw) {
+  
+  
+  
+  raw$Age[raw$Age == "33-34"] <- "33.5"
+  raw$Age[raw$Age == "22-23"] <- "22.5"
+  raw$Age <- as.numeric(raw$Age)
+  
+  D0true <- with(raw, (D0 + D0.2)/2 )
+  raw$D0[!is.na(raw$D0.2)] <- D0true[!is.na(raw$D0.2)]
+  
+  DBHtrue <- with(raw, (DBH + DBH.2)/2 )
+  raw$DBH[!is.na(raw$DBH.2)] <- DBHtrue[!is.na(raw$DBH.2)]
+  
+  DLtrue <- with(raw, (DL + DL.2)/2 )
+  raw$DL[!is.na(raw$DL.2)] <- DLtrue[!is.na(raw$DL.2)]
+  raw$a.stbc <- (pi/4) * raw$DL^2
+  
+  # Separate stem mass from branch mass:
+  raw$WS <- raw$WSB - raw$WB
+  
+  
+  raw$SiteStand <- paste(raw$Country, raw$Site, raw$Stand, sep="-")
+  
+  raw$Forest.type[raw$Forest.type == "S?"] <- "S"
+  
+  raw$Ref2 <- paste("Ishihara0000", raw$Ref, sep="-")
+  raw
+}
+

### 2013-05-17 data/Kenzo2009/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw[["n.lf.per"]]  <-  (raw[["n.lf.per"]]/100) * raw[["lf.ma.g.m2"]] # this transforms it from percentage (/unit mass) to /unit area --> final unit g/m2
+  raw[["group"]]  <-  paste(raw[["group"]], raw[["location"]], raw[["contributor"]], sep="; ")
+  
+  raw
+}
+

### 2013-05-17 data/Kenzo2009b/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw$grouping  <-  makeGroups(raw, c("location", "contributor"))
+  
+  raw
+}
+

### 2013-05-17 data/Kohyama1987/dataManipulate.r

@@ -1,0 +1,8 @@
+manipulate <- function(raw) {
+  raw[["leaf.mass"]]  <-  raw[["Wtl.g"]] + raw[["Wbl.g"]]
+  raw[["m.st"]]       <-  raw[["Wts.g"]] + raw[["Wbs.g"]]
+  
+  
+  raw
+}
+

### 2013-05-17 data/Kohyama1990/dataManipulate.r

@@ -1,0 +1,5 @@
+manipulate <- function(raw) {
+  raw$grouping  <-  makeGroups(raw, c("Year", "#"))
+  raw
+}
+

### 2013-05-17 data/Mori1991/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw[raw=="No data"] <-  NA
+  
+  raw
+}
+

### 2013-05-17 data/O'Grady2000/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw$grouping  <-  makeGroups(raw, c("SITE"))
+  
+  raw
+}
+

### 2013-05-17 data/O'Grady2006/dataManipulate.r

@@ -1,0 +1,13 @@
+manipulate <- function(raw) {
+  
+  raw[["stem"]]       <-  raw[["stem"]]+raw[["branch"]]
+  raw[["m.rf"]]       <-  raw[["5-Oct"]] + raw[["Oct-15"]]
+  raw[["m.rc"]]       <-  raw[["15 - 20"]] + raw[["20  + stump"]]
+  raw$grouping        <-  makeGroups(raw, c("Harvested","plot","treatment","seedling"))
+  
+  #Exclude data from zulu plot, these seem to be in different units. Send query to author
+  raw  <-  raw[raw[["plot"]] != "zulu",]
+  raw  <-  raw[!(raw[["plot"]] == 1 & raw[["seedling"]] ==16),]
+  raw
+}
+

### 2013-05-17 data/Osada0000/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw           <-  raw[raw[["Source"]] != "Osada et al. (2003) Forest Ecology and Management" & raw[["Source"]] != "Osada (2005) Canadian Journal of Botany", ]
+  raw[["SLA"]]  <-  (1/raw[["SLA"]])*10000
+  raw[["LMA"]]  <-  mapply(function(x,y){sum(c(x,y),na.rm=TRUE)},raw[["SLA"]],raw[["LMA"]]) 
+  raw
+}
+

### 2013-05-17 data/Osada2005/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw           <-  raw[raw[["Source"]] == "Osada (2005) Canadian Journal of Botany", ]
+  raw[["SLA"]]  <-  (1/raw[["SLA"]])*10000
+  raw[["LMA"]]  <-  mapply(function(x,y){sum(c(x,y),na.rm=TRUE)},raw[["SLA"]],raw[["LMA"]])  
+  raw
+}
+

### 2013-05-17 data/Peri2008/dataManipulate.R

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw       <-  raw[raw$ref=="peri2008",]
+  raw$n.lf  <-  (raw$n.lf*raw$m.lf)/raw$a.lf
+  raw
+}
+

### 2013-05-17 data/Peri2011/dataManipulate.R

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw       <-  raw[raw$ref=="peri2011",]
+  raw$n.lf  <-  (raw$n.lf*raw$m.lf)/raw$a.lf
+  raw
+}
+

### 2013-05-17 data/Rodriguez2003/dataManipulate.r

@@ -1,0 +1,5 @@
+manipulate <- function(raw) {
+  raw[["Bark_Stemwood"]]                    <-  raw[["Bark"]] + raw[["Stemwood"]]
+  raw
+}
+

### 2013-05-17 data/Selaya2007/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw        <-  raw[raw[["reference"]]=="Selaya et al. (2007) Annals of Botany 99:141-151; Selaya & Anten (2010) Ecology 91: 1102-1113",]
+  raw[["light"]]  <-  raw[["light"]]/55*100
+  
+  raw
+}
+

### 2013-05-17 data/Selaya2008/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw        <-  raw[raw[["reference"]]=="Selaya & Anten (2008) Functional Ecology 22: 30-39",]
+  raw[["light"]]  <-  raw[["light"]]/55*100
+  
+  raw
+}
+

### 2013-05-17 data/Selaya2008b/dataManipulate.r

@@ -1,0 +1,7 @@
+manipulate <- function(raw) {
+  raw        <-  raw[raw[["reference"]]=="Selaya et al. (2008) Journal of Ecology 96: 1211-1221; Selaya & Anten (2010) Ecology 91:1102-1113",]
+  raw[["light"]]  <-  raw[["light"]]/55*100
+  
+  raw
+}
+

### 2013-05-17 data/Sillett2010/dataManipulate.r

@@ -1,0 +1,9 @@
+manipulate <- function(raw) {
+  for(j in c(2,5:16)){
+    raw[,j]    <-  unlist(lapply(raw[,j], function(x){as.numeric(strsplit(x, "_")[[1]][1])}))
+  }
+  raw[["Tree"]]     <-  unlist(lapply(gsub("s ", "s_", raw[["Tree"]]), function(x){strsplit(x, "_")[[1]][1]}))  
+  
+  raw
+}
+

### 2013-05-17 data/Valladares2000/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw[["species"]]  <-  gsub("P.", "Psychotria", raw[["species"]])
+  
+  raw
+}
+

### 2013-05-17 data/Vanninen2005/dataManipulate.R

@@ -1,0 +1,5 @@
+manipulate <- function(raw) {
+  names(raw)[names(raw)=="species"]  <-  "wrong.sp"
+  raw
+}
+

### 2013-05-17 data/Yamada2000/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw[["stem dry mass (g)"]]  <-  raw[["stem dry mass (g)"]] + raw[["branch dry mass (g)"]]
+  
+  raw
+}
+

### 2013-05-17 data/leMaire2011/dataManipulate.r

@@ -1,0 +1,6 @@
+manipulate <- function(raw) {
+  raw[["group"]]  <-  paste(raw[["Variable.Unit"]], sep="; ")
+  
+  raw
+}
+

### 2013-05-17 report/report-bivarPlots.R

@@ -1,0 +1,14 @@
+rm(list=ls())
+source('R/import.R')
+source('report/report-fun.R')
+
+#names of all studies
+studyNames     <-  getStudyNames()
+
+#import data
+d  <-  addStudies(studyNames, reprocess = FALSE, verbose = FALSE)
+
+#Explore
+alldata  <-  d$data
+
+comparePlots(alldata)

### 2013-05-17 report/report-byStudy.R

@@ -1,0 +1,23 @@
+rm(list=ls())
+source('R/import.R')
+source('R/plotting.R')
+source('R/formatBib.R')
+source('report/report-fun.R')
+
+# Load all data
+dat <- loadStudies(reprocess=FALSE)
+
+# Markdown report (HTML)
+printStudyReport(dat, "Ilomaki2003")
+
+# All reports:
+allStudyReport(dat)
+
+
+
+#create dataNew.csv files for authors to fill in
+generateAllDataNew(dat, studyNames)
+
+#-- Not touched : will change with new email code??
+#for each study, put variable.definition, contactInfo and Reference in the email folder
+y  <-  lapply(studyNames, function(x)emailFiles(d$data, x))

### 2013-05-17 report/report-fun.R

@@ -1,0 +1,218 @@
+library(knitr, quietly=TRUE)
+
+dir.Emails    <-  "output/email"
+
+
+allStudyReport <- function(data, studynames= getStudyNames(), progressbar=TRUE){
+  
+  message("Generating ", length(studynames), " markdown reports.")
+  if(progressbar){
+    wp <- txtProgressBar(min = 0, max = length(studynames), 
+                         initial = 0, width = 50, style=3)
+  }
+  
+  for(i in seq_along(studynames)){
+    printStudyReport(data, studynames[i])
+    if(progressbar)setTxtProgressBar(wp, i)
+  }
+  
+  if(progressbar)close(wp) 
+}
+
+
+locLevelInfo  <-  function(data){
+  
+  vars <- c("location","map","mat","longitude","latitude","vegetation")
+  
+  loc <- data[!duplicated(data[,vars]),vars]
+  
+  loc[is.na(loc) | loc=="" | loc=="NA"]  <-  "????"  
+  return(loc)
+}
+
+
+standLevelInfo  <-  function(data){
+  
+  vars <- c("location","grouping","growingCondition","status")
+  sta <- data[!duplicated(data[,vars]),vars]
+  
+  
+  if(all(is.na(sta$grouping)))
+    sta  <-  sta[,c("location", "growing_condition", "status")]
+  
+  sta[is.na(sta) | sta=="" | sta=="NA"]  <-  "????"  
+  
+  return(sta)
+}
+
+
+spLevelInfo  <-  function(data){
+  spec         <-  data.frame(species=as.character(unique(data$species)), stringsAsFactors=FALSE)
+  for(z in c("family", "pft")){
+    spec[[z]]  <-  as.character(data[[z]][match(spec$species,data$species)])
+    i          <-  spec[[z]]=="" | is.na(spec[[z]]) 
+    if(any(i)){
+      spec[[z]][i]  <-  "????"
+    }
+  }
+  j            <-  spec$species=="" | is.na(spec$species) 
+  if(any(j)){
+    spec$species[j]  <-  "????"
+  }
+  spec
+}
+
+printMeta  <-  function(data){
+  dataset   <-  as.character(unique(data$dataset))
+  openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
+  openMeta
+}
+
+
+#creates html reports using knitr
[…]

### 2013-05-17 sh/add_all.sh

@@ -1,0 +1,7 @@
+# Copy import.csv into all data directories
+for filename in data/* 
+do
+   cp dataNew.csv $filename
+#   mv $filename/data_new.csv $filename/dataNew.csv
+#	touch $filename/dataManipulate.r
+done

### 2013-05-17 sh/cleanCSV.sh

@@ -1,0 +1,12 @@
+# Renames pdf to stadard name format
+#to run type bash sh/renamePdf.sh ls data/   
+cd data
+for filename in *
+do
+   cd $filename
+   for file in *.csv
+   do
+	perl -pi -e 's/\r/\n/g' $file
+   done
+   cd ..
+done

### 2013-05-17 sh/rename.sh

@@ -1,0 +1,6 @@
+# Copy import.csv into all data directories
+for filename in data/* 
+do
+# git mv $filename/extra $filename/xtra
+ git rm $filename/MetadataOLD*
+done

### 2013-05-17 sh/renamePdf.sh

@@ -1,0 +1,9 @@
+# Renames pdf to stadard name format
+#to run type bash sh/renamePdf.sh ls data/   
+cd data
+for filename in *
+do
+   echo $filename.pdf
+   mv $filename/*.pdf $filename/pdf-$filename.pdf
+#   mv $filename/import.txt $filename/import.csv
+done

### 2013-05-17 sh/zipXtra.sh

@@ -1,0 +1,17 @@
+# Renames pdf to stadard name format
+#to run type bash sh/renamePdf.sh ls data/   
+cd data
+for filename in *
+do
+   cd $filename
+   mv extra xtra
+   mkdir xtra
+   mv *.pdf xtra
+   mv *.xls xtra
+   mv *.xlsx xtra
+
+# rm -r -f extra
+   # rm *.xlsx
+   cd ..
+#   mv $filename/import.txt $filename/import.csv
+done

### 2013-05-17 tests/test-import-all.R

@@ -1,0 +1,5 @@
+rm(list=ls())
+source('R/import.R')
+
+# Load all data
+dat <- loadStudies(reprocess=TRUE)

### 2013-05-22 R/import.R
generate email content

@@ -467,15 +467,17 @@ data.path <- function(studyName, ...){
 }
 
 readReference <- function(studyName) {
-  reference <-data.path(studyName, "studyRef.bib")
-  data.frame(dataset=studyName, reference)
+  filename <-data.path(studyName, "studyRef.bib")
+  data.frame(dataset=studyName, filename,  stringsAsFactors=FALSE)
 }
 
 readContact <- function(studyName) {
-  contact <- read.csv(data.path(studyName,"studyContact.csv"),
+  filename <-data.path(studyName, "studyContact.csv")
+  
+  contact <- read.csv( filename,
                       header=TRUE, stringsAsFactors=FALSE,
                       strip.white=TRUE)
-  data.frame(dataset = studyName, contact)
+  data.frame(dataset = studyName, contact, filename,  stringsAsFactors=FALSE)
 }
 
 na.vector <- function(type, n) {

### 2013-05-22 report/report-1-email.R

@@ -1,0 +1,28 @@
+
+generateEmailSubject <-function(study){
+  paste("Biomass and allometry database -", study, ", - error checking") 
+}
+  
+generateEmailContent <-function(to=NA){
+  
+"Dear XXXXX,
+
+You will find attached an individual .html report for the study XXXXX that you provided.
+We would really appreciate if you could answer each of the questions below following the .html report.
+
+1. Is your contact information correct? If not, could you please revise the information in the attached studyContact.csv file?
+2. Is your reference information correct? If not, could you please revise the information in the attached studyContact.csv file? - Please forgive us if your name is mispelled or if it has special characters that are ignored/modified, we are working on a solution.
+3. Does the number of records for each variable seem reasonable based on the datafile you provided?
+4. Is 'Location data' complete? If not, could you please fill in the file dataNew.csv attached with this missing information? Please check the codes and legends for Vegetation_type in Variable.definition.csv file to guide you through.
+5. Are the coordinates/location complete? If the data is missing, could you please provide it by filling in the file dataNew.csv?
+6. Do your locations fall in the right spot in both world and country map?
+7. Is 'Stand description' complete? If not, could you please fill in the missing information in dataNew.csv? If needed, codes and legends for Growing_condition and Status are found within Variable.definition.csv.  
+8. Is your 'Species data' correct? What about plant functional type (pft)? If needed, please correct the data in dataNew.csv and check the relevant codes for pft in Variable.definition.csv.
+9. Are your 'Methods' accurate? Does it contain missing information? Could you please complete the Metadata file attached? Please bear in mind that these Metadata should be a short description (up to 4 lines) of each of the items listed. Additional items not listed in the report are not needed, but if you judge it really important add it to 'Other variables'.
+10. Does your data falls well within the rest of the dataset (please check 'Plots of data')? Are there outliers? If so, could you please revise the units provided in the original dataset and let us know?
+
+Additionally, could you please answer additional questions in the Question file attached and return it to us?
+
+With best regards,
+Daniel Falster"
+}

### 2013-05-22 report/report-byStudy.R

@@ -1,21 +1,19 @@
 rm(list=ls())
-source('R/import.R')
-source('R/plotting.R')
-source('R/formatBib.R')
-source('R/email.R')
 
 source('report/report-fun.R')
 
 # Load all data
 dat <- loadStudies(reprocess=FALSE)
 
-# Markdown report (HTML)
-reportPath<-printStudyReport(dat, "Ilomaki2003")
-
 # All reports:
 reportPaths <- printAllStudyReports(dat)
 
-emailReport("Ilomaki2003", reportPath= reportPath, to = c("daniel.falster@mq.edu.au"), from ="daniel.falster@mq.edu.au", cc= "daniel@falsters.net")
+
+#trial emal
+reportPath<-printStudyReport(dat, "Ilomaki2003")
+emailReport(dat, "Ilomaki2003", reportPath= reportPath)
+
+, to = c("daniel.falster@mq.edu.au"), from ="daniel.falster@mq.edu.au", cc= "daniel@falsters.net")
   
 
 

### 2013-05-22 report/report-1.Rmd
generate email content
Moved report-specific functions into Rmd file

--- 2013-05-17 report/reportmd.Rmd
@@
+```{r, echo=FALSE}
+
+locLevelInfo  <-  function(data){
+
+  vars <- c("location","map","mat","longitude","latitude","vegetation")
+
+  loc <- data[!duplicated(data[,vars]),vars]
+
+  loc[is.na(loc) | loc=="" | loc=="NA"]  <-  "????"  
+  return(loc)
+}
+
+standLevelInfo  <-  function(data){
+
+  vars <- c("location","grouping","growingCondition","status")
+  sta <- data[!duplicated(data[,vars]),vars]
+
+  if(all(is.na(sta$grouping)))
+    sta  <-  sta[,c("location", "growing_condition", "status")]
+
+  sta[is.na(sta) | sta=="" | sta=="NA"]  <-  "????"
+
+  return(sta)
+}
+
+spLevelInfo  <-  function(data){
+  spec         <-  data.frame(species=as.character(unique(data$species)), stringsAsFactors=FALSE)
+  for(z in c("family", "pft")){
+    spec[[z]]  <-  as.character(data[[z]][match(spec$species,data$species)])
+    i          <-  spec[[z]]=="" | is.na(spec[[z]]) 
+    if(any(i)){
+      spec[[z]][i]  <-  "????"
+    }
+  }
+  j            <-  spec$species=="" | is.na(spec$species) 
+  if(any(j)){
+    spec$species[j]  <-  "????"
+  }
+  spec
+}
+
+printMeta  <-  function(data){
+  dataset   <-  as.character(unique(data$dataset))
+  openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
+  openMeta
+}
+
+```
+
 # Report for study: `r .study`
 
-**Reference:** `r formatBib(paste0("../", as.character(.dat$ref$reference)))[[1]]`
+**Reference:** `r formatBib(paste0("../", as.character(.dat$ref$filename)))[[1]]`
 

### 2013-05-22 report/report-fun.R

@@ -104,54 +104,5 @@ knitThis <- function(RmdFile ="report/reportmd.Rmd", path="output/report-per-stu
 }
 
 
-locLevelInfo  <-  function(data){
-  
-  vars <- c("location","map","mat","longitude","latitude","vegetation")
-  
-  loc <- data[!duplicated(data[,vars]),vars]
-  
-  loc[is.na(loc) | loc=="" | loc=="NA"]  <-  "????"  
-  return(loc)
-}
-
-
-standLevelInfo  <-  function(data){
-  
-  vars <- c("location","grouping","growingCondition","status")
-  sta <- data[!duplicated(data[,vars]),vars]
-  
-  
-  if(all(is.na(sta$grouping)))
-    sta  <-  sta[,c("location", "growing_condition", "status")]
-  
-  sta[is.na(sta) | sta=="" | sta=="NA"]  <-  "????"  
-  
-  return(sta)
-}
-
-
-spLevelInfo  <-  function(data){
-  spec         <-  data.frame(species=as.character(unique(data$species)), stringsAsFactors=FALSE)
-  for(z in c("family", "pft")){
-    spec[[z]]  <-  as.character(data[[z]][match(spec$species,data$species)])
-    i          <-  spec[[z]]=="" | is.na(spec[[z]]) 
-    if(any(i)){
-      spec[[z]][i]  <-  "????"
-    }
-  }
-  j            <-  spec$species=="" | is.na(spec$species) 
-  if(any(j)){
-    spec$species[j]  <-  "????"
-  }
-  spec
-}
-
-printMeta  <-  function(data){
-  dataset   <-  as.character(unique(data$dataset))
-  openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
-  openMeta
-}
-
-
 
 

### 2013-05-23 R/plotting.R
Filter points < 0 from log plots

@@ -70,17 +70,25 @@ bivarPlot.Legend <- function(tmp, location="topleft", text.col = "black", pch = 
 
 
 
-bivarPlot <- function(data, xvar, yvar, xlab=xvar, ylab=yvar, col= make.transparent("grey", 0.5), pch=19, add = FALSE, ...){
+bivarPlot <- function(data, xvar, yvar, xlab=xvar, ylab=yvar, col= make.transparent("grey", 0.5), pch=19, add = FALSE, id =c("dataset", "species"), zeroWarning = FALSE, ...){
   
+  #check for negative values  
+  i <-  unique(c( which(data[,xvar] <= 0), which(data[,yvar] <= 0)))
+  
+  if(zeroWarning  & length(i) > 0 ){
+    warning("values <= 0 omitted from logarithmic plot")
+    print(data[i, c(id, xvar, yvar)])
+  }
+    
   if(!add){
-    plot(data[,xvar], data[,yvar],  type= 'n', log="xy", las=1, yaxt="n", xaxt="n", xlab=xlab, ylab=ylab,  ...)
+    plot(data[-i,xvar], data[-i,yvar],  type= 'n', log="xy", las=1, yaxt="n", xaxt="n", xlab=xlab, ylab=ylab,  ...)
     #add nice log axes
     axis.log10(1) 
     axis.log10(2)    
   }
   
   #add data
-  points(data[,xvar], data[,yvar],  type= 'p', col = col, pch=pch, ...)  
+  points(data[-i,xvar], data[-i,yvar],  type= 'p', col = col[-i], pch=pch, ...)  
 }  
 
 whichStudies <- function(alldata, var, value){

### 2013-05-28 talk/Analysis-1.Rmd
Start draft for talk to lab

```{r, echo=FALSE}
source('report/report-fun.R')
options(width =160)

# Load all data
data <- loadStudies(reprocess=FALSE)$data

spLevelInfo  <-  function(data){
  spec         <-  data.frame(species=as.character(unique(data$species)), stringsAsFactors=FALSE)
  for(z in c("family", "pft")){
    spec[[z]]  <-  as.character(data[[z]][match(spec$species,data$species)])
    i          <-  spec[[z]]=="" | is.na(spec[[z]]) 
    if(any(i)){
      spec[[z]][i]  <-  "????"
    }
  }
  j            <-  spec$species=="" | is.na(spec$species) 
  if(any(j)){
    spec$species[j]  <-  "????"
  }
  spec
}

```

## Study details

### Location

**Global map**

```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
prepdata  <-  prepMapInfo(data)
drawWorldPlot (prepdata)
```

### Species data

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
(spLevelInfo(data))

```

## Plots of data

```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
bivarPlot(data, "h.t", "a.lf")

tmp <- bivarPlotColorBy(data, "m.lf", "h.t",colorBy = data$vegetation)
bivarPlot.Legend(tmp)

tmp <- bivarPlotColorBy(data, "a.lf", "h.t",colorBy = data$pft)
bivarPlot.Legend(tmp)

tmp <- bivarPlotColorBy(data, "m.ss", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)

tmp<-bivarPlotColorBy(data, "h.t", "a.lf", colorBy = data$vegetation)
bivarPlot.Legend(tmp)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "m.t", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "m.to", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
source('report/report-fun.R')
tmp <- bivarPlotColorBy(data, "m.st", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "m.ss", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "m.lf", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "m.rt", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "m.sh", "h.t",  colorBy = data$vegetation)
bivarPlot.Legend(tmp)
tmp <- bivarPlotColorBy(data, "a.st", "a.lf",  colorBy = data$vegetation)
tmp <- bivarPlotColorBy(data, "a.ssba", "a.lf",  colorBy = data$vegetation)
tmp <- bivarPlotColorBy(data, "a.stba", "a.lf",  colorBy = data$vegetation)
tmp <- bivarPlotColorBy(data, "a.stbh", "a.lf",  colorBy = data$vegetation)
tmp <- bivarPlotColorBy(data, "m.st", "m.lf",  colorBy = data$vegetation)

```

### 2013-05-28 talk/slides.tex
Start draft for talk to lab

\documentclass[t]{beamer}

\mode<presentation>
{
\usetheme{Singapore}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
%\usepackage{graphicx}
\usepackage{movie15}  % For embedding movies http://pages.uoregon.edu/noeckel/PDFmovie.html
\usepackage[normal,tight,center]{subfigure}
\usepackage{epstopdf}
\usepackage{colortbl}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\setbeamertemplate{navigation symbols}{}  % remove navigation material in footer

\title[] % (optional, use only with long paper titles)
{ }
\author%[D~Falster] % (optional, use only with lots of authors)
{Daniel Falster\inst{1}\\
Collaborators: Remko Duursma\inst{2}, Rich FitzJohn\inst{1}, Diego Barneche\inst{1}, many others }
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different affiliation.

\institute[Macquarie University, Sydney] % (optional, but mostly needed)
{
  \inst{1}%
  Biological Sciences,\\
  Macquarie University, Sydney, Australia
  \and
  \inst{2}%
  Hawkesbury Institue for the Environment, \\
  University of Western Sydney, Australia
}

\date[Labchat 2013] % (optional, should be abbreviation of conference name)
{}

\subject{Theoretical Evolutionary Ecology}

%----------------------------------------------------------------------------------------
\begin{document}
%----------------------------------------------------------------------------------------

\titlepage{}

\section{Problem} \subsection{intro}
\begin{frame}[c]  \frametitle{}
\end{frame}

\end{document}

### 2013-05-28 talk/talk.R

@@ -1,0 +1,36 @@
+source('report/report-fun.R')
+data<- loadStudies(reprocess=FALSE)$data
+
+
+to.pdf(
+  drawWorldPlot (prepMapInfo(data))
+  ,"talk/figures/map.pdf", height= 4, width =6)
+
+to.pdf(
+  highlightStudies(data, "h.t","a.lf", "species", c("Martin1998", "Aiba2005"))
+  ,"talk/figures/H-A.pdf", height= 6, width =8)
+# add  pas used  a1 = 5.44;      B1= 0.306; 
+
+bivarPlotColorBy(data[data$dataset=="Martin1998",], "a.lf", "h.t","species", pch=19)
+
+tmp <- bivarPlotColorBy(data, "a.lf", "h.t", group = "vegetation", pch=19)       
+
+  
+--- 
+
+## Nice plots
+```{r, echo=FALSE, warning=FALSE}
+tmp <- bivarPlotColorBy(data, "m.st", "h.t",colorBy = data$vegetation)
+bivarPlot.Legend(tmp)
+```
+
+--- 
+
+## The end of 3/4 scaling?
+```{r, echo=FALSE, warning=FALSE}
+tmp <- bivarPlotColorBy(data, "m.st", "m.lf",colorBy = data$pft)
+bivarPlot.Legend(tmp)
+```
+
+
+

### 2013-05-28 talk/slides.tex
Talk for labgroup

@@
 %\usepackage{graphicx}
-\usepackage{movie15}  % For embedding movies http://pages.uoregon.edu/noeckel/PDFmovie.html
 \usepackage[normal,tight,center]{subfigure}
@@
 \title[] % (optional, use only with long paper titles)
-{ }
+{A biomass and allometry database for woody plants}
 \author%[D~Falster] % (optional, use only with lots of authors)
-{Daniel Falster\inst{1}\\
+{Daniel Falster\inst{1}
 Collaborators: Remko Duursma\inst{2}, Rich FitzJohn\inst{1}, Diego Barneche\inst{1}, many others }
@@
 
-\section{Problem} \subsection{intro}
+%----------------------------------------------------------------------------------------
+\section{Motivation} \subsection{intro}
 \begin{frame}[c]  \frametitle{}
+\vskip5pt
+{\Huge \color{structure.fg} Challenge: Modelling plant life cycles }
+\end{frame}
+
+%----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{Mechanistic model capturing common processes}
+{\tiny   \color{grey} Falster {\em et al}, J Ecol 2011}
+\begin{center}
+	 \includegraphics[width=\textwidth]{/Users/dfalster/Dropbox/Documents/talks/archive/12_Falster_ESA/analysis/plantModel.pdf}
+\end{center}
+\end{frame}
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{Two assumptions about biomass allocation}
+\begin{center}
+	\includegraphics<1>[width=\textwidth]{/Users/dfalster/Dropbox/Documents/talks/archive/12_Falster_ESA/analysis/output/Assumption-2.pdf}
+\end{center}
+{\tiny   \color{grey} Falster {\em et al}, J Ecol 2011}
+\end{frame}%
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{But - what happens at smaller sizes?}
+\begin{center}
+	\includegraphics<1>[width=\textwidth]{figures/H-A-Aiba-Martin-only.pdf}
+	\includegraphics<2>[width=\textwidth]{figures/H-A-Aiba-Martin-all.pdf}
+\end{center}
+\end{frame}%
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{Problem - few datasets available}
+
+Mostly not suitable 
+\begin{center}
+\begin{itemize}
+		\item Stand avarges (e.g. Niklas \& Enquist; Gentry) \\[\baselineskip] 
+		\item small plants (Poorter 2012)\\[\baselineskip] 
+		\item Multi-species (Chave 2005)\\[\baselineskip] 
+		\item Focus on  DBH and AGB (not so helpful)\\[\baselineskip] 
+	\end{itemize}
+\end{center}
+
+\pause
+But $\ldots$ lots of data in existence
+\end{frame}%
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{Liberating data}
+\begin{center}
+	\includegraphics<1>[width=\textwidth]{figures/H-A-Aiba-Martin-all.pdf}
+\end{center}
+\end{frame}%
+
+%----------------------------------------------------------------------------------------
+\section{Collaboration} \subsection{intro}
+\begin{frame}[c]  \frametitle{}
+\vskip5pt
+{\Huge \color{structure.fg} Idea: A data paper}
+\end{frame}
+
+\begin{frame} [c] \frametitle{Contributions so far}
+\begin{center}
+	\includegraphics[width=0.9\textwidth]{figures/map.pdf}
+\end{center}
+
+{\small \color{red} 86 studies, 14156 individuals, 205 locations, 476 species.}
+\end{frame}
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{Principles}
+\begin{enumerate}
+		\item Recognition for contributors \\$\rightarrow$ coauthorship on data paper (only)\\[\baselineskip] 
+		\item Public data \\$\rightarrow$ open and unrestricted use, citation credit \\[\baselineskip] 
+		\item Transparent data handling \\$\rightarrow$ text and scripts \\[\baselineskip] 
+	\end{enumerate}
+
+\end{frame}
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{}
+\vskip5pt
+{\Huge \color{structure.fg} Mechanics: Scalable and reproducible}
+\end{frame}
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{An R package: dataMashR ??}
+
+\end{frame}
+
+%----------------------------------------------------------------------------------------
+\section{Overview of dataset} \subsection{intro}
+\begin{frame}[c]  \frametitle{}
+\vskip5pt
+{\Huge \color{structure.fg} The dataset}
+\end{frame}
+
+% ----------------------------------------------------------------------------------------
+\begin{frame}[c]  \frametitle{First plots}
+	\includegraphics<1>[width=\textwidth]{figures/plot-1.pdf}
+	\includegraphics<2>[width=\textwidth]{figures/plot-2.pdf}
+	\includegraphics<3>[width=\textwidth]{figures/plot-3.pdf} 
+	\includegraphics<4>[width=\textwidth]{figures/plot-4.pdf}
+	\includegraphics<5>[width=\textwidth]{figures/plot-5.pdf} 
+	\includegraphics<6>[width=\textwidth]{figures/plot-6.pdf}
+	\includegraphics<7>[width=\textwidth]{figures/plot-7.pdf} 
+	\includegraphics<8>[width=\textwidth]{figures/plot-8.pdf}
+	\includegraphics<9>[width=\textwidth]{figures/plot-9.pdf}
+\end{frame}
+
+%----------------------------------------------------------------------------------------
+\section{Future} \subsection{intro}
+\begin{frame}[c]  \frametitle{What next}
+\begin{center}
+\begin{itemize}
+		\item Email report to collabortaors \\[\baselineskip] 
+		\item Draft data paper \\[\baselineskip] 
+		\item Draft dataMashR package and paper \\[\baselineskip] 
+		\item Analysis papers \\[\baselineskip] 
+	\end{itemize}
+\end{center}
 \end{frame}

### 2013-05-28 R/plotting.R

@@ -62,14 +62,14 @@ getLabel<-function(xvar){
   paste0(var.def[i, "label"], " (", var.def[i, "Units"], ")")
 }
 
-highlightStudies <- function(data, xvar, yvar, group, studies, col = niceColors(length(studies)), type='b'){
+highlightStudies <- function(data, xvar, yvar, group, studies, col = niceColors(length(studies)), type='b', pch=1){
   
-  bivarPlotColorBy(data, xvar, yvar, group =group, col=make.transparent("grey", 0.5), legend= FALSE, type = 'p')
+  bivarPlotColorBy(data, xvar, yvar, group =group, col=make.transparent("grey", 0.5), legend= FALSE, type = 'p', pch=pch)
   
   for(i  in seq_len(length(studies))){
     ii <- data$dataset==studies[i]
     if(sum(ii) > 0)
-      bivarPlotColorBy(data[ii,], xvar, yvar, group = group, add = TRUE, legend = FALSE, col=col[i], type=type)  
+      bivarPlotColorBy(data[ii,], xvar, yvar, group = group, add = TRUE, legend = FALSE, col=col[i], type=type, pch=pch)  
   }
   
   leg <- data.frame(group = studies, col = col, stringsAsFactors=FALSE)
@@ -101,7 +101,7 @@ bivarPlotColorBy <- function(data, xvar, yvar, group, type='b', col = make.trans
   out <- out[order(out$group),]
   
   if(legend)
-    bivarPlot.Legend(out, pch=pch)
+    bivarPlot.Legend(out)
   
   if(type %in% c("o","b", "l"))
      add.sma(data, xvar, yvar, colorBy, out, from=from, to=to,...)   
@@ -139,8 +139,8 @@ add.sma <-function (data, xvar, yvar, colorBy, colours, from=NA, to=NA,...){
   invisible(out)
 } 
 
-bivarPlot.Legend <- function(tmp, location="topleft", text.col = "black", pch = 1, lwd=0, bty ="n"){
-  legend(location, tmp$group , col = tmp$col, text.col = text.col, pch = pch, merge = TRUE, lwd=lwd, bty =bty)  
+bivarPlot.Legend <- function(tmp, location="topleft", text.col = "black", lwd=0, bty ="n"){
+  legend(location, tmp$group , col = tmp$col, text.col = text.col,  merge = TRUE, lwd=lwd, bty =bty, pch=19)  
 }
 
 findPositive<-function(data, xvar, yvar){

### 2013-05-28 report/report-bivarPlots.R

@@ -1,14 +1,8 @@
 rm(list=ls())
-source('R/import.R')
+
 source('report/report-fun.R')
 
 #names of all studies
-studyNames     <-  getStudyNames()
+data <- loadStudies(reprocess=FALSE)
 
-#import data
-d  <-  addStudies(studyNames, reprocess = FALSE, verbose = FALSE)
-
-#Explore
-alldata  <-  d$data
-
-comparePlots(alldata)
+comparePlots(data)

### 2013-05-28 report/report-byStudy.R

@@ -3,11 +3,11 @@ rm(list=ls())
 source('report/report-fun.R')
 
 # Load all data
-dat <- loadStudies(reprocess=FALSE)
+dat <- loadStudies(reprocess=TRUE)
 
 # single study
-printStudyReport(dat, "Aiba2007", reprocess=TRUE)
-emailReport(dat, "Aiba2005")
+#printStudyReport(dat, "Aiba2007", reprocess=TRUE)
+#emailReport(dat, "Aiba2005")
 
 # All reports:
 reportPaths <- printAllStudyReports(dat)

### 2013-05-28 talk/talk.R

@@ -2,35 +2,47 @@ source('report/report-fun.R')
 data<- loadStudies(reprocess=FALSE)$data
 
 
+#World map
 to.pdf(
   drawWorldPlot (prepMapInfo(data))
   ,"talk/figures/map.pdf", height= 4, width =6)
 
 to.pdf(
-  highlightStudies(data, "h.t","a.lf", "species", c("Martin1998", "Aiba2005"))
-  ,"talk/figures/H-A.pdf", height= 6, width =8)
-# add  pas used  a1 = 5.44;      B1= 0.306; 
+  highlightStudies(data[data$dataset %in% c("Martin1998", "Aiba2005"), ], "h.t","a.lf", "species", c("Martin1998", "Aiba2005"), pch=19)
+  ,"talk/figures/H-A-Aiba-Martin-only.pdf", height= 6, width =8)
 
-bivarPlotColorBy(data[data$dataset=="Martin1998",], "a.lf", "h.t","species", pch=19)
+to.pdf(
+  highlightStudies(data, "h.t","a.lf", "species", c("Martin1998", "Aiba2005"), pch=19)
+  ,"talk/figures/H-A-Aiba-Martin-all.pdf", height= 6, width =8)
 
-tmp <- bivarPlotColorBy(data, "a.lf", "h.t", group = "vegetation", pch=19)       
+n = 20
+v <- data.frame(x = rep("",n), y= rep("",n), g = rep("",n), stringsAsFactors=FALSE)
+v[1,] <- c("h.t", "a.lf", "pft") 
+v[2,] <- c("h.t", "a.lf", "vegetation") 
+v[3,] <- c("h.t", "m.lf", "vegetation") 
+v[4,] <- c("h.t", "m.st", "vegetation") 
+v[5,] <- c("m.lf", "m.st", "vegetation") 
+v[6,] <- c("a.lf", "m.st", "vegetation") 
+v[7,] <- c("a.stbh", "m.st", "vegetation") 
+v[8,] <- c("a.stbh", "a.lf", "vegetation") 
+v[9,] <- c("a.ssbh", "a.lf", "vegetation") 
 
-  
---- 
 
-## Nice plots
-```{r, echo=FALSE, warning=FALSE}
-tmp <- bivarPlotColorBy(data, "m.st", "h.t",colorBy = data$vegetation)
-bivarPlot.Legend(tmp)
-```
+for(i in seq_len(9)){
+  to.pdf(
[…]

### 2013-06-07 report/report-1.Rmd
Refine Emial and report

--- 2013-05-17 report/reportmd.Rmd
@@
+```{r, echo=FALSE}
+
+locLevelInfo  <-  function(data){
+
+  vars <- c("location","longitude","latitude","vegetation")
+
+  loc <- data[!duplicated(data[,vars]),vars]
+
+  loc[is.na(loc) | loc=="" | loc=="NA"]  <-  "missing"  
+  return(loc)
+}
+
+standLevelInfo  <-  function(data){
+
+  vars <- c("location","grouping","growingCondition","status")
+  sta <- data[!duplicated(data[,vars]),vars]
+
+  if(all(is.na(sta$grouping)))
+    sta  <-  sta[,-which(names(sta) == "grouping")]
+
+  sta[is.na(sta) | sta=="" | sta=="NA"]  <-  "missing"
+
+  return(sta)
+}
+
+spLevelInfo  <-  function(data){
+  spec         <-  data.frame(species=as.character(unique(data$species)), stringsAsFactors=FALSE)
+  for(z in c("family", "pft")){
+    spec[[z]]  <-  as.character(data[[z]][match(spec$species,data$species)])
+    i          <-  spec[[z]]=="" | is.na(spec[[z]]) 
+    if(any(i)){
+      spec[[z]][i]  <-  "missing"
+    }
+  }
+  j            <-  spec$species=="" | is.na(spec$species) 
+  if(any(j)){
+    spec$species[j]  <-  "missing"
+  }
+  spec
+}
+
+printMeta  <-  function(data){
+  dataset   <-  as.character(unique(data$dataset))
+  openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
+  openMeta
+}
+
+```
+
 # Report for study: `r .study`
 
-**Reference:** `r formatBib(paste0("../", as.character(.dat$ref$reference)))[[1]]`
+## Contact Information
 
@@
 **Address:** `r .dat$contact$address`.
+
+## Citation
+Information about the primary reference for the data you provided. 
+```{r echo=FALSE, comment=''}
+  bibfile<- paste0("../", as.character(.dat$ref$filename))
+```
+**Reference:** `r formatBib(bibfile)[[1]]`
+
+**DOI:** `r  bibGetElement(bibfile, element = "doi")`
+
+**Abstract:** `r  bibGetElement(bibfile, element = "abstract")`
 
@@
 **Stand description**
+Information about the growing condition of plants in your study, i.e. their local environment.
 
@@
 
-This is how the study `r .study` (each color represents a species) fits in the entire dataset (grey).
+This is how the study `r .study` (each colour represents a species) fits in the entire dataset (grey).
 

### 2013-06-07 report/report-1-email.R

@@ -13,10 +13,7 @@ getEmailDetails <- function(alldata, study, reprocess= FALSE){
        content = generateEmailContent(alldata, study, dat),
        filesToSend = c(reportPath, 
                        makeQuestionsFile(study),
-                       dat$contact$filename,
-                       dat$ref$filename, 
                        file.path(data.path(study, "studyMetadata.csv")),
-                       generateDataNew(alldata, study),
                        file.path(data.path(study, "data.csv")),
                        paste0("output/data/", study, ".csv"),
                        "config/variableDefinitions.csv"
@@ -35,13 +32,15 @@ generateEmailContent <-function(allData, study, studyData){
 
 Last year you kindly agreed to contribute data to a biomass and allometry database for woody plants, for publication as a data paper in the journal Ecology.  Overall we had a fantastic response to our request for data, with data now collected from ", length(unique(allData$ref$dataset)), " studies, with information for ", length(allData$data$dataset), " individuals from ", length(unique(allData$data$location)), " locations, covering ",length(unique(allData$data$species)), " species. We are certain this dataset will be a widely used resource and thank you for your contribution.
 
-The purpose of this email is to provide you with a report on the data you contributed to the study. The database is organised by publication, so if you contributed data from more than one publication, you will receive multiple emails. We kindly request that you review each of these before we submit our paper for publication, to ensure all the information is correct. Unless there are major issues with your data, it should only take 30 mins to complete each review. 
+The purpose of this email is to provide you with a report on the data you contributed to the study. The database is organised by publication, so if you contributed data from more than one publication, you will receive multiple emails. We kindly request that you review each of these before we submit our paper for publication, to ensure all the information is correct. Unless there are major issues with your data, it should only take a few minutes to complete each review.
 
 This email contains a report for the following study, for which you are listed as contact person: ",
         formatBib(studyData$ref$filename)[[1]],
          ".
 
-To assist us in preparing the data for publication, we ask that you review the attached report of your data and provide answers to ALL of the questions listed in the file 'report1-questions.txt' attached. We require all questions be answered for a study to be included in the final dataset. Please type your answers directly into that file or into one of the other attached files, where appropriate.  By returning the requested information in this way, you are assisting us in handling a large number of files and corrections.
[…]

### 2013-06-07 report/report-byStudy.R

@@ -6,15 +6,20 @@ source('report/report-fun.R')
 dat <- loadStudies(reprocess=TRUE)
 
 # single study
-printStudyReport(dat, "Aiba2007", reprocess=TRUE)
-printStudyReport(dat, "Ishihara0000", reprocess=TRUE)
+printStudyReport(dat, "Aiba2007", reprocess=FALSE)
+printStudyReport(dat, "Lusk2011", reprocess=FALSE)
 
-#emailReport(dat, "Aiba2005")
+printStudyReport(dat, "Ishihara0000", reprocess=TRUE)
+emailReport(dat, "Gargaglione2010")
 
 # All reports:
 reportPaths <- printAllStudyReports(dat)
 
+
 # Reprocess one study
 tmp <- loadStudy("Gargaglione2010", reprocess= TRUE)
 dat <- loadStudies(reprocess=FALSE)
+
+
+data = extractStudy(dat, "Bond-Lamberty2002")$data
 

### 2013-06-26 report/report-1.Rmd
'status' removed from stand variables in report Rmd file.
Gloabl vars removed, var.def and dir.rawData now accessed via function mashrDetail("..")

@@
 
-  vars <- c("location","grouping","growingCondition","status")
+  vars <- c("location","grouping","growingCondition") #,"status")
   sta <- data[!duplicated(data[,vars]),vars]
@@
   dataset   <-  as.character(unique(data$dataset))
-  openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
+  openMeta  <-  read.csv(paste0("../",mashrDetail("dir.raw"),"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
   openMeta
@@
 counts  <-  counts[counts>0]  #only include no zeros
-counts  <-  counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
+counts  <-  counts[names(counts)%in%mashrDetail("var.def")$Variable[mashrDetail("var.def")$Group=="tree"]]
 cbind(counts)

### 2013-06-26 R/plotting.R

@@ -28,7 +28,7 @@ makePlotPanel<-function(data, study, dir="report-per-study", col="grey", pdf=TRU
   }
   
   dat        <-  data[data$dataset==study,]
-  plot.vars  <-  var.def$Variable[var.def$Group=="tree"]
+  plot.vars  <-  mashrDetail("var.def")$Variable[mashrDetail("var.def")$Group=="tree"]
   plot.vars  <-  plot.vars[plot.vars %in% c("growingCondition","status","light") == FALSE]
   available  <-  dat[,names(dat) %in% plot.vars]   
   available  <-  apply(available,2,as.numeric)
@@ -50,8 +50,8 @@ makePlotPanel<-function(data, study, dir="report-per-study", col="grey", pdf=TRU
       if(length(testNa[!is.na(testNa)]) != 0){
         z  <-  z + 1
         if(z==5){par(mfrow=c(2,2)); z  <- 1}
-        varjx  <-  paste0(var.def$label[var.def$Variable==colnames(available)[j]], " (", var.def$Units[var.def$Variable==colnames(available)[j]], ")")
-        varjy  <-  paste0(var.def$label[var.def$Variable==colnames(available)[k]], " (", var.def$Units[var.def$Variable==colnames(available)[k]], ")")
+        varjx  <-  paste0(mashrDetail("var.def")$label[mashrDetail("var.def")$Variable==colnames(available)[j]], " (", mashrDetail("var.def")$Units[mashrDetail("var.def")$Variable==colnames(available)[j]], ")")
+        varjy  <-  paste0(mashrDetail("var.def")$label[mashrDetail("var.def")$Variable==colnames(available)[k]], " (", mashrDetail("var.def")$Units[mashrDetail("var.def")$Variable==colnames(available)[k]], ")")
         makePlot(data, available, studycol=colorvec, xvar = colnames(available)[j], yvar = colnames(available)[k],  xlab=varjx, ylab=varjy)
       }
     }
@@ -61,8 +61,8 @@ makePlotPanel<-function(data, study, dir="report-per-study", col="grey", pdf=TRU
 }
 
 getLabel<-function(xvar){
-  i<-which(var.def$Variable==xvar)
-  paste0(var.def[i, "label"], " (", var.def[i, "Units"], ")")
+  i<-which(mashrDetail("var.def")$Variable==xvar)
+  paste0(mashrDetail("var.def")[i, "label"], " (", mashrDetail("var.def")[i, "Units"], ")")
 }
 
[…]

### 2013-06-26 report/report-1-email.R

@@ -229,7 +229,7 @@ generateDataNew  <-  function(alldata, studyName, levels=c("site","species","tre
 
 checkLevels  <-  function(site.data, group){
   #get essential variables for the chosen group
-  chosen  <-  var.def$Variable[var.def$Group==group & var.def$essential==TRUE]
+  chosen  <-  mashrDetail("var.def")$Variable[mashrDetail("var.def")$Group==group & mashrDetail("var.def")$essential==TRUE]
   #find NA's
   nas     <-  apply(site.data[names(site.data) %in% chosen], 2, function(x){which(is.na(x))})
   if(class(nas) == "matrix"){

### 2013-06-26 report/report-byStudy.R

@@ -3,8 +3,10 @@ rm(list=ls())
 source('report/report-fun.R')
 source('report/report-1-email.R')
 
+mashrDetail("var.def")
+
 # Load all data
-dat <- loadStudies(reprocess=TRUE)
+dat <- loadStudies(reprocess=FALSE)
 
 # single study
 printStudyReport(dat, "Aiba2007", reprocess=FALSE)

### 2013-10-03 website/index.Rmd
Send invites to new studies

```{r, results="hide", echo=FALSE, warnings=FALSE}
opts_chunk$set(warnings = FALSE)
```

# A Biomass and Allometry Database for Woody Plants

# Invitation to participate
Thank you for showing interest in our project. Our aim is to submit a co-authored 'Data paper' to the journal Ecology. You are hereby invited to contribute. Your contribution will help build an important resource for the research on plant growth, ecosystem dynamics and carbon cycling in the worlds vegetation.

Below follows:

* a draft abstract for a co-authored "data paper" to Ecology. A data paper consists of an abstract, metadata and data. An example can be found [here](A Biomass and Allometry Database for Woody Plants).
* instructions for contributing data, with further details regarding participation and information on focal variables.

Any questions or correspondence should be directed to:

Dr Daniel Falster - daniel.falster@mq.edu.au - ARC Postdoctoral Research Fellow, Department of Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia

Dr Remko Duursma - remkoduursma@gmail.com - Lecturer, Hawkesbury Institue for the Environment, University of Western Sydney

# Draft abstract

The allocation of primary production (biomass) between different plant tissues (leaf, sapwood, bark, roots, reproductive material) determines the nature of plant growth, the outcome of competitive interactions and the flow of energy in the terrestrial biosphere. Biomass allocation is thought to change along environmental gradients and also to differ among coexisting species, such that individuals within a species exhibit a typical trajectory of biomass deployment through ontogeny. Despite decades of interest in the allometry of plant growth, few data sets enable the growth-trajectories of different species to be compared with one another.

Here we present an individual-level data set capturing ontogenetic growth trajectories in woody plant species from across the globe. The data were compiled from published and unpublished sources, focusing on studies where: i) measurements were for individual plants rather than whole stands; ii) individuals spanning a range of sizes were measured, ideally from seed to mature plant; and iii) size was measured directly, i.e. all biomass measurements involve a complete or partial destructive harvest. We included both plants grown in their natural habitat and in artificial plantings. The data set contains the following variables: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; total plant height; height of crown base; projected crown area; lateral crown surface area; and total dry mass of leaf, stem, branch, sapwood, heartwood, bark, coarse root and fine root tissues. We also report other properties of individuals, where available (plant age, mean leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (status, light, experimental treatment, site leaf area index, vegetation type, mean annual rainfall). Overall, the data set contains xxx measurements, from xxx individuals, xxx species, and xxx sites.

Our size and biomass data set represents a vast, untapped scientific resource characterising the growth trajectories of  xxxx woody plant species from across the globe. We hope that compiling this data set and making it available will advance our understanding of plant growth, ecosystem dynamics and carbon cycling in the worlds vegetation.

# Instructions

This study aims to characterise the typical growth trajectory of a species through ontogeny. We are interested in a number of variables (see table below), but as a minimum:
 - you must have recorded ALL of i) total leaf area or mass, ii) plant height, and iii) stem diameter (or cross sectional area) at base, breast height or base of crown.
 - variables must have been measured directly and compiled through partial or complete destructive harvest, i.e. no allometric equations.
 - variables exist for a number of individuals spanning a range of sizes for each species.
 - individuals sampled for each species come from the same area of forest or experiment.

The following table outlines all the  variables we are collecting, and any equivalent (related) variables. You do not need all of these to participate, indeed we expect only a few papers will have the complete set, but the more the better.

If your data does not have the same units or variable names, simply send the data in the units as they appear in your paper. We will later convert it to these preferred dimensions.

## List of variables included in the study

```{r,results="hide",echo=FALSE}
	data <- read.csv("../config/variableDefinitions.csv", stringsAsFactors=FALSE)
	data$Units <- gsub(".", "/", data$Units, fixed = TRUE)
	text <- ascii(data[, c("Variable","Units", "Description")] , include.rownames = FALSE)
```
```{r,results="asis",echo=FALSE}
	print(text, type = 'pandoc')
```

### 2013-10-03 report/invitation.R

@@ -1,0 +1,58 @@
+rm(list=ls())
+
+source('R/email.R')
+
+sendInvitation <- function(dat){
+  cols <- names(dat)[10:57]
+  vars <- dat=='x' & names(dat) %in% cols
+  collected <- paste(names(dat)[vars], collapse = ", ")
+
+  email.text  =  paste0("Dear ", dat[["Contact"]],",
+
+We are writing to invite you to co-author a 'Data paper' in the journal Ecology, gathering together size and biomass data for woody plants from across the globe. Literally tens of thousands of plants have been measured or harvested in different experiments and studies; gathered together such data would have enormous value and represent an important resource for the community.
+
+We are contacting you because we read your ", dat[["Year"]]," paper '", dat[["Title"]],"' in the journal '", dat[["Journal"]],"', in which you measured ", collected, " for ", dat[["Individuals"]]," individuals from ", dat[["Species"]]," species. This data seems highly suitable for inclusion in this project.
+
+We have already collated data from almost 100 studies, covering over 400 species and 10,000 individuals. We are contacting you as part of our last round of recruitment for the study, having identified your study as making a quality contribution to the database, in addition to the datasets already included.
+
+Why should I participate?
+  * We are planning to submit a 'Data paper' to the journal Ecology. The dataset will be open access and we expect the paper to be widely cited. All data contributors will be included as co-authors, so for little additional effort you could add a well-cited Ecology paper to your CV.
+  * By participating you ensure your data remains available (via the ESA archives) into the future, thereby increasing the impact of your work.
+  * Once the dataset has been compiled, we will send you a report showing how your data compares to other studies. This will give additional insight into your study system.
+
+Who will use the data?
+  * The dataset will be open-access. Once published, anyone can pursue and publish his or her own analyses, including you.
+  * Beyond regular papers, we expect many groups developing vegetation models will use the data set to parameterise and test parts of their model.
+
+Should I worry about sending you my data?
+  * No. We have intentionally chosen a publication model that recognises the hard work of data collectors.  By making you a co-author on the data paper, we ensure you receive proper credit (via citation) anytime the data is used. By making the dataset open to the public, we hope to prevent anyone from thinking we have ownership of your data.
[…]

### 2014-01-31 analysis/Outline.rmd
outline is now an rmd

% How to build a plant
% Daniel Falster; Remko Duursma; Diego Barneche

Goal: Establish foundations for modelling plant growth through ontogeny, emphasising simple hypotheses, areas where important differences lie (e.g. between species, between vegetation, light env)

# Relationships to test

## An allometric relation between total leaf area and plant height

- is there linear log-log relationship between leaf area and height across size range?
- slope seems to increase in the upper range of heights, leaf area and (less distinctly) leaf mass increase more steeply relative to height gain. Or to put that the other way round, height gain slows even though canopy expansion is continuing
  - how much of this is higher SLA in the taller plants?
  - how much of it is a between species rather than a within-species effect?
  - it should be apparent also in the relation between stem or sapwood cross

- how much variation between species within sites; species across sites; vegetation types; light env within sites?

## The pipe model, relating cross-sectional area of sapwood and bark to total leaf area.

- is there linear log-log relationship across size range between sapwood cross section and total leaf area?
  - for smaller stems: use total area

- expect slope different to 1 for total stem cross section, because incorporates heartwood
  - better with stem area at crown base (as per Makela hypothesis)

- is there linear log-log relationship across size range between bark cross section and total leaf area?
  - for smaller stems: use total area

- how much variation between species within sites; species across sites; vegetation types; light env within sites?

## Ratio of root mass to leaf area.

- is there linear log-log relationship across size range between root mass and total leaf area?

- how much variation between species within sites; species across sites; vegetation types; light env within sites?

## Crown width and surface area

- is there linear log-log relationship across size range between root mass and total leaf area?

## Leaf mass fraction

## Equations for total biomass

- stem biomass predicted from height & traits via simple equation:
- leaf biomass?

### 2014-01-31 analysis/Outline.rmd
outline.rmd has leaf mass fraction figure

@@
-% How to build a plant
-% Daniel Falster; Remko Duursma; Diego Barneche
+How to build a plant
+=======================
+Daniel Falster, Remko Duursma, Diego Barneche
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+source("../R/fillDerivedVariables.R")
+source("../R/import.R")
+library(gplots)
+library(scales)
+library(magicaxis)
+
+if(!file.exists("../cache/allomdata.rds")){
+  dat <- loadStudies(reprocess=FALSE)
+  data <- fillDerivedVariables(dat$data)
+  saveRDS(data, file="cache/allomdata.rds")
+} else {
+  data <- readRDS("../cache/allomdata.rds")
+}
+
+```
 
@@
 
+```{r echo=FALSE}
+# subset/refine data.
+dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & pft != "" & m.lf > 0))
+dfr <- within(dfr, {
+  dataset <- as.factor(dataset)
+  species <- as.factor(species)
+  pft <- as.factor(pft)
+  m.lf.fraction <- with(dfr, m.lf / m.so)
+  log_m.so <- log10(m.so)
+  log_m.lf.fraction <- log10(m.lf.fraction)
+})
+
+with(dfr, plot(log_m.so, log_m.lf.fraction, axes=FALSE,
+               xlab="Aboveground biomass (kg)",
+               ylab="Leaf mass fraction (-)",
+               pch=19, col=pft
+))
+magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
+box()
+legend("topright", levels(dfr$pft), pch=19, col=palette())
+
+```
+
 ## Equations for total biomass

### 2014-01-31 analysis/Outline.rmd
Biomass fractions for root, leaf stem added to outline.rmd

@@
 legend("topright", levels(dfr$pft), pch=19, col=palette())
+```
 
+```{r echo=FALSE}
+dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & !is.na(m.rt) & 
+                           m.lf > 0 & m.rt > 0 & pft != "" & 
+                           m.st < m.to))
+dfr$m.to <- with(dfr, m.rt + m.st + m.lf)
+
+dfr$dataset <- as.factor(dfr$dataset)
+dfr$species <- as.factor(dfr$species)
+dfr$pft <- as.factor(dfr$pft)
+dfr$m.lf.fraction <- with(dfr, m.lf / m.to)
+dfr$m.st.fraction <- with(dfr, m.st / m.to)
+dfr$m.rt.fraction <- with(dfr, m.rt / m.to)
+
+par(mfrow=c(2,2))
+plotit <- function(Pft, dataset){
+  x <- subset(dataset, pft == Pft)
+  with(x, plot(log10(m.to), m.rt.fraction, col="red", ylim=c(0,1.1),
+               xlab="Total biomass (kg)",
+               ylab="Biomass fractions (-)",
+               main=Pft,
+               axes=FALSE))
+
+  magaxis(side=1, unlog='x', tcl=-0.4)
+  axis(2, at=seq(0,1,by=0.2))
+  box()
+  with(x, points(log10(m.to), m.rt.fraction + m.lf.fraction, col="blue"))
+  abline(h=1, lwd=2, col="forestgreen")
+}
+for(z in c("DA","EA","EG"))plotit(z,dfr)
+
+plot(1, type='n', ann=FALSE, axes=FALSE)
+legend("left", c("Root","Root+Leaf","Root+Leaf+Stem"), pch=c(19,19,-1),
+                 lty=c(-1,-1,1), col=c("red","blue","forestgreen"), cex=0.8)
 ```

### 2014-01-31 analysis/Outline.rmd
Functional balance (fine root - leaf mass) in outline.rmd

@@
 
+```{r echo=FALSE}
+dfr <- subset(data, !is.na(m.rf) & !is.na(m.lf) & pft != "")
+dfr$pft <- as.factor(dfr$pft)
+
+palette(rich.colors(4))
+with(dfr, plot(log10(m.lf), log10(m.rf),
+               axes=FALSE,
+               xlab="Leaf biomass (kg)",
+               ylab="Fine root biomass (kg)",
+               pch=19,
+               col=pft))
+zs <- c(0.05, 0.25, 0.5, 0.75,1)
+for(z in zs)abline(log10(z),1)
+legend("bottomright", levels(dfr$pft), pch=19, col=palette())
+magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
+box()
+```
+Figure X. Lines are with exponent of 1, and represent fixed ratios between fine root biomass and leaf biomass (from top to bottom: 1, 0.75, 0.5, 0.25, 0.05).
+
 - how much variation between species within sites; species across sites; vegetation types; light env within sites?
-
-## Crown width and surface area
-
-- is there linear log-log relationship across size range between root mass and total leaf area?
 
@@
 ```
+Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.
 

### 2014-01-31 analysis/Outline.rmd
updated scaling of fine root vs. leaf mass, and leaf mass fraction of abvground biomass

@@
 }
+data$dataset_species <- as.factor(paste(data$dataset, data$species, sep="-"))
 
@@
 
-palette(rich.colors(4))
+sm <- lapply(split(dfr, dfr$pft), function(x){
+      sma(log10(m.rf) ~ log10(m.lf)*dataset_species, 
+           data=x)
+      })
+
+palette(alpha(rich.colors(4),0.5))
 with(dfr, plot(log10(m.lf), log10(m.rf), 
                axes=FALSE,
-               xlab="Leaf biomass (kg)",
-               ylab="Fine root biomass (kg)",
-               pch=19,
-               col=pft))
-zs <- c(0.05, 0.25, 0.5, 0.75,1)
-for(z in zs)abline(log10(z),1)
-legend("bottomright", levels(dfr$pft), pch=19, col=palette())
+               xlab="Aboveground biomass (kg)",
+               ylab="Foliage mass (kg)",
+               type='n',col=pft, panel.first={
+                 for(z in c(0.05, 0.25, 0.5, 0.75, 1))abline(log10(z),1)
+               }))
 magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
+for(i in 1:3)plot(sm[[i]], add=TRUE, col=palette()[i], type='l', lwd=2)
 box()
+legend("bottomright", levels(dfr$pft), fill=alpha(palette(),1))
+
 ```
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, eval=FALSE}
 # subset/refine data.
@@
 legend("topright", levels(dfr$pft), pch=19, col=palette())
+#Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.
 ```
-Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.
 
-```{r echo=FALSE}
+```{r echo=FALSE, eval=FALSE}
 dfr <- droplevels(subset(data, !is.na(m.st) & !is.na(m.lf) & !is.na(m.rt) & 
@@
 
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+dfr <- droplevels(subset(data, !is.na(m.so) & !is.na(m.lf) & pft != "" &
+                           m.lf > 0 & m.so > 0))
+dfr$pft <- as.factor(dfr$pft)
+
+sm <- lapply(split(dfr, dfr$pft), function(x){
+      sma(log10(m.lf) ~ log10(m.so)*dataset_species, 
+           data=x)
+      })
+
+palette(alpha(rich.colors(4), 0.4))
+with(dfr, plot(log10(m.so), log10(m.lf), 
+               axes=FALSE,
+               xlab="Aboveground biomass (kg)",
+               ylab="Foliage mass (kg)",
+               type='n',col=pft, panel.first={
+                 for(z in c(-2,-1,0))abline(z,1)
+
+               }))
+magaxis(side=c(1,2), unlog='xy', tcl=-0.4)
+for(i in 1:4)plot(sm[[i]], add=TRUE, col=palette()[i], type='l')
+box()
+legend("bottomright", levels(dfr$pft), fill=alpha(palette(),1))
+
+```
+
 ## Equations for total biomass

### 2014-01-31 analysis/Outline.rmd
Fixed glitches in outline.rmd

@@
                axes=FALSE,
-               xlab="Aboveground biomass (kg)",
-               ylab="Foliage mass (kg)",
+               xlab="Leaf mass (kg)",
+               ylab="Fine root mass (kg)",
                type='n',col=pft, panel.first={
@@
 
-```{r echo=FALSE, eval=FALSE}
+```{r echo=FALSE, eval=TRUE}
 # subset/refine data.
@@
 legend("topright", levels(dfr$pft), pch=19, col=palette())
-#Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.
+
 ```
+Figure X. Much higher leaf mass fraction for evergreen gymnosperms at a given total aboveground plant biomass.
 

### 2014-01-31 analysis/Outline.rmd
smatr messages gone

@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+if(packageVersion("smatr") < "3.4-4")
+  stop("Please update smatr! We use the new quiet argument.")
 source("../R/fillDerivedVariables.R")
@@
 sm <- lapply(split(dfr, dfr$pft), function(x){
-      sma(log10(m.rf) ~ log10(m.lf)*dataset_species, 
+      sma(log10(m.rf) ~ log10(m.lf)*dataset_species, quiet=TRUE,
            data=x)
@@
 sm <- lapply(split(dfr, dfr$pft), function(x){
-      sma(log10(m.lf) ~ log10(m.so)*dataset_species, 
+      sma(log10(m.lf) ~ log10(m.so)*dataset_species, quiet=TRUE,
            data=x)

### 2014-03-21 analysis/stem_leaf_scaling.Rnw
Stem-leaf scaling in knitr-latex (Rnw) in analysis/.

\documentclass{article}

\begin{document}

\title{Scaling of leaf and above-ground woody biomass}
\maketitle

<<load, echo=FALSE, message=FALSE>>=
setwd("..//")
dat <- readRDS("cache/allomdata_post.rds")

library(smatr)
#library(tables)
library(magicaxis)
#library(Hmisc)
#library(MIfuns)
source("R/remko_functions.R")
source("R/plotting.R")
@

<<prep, echo=FALSE>>=
# Stem-leaf scaling

# Select datasets that have stem, leaf biomass, and height
x <- suppressMessages(studyWithVars(dat, c("m.st","m.lf","h.t")))
x$dataset_species <- paste0(x$dataset,"_",x$species)

# clean
x <- subset(x, m.lf > 0 & m.st > 0 & !is.na(m.lf) & !is.na(m.st) & !is.na(h.t))

# Keep only dataset_species with at least 20 observations
tab <- table(x$dataset_species)
x <- x[x$dataset_species %in% names(tab)[tab>19],]
@

The selected data has \Sexpr{nrow(x)} rows, and \Sexpr{length(unique(x$dataset_species))} unique dataset-species combinations.

<<echo=FALSE>>=
xtabs(~ pft + vegetation,data=x)
@

<<analysis, echo=FALSE>>=
sm_mlf <- sma(m.lf ~ h.t * dataset_species, log="xy", data=x)
sm_mst <- sma(m.st ~ h.t * dataset_species, log="xy", data=x)

# sma coefficients
makedfr <- function(z){
  z$dataset_species <- rownames(z)
  as.data.frame(z)
}
p_mlf <- coef(sm_mlf)
names(p_mlf) <- c("b0_mlf","b1_mlf")
p_mst <- coef(sm_mst)
names(p_mst) <- c("b0_mst","b1_mst")
p <- merge(makedfr(p_mlf),makedfr(p_mst))

# Find ranges for h.t
r <- t(sapply(p$dataset_species, function(lev){
    range(x$h.t[x$dataset_species==lev], na.rm=TRUE)
}))
colnames(r) <- c("ht_min","ht_max")

# P is a dataframe with leaf, stem scaling, and ranges of h.t
P <- cbind(p,r)
rownames(P) <- NULL

# Add some metadata
dfr <- dat[,c("dataset","species","pft","vegetation")]
dfr$dataset_species <- with(dfr, paste0(dataset,"_",species))
dfr <- dfr[!duplicated(dfr$dataset_species),]
P <- merge(P, dfr)
P$pft <- as.factor(P$pft)
P$vegetation <- as.factor(P$vegetation)

# Plot : save predictions in a list
l <- lapply(1:nrow(P), function(i){
  h <- seq(P$ht_min[i], P$ht_max[i], length=101)
  a <- 10^(P$b0_mlf[i])
  b <- P$b1_mlf[i]
  c <- 10^(P$b0_mst[i])
  d <- P$b1_mst[i]
  wfrel <- a*h^b/(a*h^b + c*h^d)
  mst <- c*h^d
  mlf <- a*h^b
  return(data.frame(h=h,wfrel=wfrel,mlf=mlf,mst=mst))
})

@

<<plotting, echo=FALSE>>=
# natural scale
plot(1, xlim=c(0,50), ylim=c(0,1), type='n',
     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)")
for(i in 1:length(l))points(l[[i]]$h, l[[i]]$wfrel, type='l')

# log-log scale
plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
     axes=FALSE)
addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l')

# log-natural scale
plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
     axes=FALSE)
addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
axis(2)
for(i in 1:length(l))points(log10(l[[i]]$h), l[[i]]$wfrel, type='l')

# Coloured by pft

# log-log scale
palette(c("blue","royalblue","red","darkorange"))
plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
     axes=FALSE)
addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
for(i in 1:length(l)){
  points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l',
         col=P$pft[i])
}
legend("bottomleft", c("Deciduous angiosperm","Deciduous gymnosperm",
                       "Evergreen angiosperm","Evergreen gymnosperm"),
       fill=palette())
# dev.copy2pdf(file="analysis/output/figures/LMF_height_loglog.pdf")

# log-log scale
plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
     axes=FALSE)
addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
axis(2)
for(i in 1:length(l)){
  points(log10(l[[i]]$h), l[[i]]$wfrel, type='l',
         col=P$pft[i])
}
legend("bottomleft", c("Deciduous angiosperm","Deciduous gymnosperm",
                       "Evergreen angiosperm","Evergreen gymnosperm"),
       fill=palette())
# dev.copy2pdf(file="analysis/output/figures/LMF_height_lognat.pdf")

# m.lf, m.st vs. h.t by pft.
plot(1, xlim=c(-3,2), ylim=c(-7,5), type='n',
     xlab="Plant height (m)", ylab="Leaf mass (kg)",
     axes=FALSE)
magaxis(side=1:2, unlog='xy',box=TRUE)
for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$mlf), type='l',
                            col=P$pft[i])
legend("topleft", c("Deciduous angiosperm","Deciduous gymnosperm",
                       "Evergreen angiosperm","Evergreen gymnosperm"),
       fill=palette(), inset=0.04)
# dev.copy2pdf(file="analysis/output/figures/MF_height_loglog.pdf")

plot(1, xlim=c(-3,2), ylim=c(-7,5), type='n',
     xlab="Plant height (m)", ylab="Above-ground woody mass (kg)",
     axes=FALSE)
magaxis(side=1:2, unlog='xy',box=TRUE)
for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$mst), type='l',
                            col=P$pft[i])
legend("topleft", c("Deciduous angiosperm","Deciduous gymnosperm",
                    "Evergreen angiosperm","Evergreen gymnosperm"),
       fill=palette(), inset=0.04)
# dev.copy2pdf(file="analysis/output/figures/MS_height_loglog.pdf")
@

\end{document}

### 2014-03-21 R/load.R

@@ -1,6 +1,7 @@
 rm(list=ls())
 source('R/import.R')  
 
+
 #identify new studies
 newStudies  <-  identifyNewStudies()
 if(length(newStudies) > 0){
@@ -15,6 +16,7 @@ d <- loadStudies(reprocess=TRUE)
 
 
 # Post-process.
+source('R/remko_functions.R')
 dat <- postProcessAllom(d$data)
 #saveRDS(dat, "cache/allomdata_post.rds")
 

### 2014-03-21 analysis/metadata_and_tables.R

@@ -1,5 +1,6 @@
 
 source("R/plotting.R")
+source("R/remko_functions.R")
 # dat <- readRDS("cache/allomdata_post.rds")
 
 # Sample size
@@ -89,33 +90,6 @@ with(datEA, plot(log10(m.rt), log10(m.lf), pch=19,cex=0.8, col=TempTrop))
 #-------------------------------------------------------------------------------------------#
 
 
-# Allometry plot as in Wolf
-
-x <- studyWithVars(dat, c("m.st","m.lf","h.t"))
-
-# how many species per study
-nspecies <- sapply(x, function(arg)length(unique(arg$species)))
-this1 <- nspecies==1
-
-y <- x[[this1[1]]]
-y <- y[complete.cases(y),]
-
-
-sm_mlf <- sma(m.lf ~ h.t, log="xy", data=y)
-sm_mst <- sma(m.st ~ h.t, log="xy", data=y)
-
-p_mlf <- coef(sm_mlf)
-p_mst <- coef(sm_mst)
-
-r <- range(y$h.t)
-ht <- seq(r[1],r[2],length=101)
-
-mlfs <- 10^(p_mlf[1] + p_mlf[2]*ht)
-msts <- 10^(p_mst[1] + p_mst[2]*ht)
-mtot <- mlfs+msts
-
-plot(ht, mlfs/mtot, type='l', ylim=c(0,1))
-points(ht, (mlfs+msts)/mtot, type='l')
 
 
 

### 2014-03-21 analysis/stem_leaf_scaling.Rnw
Added comparison between tropical and temperate evergreen angiosperms

@@
 
+\section{By plant functional type}
+
 <<plotting, echo=FALSE>>=
-# natural scale
-plot(1, xlim=c(0,50), ylim=c(0,1), type='n',
-     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)")
-for(i in 1:length(l))points(l[[i]]$h, l[[i]]$wfrel, type='l')
-
-# log-log scale
-plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
-     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
-     axes=FALSE)
-addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
-for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l')
-
-# log-natural scale
-plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
-     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
-     axes=FALSE)
-addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
-axis(2)
-for(i in 1:length(l))points(log10(l[[i]]$h), l[[i]]$wfrel, type='l')
-
-# Coloured by pft
+# # natural scale
+# plot(1, xlim=c(0,50), ylim=c(0,1), type='n',
+#      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)")
+# for(i in 1:length(l))points(l[[i]]$h, l[[i]]$wfrel, type='l')
+# 
+# # log-log scale
+# plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
+#      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
+#      axes=FALSE)
+# addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
+# for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l')
+# 
+# # log-natural scale
+# plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
+#      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
+#      axes=FALSE)
+# addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
+# axis(2)
+# for(i in 1:length(l))points(log10(l[[i]]$h), l[[i]]$wfrel, type='l')
 
@@
 
+\section{Tropical vs. temperate evergreens}
+
+<<echo=FALSE>>=
+ii <- which(P$pft == "EA" & P$vegetation %in% c("TempF","TempRF","TropRF","TropSF"))
+l <- l[ii]
+P <- P[ii,]
+
+P$TempTrop <- as.factor(ifelse(P$vegetation %in% c("TempF","TempRF"), "Temp", "Trop"))
+palette(c("blue","red"))
+
+@
+
+<<plotEA, echo=FALSE>>=
+
+# log-log scale
+plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
+     xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
+     axes=FALSE)
+addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
+axis(2)
+for(i in 1:length(l)){
+  points(log10(l[[i]]$h), l[[i]]$wfrel, type='l',
+         col=P$TempTrop[i])
+}
+legend("bottomleft", c("Temperate evergreen angiosperm",
+                       "Tropical evergreen angiosperm"), fill=palette())
+
+plot(1, xlim=c(-3,2), ylim=c(-7,5), type='n',
+     xlab="Plant height (m)", ylab="Leaf mass (kg)",
+     axes=FALSE)
+magaxis(side=1:2, unlog='xy',box=TRUE)
+for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$mlf), type='l',
+                            col=P$TempTrop[i])
+legend("topleft", c("Temperate evergreen angiosperm",
+                       "Tropical evergreen angiosperm"), fill=palette(),inset=0.03)
+
+plot(1, xlim=c(-3,2), ylim=c(-7,5), type='n',
+     xlab="Plant height (m)", ylab="Above-ground woody mass (kg)",
+     axes=FALSE)
+magaxis(side=1:2, unlog='xy',box=TRUE)
+for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$mst), type='l',
+                            col=P$TempTrop[i])
+legend("topleft", c("Temperate evergreen angiosperm",
+                       "Tropical evergreen angiosperm"), fill=palette(),inset=0.03)
+
+@
+
 \end{document}

### 2014-03-25 analysis/stem_leaf_scaling.Rnw

@@
 library(magicaxis)
+library(scales)
 #library(Hmisc)
@@
 <<analysis, echo=FALSE>>=
+
+# sma by dataset_species
 sm_mlf <- sma(m.lf ~ h.t * dataset_species, log="xy", data=x)
 sm_mst <- sma(m.st ~ h.t * dataset_species, log="xy", data=x)
+
+# quadratic across all data
 
@@
 # log-log scale
-palette(c("blue","royalblue","red","darkorange"))
+# windows(7,6)
+# par(mar=c(5,5,2,2), cex.lab=1.2)
+palette(alpha(c("blue","royalblue","red","darkorange"),0.4))
 plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
@@
                        "Evergreen angiosperm","Evergreen gymnosperm"),
-       fill=palette())
+       fill=alpha(palette(),0.9))
+
+# Some flourishes for pdf in a presentation
+# l_eve <- rbind.fill(l[P$pft %in% c("EA","EG")])
+# l_dec <- rbind.fill(l[P$pft %in% c("DA","DG")])
+# 
+# library(plantecophys)
+# l_eve$logwfrel <- log10(l_eve$wfrel)
+# l_eve$logh <- log10(l_eve$h)
+# l_dec$logwfrel <- log10(l_dec$wfrel)
+# l_dec$logh <- log10(l_dec$h)
+# gamplot("logh","logwfrel",dfr=l_eve,linecolor="red",add=TRUE)
+# gamplot("logh","logwfrel",dfr=l_dec,linecolor="blue",add=TRUE)
+# 
 # dev.copy2pdf(file="analysis/output/figures/LMF_height_loglog.pdf")

### 2014-04-11 analysis/stem_leaf_scaling.Rnw
split stem-leaf scaling by glasshouse/common garden plants and field-grown plants.

@@
 # Add some metadata
-dfr <- dat[,c("dataset","species","pft","vegetation")]
+dfr <- dat[,c("dataset","species","pft","vegetation","growingCondition")]
 dfr$dataset_species <- with(dfr, paste0(dataset,"_",species))
@@
 P$pft <- as.factor(P$pft)
+P$vegetation[is.na(P$vegetation)] <- "GH"
 P$vegetation <- as.factor(P$vegetation)
@@
 <<plotting, echo=FALSE>>=
-# # natural scale
-# plot(1, xlim=c(0,50), ylim=c(0,1), type='n',
-#      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)")
-# for(i in 1:length(l))points(l[[i]]$h, l[[i]]$wfrel, type='l')
-# 
-# # log-log scale
-# plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
-#      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
-#      axes=FALSE)
-# addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
-# for(i in 1:length(l))points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l')
-# 
-# # log-natural scale
-# plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
-#      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
-#      axes=FALSE)
-# addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
-# axis(2)
-# for(i in 1:length(l))points(log10(l[[i]]$h), l[[i]]$wfrel, type='l')
 
-# log-log scale
-# windows(7,6)
-# par(mar=c(5,5,2,2), cex.lab=1.2)
+# nonfield <- union(which(P$vegetation == "GH"), which(P$dataset == "Poorter1999"))
+nonfield <- which(P$growingCondition %in% c("CG","GH"))
+field <- setdiff(1:nrow(P), nonfield)
+
+# Non-glasshouse
 palette(alpha(c("blue","royalblue","red","darkorange"),0.4))
@@
      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
-     axes=FALSE)
+     axes=FALSE, main="Field grown")
 addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
-for(i in 1:length(l)){
+for(i in field){
   points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l',
@@
        fill=alpha(palette(),0.9))
-
 # Some flourishes for pdf in a presentation
@@
 
-# log-log scale
-plot(1, xlim=c(-3,2), ylim=c(0,1), type='n',
+palette(alpha(c("blue","royalblue","red","darkorange"),0.4))
+plot(1, xlim=c(-3,2), ylim=c(-3,0), type='n',
      xlab="Plant height (m)", ylab="Leaf mass / aboveground mass (-)",
-     axes=FALSE)
-addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015,addY=F)
-axis(2)
-for(i in 1:length(l)){
-  points(log10(l[[i]]$h), l[[i]]$wfrel, type='l',
+     axes=FALSE, main="Glasshouse and Common garden")
+addlogaxes(10^-3, 10^2, 10^-3, 10^0, tck1=-0.03, tck2=-0.015)
+for(i in nonfield){
+  points(log10(l[[i]]$h), log10(l[[i]]$wfrel), type='l',
          col=P$pft[i])
@@
                        "Evergreen angiosperm","Evergreen gymnosperm"),
-       fill=palette())
-# dev.copy2pdf(file="analysis/output/figures/LMF_height_lognat.pdf")
+       fill=alpha(palette(),0.9))
 
@@
        fill=palette(), inset=0.04)
-# dev.copy2pdf(file="analysis/output/figures/MF_height_loglog.pdf")
 
@@
        fill=palette(), inset=0.04)
-# dev.copy2pdf(file="analysis/output/figures/MS_height_loglog.pdf")
 @

### 2014-05-01 analysis/metadata_tables.Rnw
Added Rnw file with some latex summary tables, and a map (analysis/metadata_tables.Rnw)

\documentclass{article}

\begin{document}

\title{Metadata, tables, maps}
\maketitle

<<load, echo=FALSE, message=FALSE>>=
dat <- readRDS("../cache/allomdata_post.rds")

# library(smatr)
# library(magicaxis)
library(scales)
library(xtable)
source("../R/remko_functions.R")
source("../R/plotting.R")
@

\section{Tables}

<<results='asis', echo=FALSE>>=

# Family
topfam <- names(sort(table(dat$family),T)[1:15])
datx <- droplevels(subset(dat, family %in% topfam))
datx$family <- with(datx, reorder(as.factor(family),family,function(x)1/length(x)))

famtab <- xtable(xtabs(~ family, data=datx), digits=0,
                 caption="Most represented families (number of observations).",
                 label="table:familytable")
print(famtab, caption.placement="top",table.placement="h")

# Vegetation and plant functional type
vegpfttab <- xtable(xtabs(~ vegetation + pft, data=subset(dat, !is.na(pft) & pft != "")), digits=0,
                    caption="Number of observations by family and plant functional type.",
                    label="table:vegpfttab")
print(vegpfttab, caption.placement="top",table.placement="h")

# growinCondition and plant functional type
gcpfttab <- xtable(xtabs(~ growingCondition + pft, data=subset(dat, !is.na(pft) & pft != "")), digits=0,
                    caption="Number of observations by growing condition and plant functional type.",
                    label="table:gcpfttab")
print(gcpfttab, caption.placement="top",table.placement="h")

# Table of variables and nr of observations.
nr <- function(x)sum(!is.na(dat[,x]))
cfg <- read.csv("../config/variabledefinitions.csv")

vars <- c("m.st","m.so","m.lf","a.lf","m.rt","h.t","c.d","d.bh", "ma.ilf","r.st")
nrdf <- data.frame(Variable=vars, nobs=unname(sapply(vars,nr)))

nrdf <- merge(nrdf, cfg[,c("Variable","label")])
nrdf <- nrdf[order(nrdf$nobs),]
nrdftab <- xtable(nrdf, caption = "Number of observations by variable.", 
                  label = "nrdftable")
print(nrdftab,caption.placement = "top", include.rownames= FALSE, table.placement = "h")
@

\section{Maps}

<<echo=FALSE,  out.width='1.3\\linewidth', out.height='1.3\\linewidth'>>=
# Colored by vegetation type
cols <- alpha(c("blue","brown","forestgreen","darkorange","dodgerblue2",
                "navajowhite4","lightgoldenrod3","lavender"),0.67)
vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
vegs_labels <- c("Boreal forest","Savannah","Temperate forest","Temperate rainforest","Tropical rainforest",
                 "Tropical seasonal forest","Woodland","Glasshouse")

drawWorldPlot(dat[dat$vegetation == vegs[1],], horlines=FALSE, sizebyn=TRUE, pchcol=cols[1])
for(i in 2:length(vegs)){
  drawWorldPlot(dat[dat$vegetation == vegs[i],], horlines=FALSE, sizebyn=TRUE, pchcol=cols[i], add=TRUE)
}
drawWorldPlot(dat[dat$growingCondition == "GH",], horlines=FALSE, sizebyn=TRUE, 
              pchcol=cols[8], add=TRUE)
par(xpd=NA)
legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
legend(-50, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')

@

\end{document}

### 2014-05-01 analysis/metadata_tables.Rnw
Fixed tables

@@
 
-\section{Tables}
-
 <<results='asis', echo=FALSE>>=
 
-# Family
+# Top 15 families
 topfam <- names(sort(table(dat$family),T)[1:15])
@@
 
+# Top 15 species
+topspec <- names(sort(table(dat$species_Fixed),T)[1:15])
+datx <- droplevels(subset(dat, species_Fixed %in% topspec))
+datx$Species <- with(datx, reorder(as.factor(species_Fixed),species_Fixed,
+                                   function(x)1/length(x)))
+
+spectab <- xtable(xtabs(~ Species, data=datx), digits=0,
+                 caption="Most represented species (number of observations).",
+                 label="table:spectable")
+print(spectab, caption.placement="top",table.placement="h")
+
+dat_pft <- within(dat, {
+  pft[is.na(pft) | pft == ""] <- "None"
+  growingCondition[is.na(growingCondition) | growingCondition == ""] <- "None"
+  vegetation[is.na(vegetation) | vegetation == ""] <- "None"
+
+  pft <- relevel(as.factor(pft), "None")
+  growingCondition <- relevel(as.factor(growingCondition), "None")
+  vegetation <- relevel(as.factor(vegetation), "None")
+})
+
 # Vegetation and plant functional type
-vegpfttab <- xtable(xtabs(~ vegetation + pft, data=subset(dat, !is.na(pft) & pft != "")), digits=0,
+vegpfttab <- xtable(addmargins(xtabs(~ vegetation + pft, data=dat_pft)), 
+                    digits=0,
                     caption="Number of observations by family and plant functional type.",
@@
 # growinCondition and plant functional type
-gcpfttab <- xtable(xtabs(~ growingCondition + pft, data=subset(dat, !is.na(pft) & pft != "")), digits=0,
+gcpfttab <- xtable(addmargins(xtabs(~ growingCondition + pft, data=dat_pft)), 
+                   digits=0,
                     caption="Number of observations by growing condition and plant functional type.",
@@
 nrdf <- merge(nrdf, cfg[,c("Variable","label")])
-nrdf <- nrdf[order(nrdf$nobs),]
-nrdftab <- xtable(nrdf, caption = "Number of observations by variable.", 
+nrdf <- nrdf[order(nrdf$nobs,decreasing=TRUE),]
+nrdftab <- xtable(nrdf, caption = "Number of observations by selected variables.", 
                   label = "nrdftable")
@@
 
-\section{Maps}
-
-<<echo=FALSE,  out.width='1.3\\linewidth', out.height='1.3\\linewidth'>>=
+%,  out.width='1.3\\linewidth', out.height='1.3\\linewidth'>>=
+<<echo=FALSE, fig.cap="Global distribution of datasets by vegetation type and sample size.",out.width='1.1\\linewidth'>>=  
 # Colored by vegetation type

### 2014-05-07 ms/ms.Rmd
Outline for paper. #82

# *Ecological Archives* Exxxxx
# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants.*Ecology* xxxx

```{r, results="hide", echo=FALSE, message=FALSE}
library(knitr, quietly=TRUE)
knitr:::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="", cache=TRUE)

setwd("..")
source('R/import.R')
source("R/MS-fun.R")
dat <- loadData(reprocess=FALSE)

vars  <-  mashrDetail("var.def")$Variable[mashrDetail("var.def")$Type=="numeric"]
vars <- vars[-c(match(c("map","mat","lai","age"),vars))]

library(plyr)

data  <-  dat$data
```

## INTRODUCTION

motivation text.

## METHODS

## METADATA CLASS I. DATA SET DESCRIPTORS

### A. DATA SET IDENTITY
BAAD: a Biomass And Allometry Database for woody plants

### B. DATA SET IDENTIFICATION CODE:

### C. DATA SET DESCRIPTORS:
**1. Originators**: The study was initiated by Daniel Falster and Remko Duursma. Diego Barneche, Rich FitzJohn and Angelica Vårhammar assisted with literature searching, data compilation, and coding. All other authors are involved because they contributed data to the project.

**2. Abstract**: The allocation of primary production between different plant tissues (leaf, sapwood, bark, roots, reproductive material) determines the nature of plant growth, the outcome of competitive interactions, and the flow of energy in the terrestrial biosphere. Allocation patterns can differ widely among coexisting species in any single forest. Large-scale shifts with environment are also evident, e.g. between tropical versus temperate stands. The effect of allocation differences on the performance of individual plants and whole forest stands can now be studied using new vegetation models that trace growth across the entire life cycle from seed to adult tree. The success of such models, however, requires suitable datasets to validate model assumptions, parameterise equations, and test predictions about plant growth and development. Here we describe an individual-level data set capturing ontogenetic profiles of biomass allocation in woody plant species from biomes across the globe. The data were compiled from published and unpublished sources, focussing on studies where: i) measurements were for individual plants rather than whole stands; ii) individuals spanning a range of sizes were measured; and iii) size was estimated directly, i.e. not indirectly via allometric equations. We included both wild and artificially grown plants. The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues. We also report other properties of individuals (age, maximum height, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (status, light, experimental treatment, site leaf area index, vegetation type, mean annual rainfall) where available. Overall, the dataset contains `r sum(!is.na(dat$data[,vars]))` measurements collected in `r length(unique(dat$data$dataset))` different studies, from `r nrow(dat$data)` individuals across `r length(unique(dat$data$species))` species. It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the worlds vegetation.

### D. Key words

allometry; trait; leaf area; height; size; architecture; vegetation
model; carbon; growth; wood

## METADATA CLASS II. RESEARCH ORIGIN DESCRIPTORS

### B. SPECIFIC SUBPROJECT DESCRIPTION

```{r,results="asis",echo=FALSE}
setwd("..")
details2B  <-  daply(data, .(dataset), function(x)class2Bdetails(x))
writeLines(details2B,sep="\n\n")
```

## CLASS III. DATA SET STATUS AND ACCESSIBILITY

### A. Status

**Latest update**: xx for all data files.

**Metadata status**: Metadata are complete to last update and are stored
with data.

**Data verification**:

### B. Accessibility

**1. Storage location and medium**: The data published here is the first release from the Biomass and allometry database stored at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/). The version of the data made available for this publication is available here XXXXX. For scientific analyses, we suggest you use this version.

Development versions of the dataset will continue to be available at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).

If you notice any problem with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).

**Contact persons**: Queries about individual specific data points can be directed to the contributing author for that study. For queries about the entire dataset, please contact either

-  Daniel Falster, email: Daniel.Falster@mq.edu.au, tel:
  +61-2-98509258, Biological Sciences, Macquarie University NSW 2109,
  Australia
-  Remko Duursma, email: remkoduursma@gmail.com, tel: +61-2-4570 1806,
  Centre for Plants and the Environment, University of Western Sydney,
  NSW 1797, Australia.

**Copyright restrictions**: The dataset is released under the [Creative Commons CC0 license](http://creativecommons.org/choose/zero/?lang=en). As such, it may be freely used and redistributed. When using the dataset, we ask that you cite the original article, recognising the hard work that went into collecting the data and the author's willingness to make it publicly available.

**Costs**: None. You are kindly requested to cite this paper when using these data if possible.

## CLASS IV. DATA STRUCTURAL DESCRIPTORS

### A. Data Set Files

Identity: allData.txt
Size:
Format and storage mode: ASCII text, comma delimited. Missing values
are denoted as 'NA'.
Header information:

### B. Variable definitions

## CLASS V. SUPPLEMENTAL DESCRIPTORS

## ACKNOWLEDGMENTS

D.S. Falster, A. Vårhammer and D. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to Duursma.

## LITERATURE CITED

### 2014-05-07 ms/ms.Rmd
Update manuscript building

@@
 library(knitr, quietly=TRUE)
-knitr:::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="", cache=TRUE)
+knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")
+```
 
-setwd("..")
-source('R/import.R')
-source("R/MS-fun.R")
-dat <- loadData(reprocess=FALSE)
+```{r, results="hide", echo=FALSE, message=FALSE}
+source("MS-fun.R")
 
-vars  <-  mashrDetail("var.def")$Variable[mashrDetail("var.def")$Type=="numeric"]
-vars <- vars[-c(match(c("map","mat","lai","age"),vars))]
+baad <- readRDS("../output/baad.rds")
+data <- baad$data
 
-library(plyr)
-
-data  <-  dat$data
+var.def <- read.csv("../config/variableDefinitions.csv",
+                    stringsAsFactors=FALSE)
+vars <- var.def$Variable[var.def$Type == "numeric"]
+vars <- setdiff(vars, c("map","mat","lai","age"))
 ```
@@
 We also report other properties of individuals (age, maximum height, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (status, light, experimental treatment, site leaf area index, vegetation type, mean annual rainfall) where available.
-Overall, the dataset contains `r sum(!is.na(dat$data[,vars]))` measurements collected in `r length(unique(dat$data$dataset))` different studies, from `r nrow(dat$data)` individuals across `r length(unique(dat$data$species))` species.
+Overall, the dataset contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$dataset))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
 It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the worlds vegetation.
@@
 
-```{r,results="asis",echo=FALSE}
-setwd("..")
+```{r,results="asis",echo=FALSE, message=FALSE}
+library(plyr)
 details2B  <-  daply(data, .(dataset), function(x)class2Bdetails(x))

### 2014-05-07 ms/MS-fun.R

@@ -1,8 +1,7 @@
-source('R/import.R')
-library(plyr)
-
-d     <- loadData(reprocess=TRUE)
-data  <-  d$data
+## Mimic dataMashR function here, but relative to *this* directory.
+data.path <- function(...) {
+  file.path("../data", ...)
+}
 
 pasteAndCollapse  <-  function(values, binder) {
 	paste0(unique(values), collapse=binder)
@@ -45,9 +44,3 @@ class2Bdetails  <-  function(data) {
 		   '\ta. Field Laboratory ', '\n\t\t',              getMetadataMethods(unique(data$dataset)),              '\n',
 		   '\t4. Project personnel: ',                      getPersonell(unique(data$dataset)),                    '\n')	
 }
-
-details2B  <-  daply(data, .(dataset), function(x)class2Bdetails(x))
-
-#test to check it out
-test  <-  pasteAndCollapse(details2B, '\n\n')
-write(test, '~/Desktop/test.md')

### 2014-05-07 ms/formatBib.R

@@ -1,4 +1,3 @@
-
 library(gdata, quietly=TRUE)
 library(bibtex, quietly=TRUE)
 
@@ -87,6 +86,3 @@ formatBib <- function(bibfile,
   
 return(f)
 }
-
-
-

### 2014-05-08 ms/ms.Rmd
added whisker package to ms.Rmd file

@@
 library(plyr)
+library(whisker)
 details2B  <-  daply(data, .(dataset), function(x)class2Bdetails(x))

### 2014-05-08 ms/ms.Rmd
Modify paper: complete abstract and intro. #82

@@
 
-motivation text.
+Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science.
+While many thousands of trees have been harvested in the name of science, most of this data has only been made available in the form of summary tables or plots included in reports and publications, the raw data residing piecemeal on the hard drives of individual scientists spread around the world.
+Several researchers have gathered together the fitted equations within the data; yet few have attempted to liberate the raw data underpinning these relationships.
+Gathered together such data would have enormous value and represent an important resource for the community.
+
+We (D Falster, R Duursma) set out to create such a resource, contacting the authors of publications to see if they would make their raw data files available, with the help of D Barneche, A Vårhammer and R FitzJohn.
+The response was overwhelming: nearly everyone we contacted was interested to contribute.
+As a result, we are happy to release BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$dataset))` different studies.
+
+Combining data from many different studies, each using it's own set of terms and units is no trivial task.
+This compilation would not have been possible without several new technologies.
+In particular,
+
+1. Our entire workflow is scripted, using the program `R`. The database can be rebuilt in an instant from the raw datafiles, using a single command.
+1.
+We wrote a new R package [dataMashR](https://github.com/dfalster/dataMashR) for merging data from different studies in a transparent and reproducible fashion.
+While it was written with BAAD in mind, the package could easily be deployed for new compilations.
+3.
+We used version control and the code sharing website [github](www.github.com) to track the evolution of the database over time.
+Datasets inevitably contain errors and modifications are made.
+All of these changes have been tracked and can thus be traced to particular sources and modification dates.
+3.
+We used the package `knitr` within the program `R` to produce reports comparing the data provided from a single study to the rest of the database.
+These reports, which were also sent to data contributors, made it easy to identify errors within the dataset.
+
+Until recently, scientific publications in biology, by their nature have mostly been concerned with `small data`.
+Datasets arrive in small parcels, corresponding to the work of an individual study.
+It is exciting to be able to put these small packages together, to create something bigger.
+We hope other people will use this dataset and build on it.
+See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
 
@@
 Diego Barneche, Rich FitzJohn and Angelica Vårhammar assisted with literature searching, data compilation, and coding.
-All other authors are involved because they contributed data to the project.
+All other authors  contributed data to the project.
 
 **2.
-Abstract**: The allocation of primary production between different plant tissues (leaf, sapwood, bark, roots, reproductive material) determines the nature of plant growth, the outcome of competitive interactions, and the flow of energy in the terrestrial biosphere.
-Allocation patterns can differ widely among coexisting species in any single forest.
-Large-scale shifts with environment are also evident, e.g.
-between tropical versus temperate stands.
-The effect of allocation differences on the performance of individual plants and whole forest stands can now be studied using new vegetation models that trace growth across the entire life cycle from seed to adult tree.
-The success of such models, however, requires suitable datasets to validate model assumptions, parameterise equations, and test predictions about plant growth and development.
-Here we describe an individual-level data set capturing ontogenetic profiles of biomass allocation in woody plant species from biomes across the globe.
-The data were compiled from published and unpublished sources, focussing on studies where: i) measurements were for individual plants rather than whole stands; ii) individuals spanning a range of sizes were measured; and iii) size was estimated directly, i.e.
-not indirectly via allometric equations.
+Abstract**: Understanding how different plants are plant constructed -- i.e.
+how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modelling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere.
+Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
+While a variety of models dealing with allocation exist, a synethtic understanding about plant allocation is lacking, in part due to the lack of suitable datasets for validating and parametrising models.
+To that end, we present the Biomass And Allometry Database (BAAD) for woody plants.
+BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$dataset))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
+Most of this data is associated with an existing publication, however, mostly raw data files were not released at time of publication.
+Thus BAAD contains individual level data from different studies, transformed into standard units and variable names.
+The transformations were achieved using a new R package dataMashR.
+Other features which distinguish BAAD are that we focussed on studies where: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
+not indirectly via allometric equations (with one exception for trees > 60m).
 We included both wild and artificially grown plants.
 The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues.
-We also report other properties of individuals (age, maximum height, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (status, light, experimental treatment, site leaf area index, vegetation type, mean annual rainfall) where available.
-Overall, the dataset contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$dataset))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
+We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, status, light, experimental treatment, vegetation type) where available.
 It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the worlds vegetation.
@@
 
-**Contact persons**: Queries about individual specific data points can be directed to the contributing author for that study. For queries about the entire dataset, please contact either
-
--  Daniel Falster, email: Daniel.Falster@mq.edu.au, tel:
-  +61-2-98509258, Biological Sciences, Macquarie University NSW 2109,
-  Australia
--  Remko Duursma, email: remkoduursma@gmail.com, tel: +61-2-4570 1806,
-  Centre for Plants and the Environment, University of Western Sydney,
-  NSW 1797, Australia.
+**Contact persons**: Queries about individual specific data points can be directed to the contributing author for that study.
+For queries about the entire dataset, please contact either Daniel Falster or Remko Duursma.
+Contact details for all authors are available in the attached table.
 
@@
 As such, it may be freely used and redistributed.
-When using the dataset, we ask that you cite the original article, recognising the hard work that went into collecting the data and the author's willingness to make it publicly available.
+When using the dataset, we kindly request that you cite this article, recognising the hard work that went into collecting the data and the author's willingness to make it publicly available.
 
-**Costs**: None. You are kindly requested to cite this paper when using these data if possible.
+**Costs**: None.
 
@@
 
-D.S. Falster, A. Vårhammer and D. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to Duursma.
+D.S. Falster, A. Vårhammer and D. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to Duursma. R FitzJohn was employed on a XXXXXX.
 

### 2014-05-08 ms/ms.Rmd
added list of authors to rmd file
Update ms.Rmd

@@
 
+```{r,results="asis",echo=FALSE, message=FALSE}
+library(plyr)
+lisOfAuthors(vardef=var.def)
+```
+
 ## INTRODUCTION
@@
 Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science.
-While many thousands of trees have been harvested in the name of science, most of this data has only been made available in the form of summary tables or plots included in reports and publications, the raw data residing piecemeal on the hard drives of individual scientists spread around the world.
-Several researchers have gathered together the fitted equations within the data; yet few have attempted to liberate the raw data underpinning these relationships.
-Gathered together such data would have enormous value and represent an important resource for the community.
+The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants.
+While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications, or as totals for a forest plot.
+Until now, much of the raw data resides piecemeal on the hard drives of individual scientists spread around the world.
 
-We (D Falster, R Duursma) set out to create such a resource, contacting the authors of publications to see if they would make their raw data files available, with the help of D Barneche, A Vårhammer and R FitzJohn.
-The response was overwhelming: nearly everyone we contacted was interested to contribute.
-As a result, we are happy to release BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$dataset))` different studies.
+Several studies have gathered together the fitted (allometric) equations for separate datasets, yet few have attempted to organize and share the raw individual plant data underpinning these equations.
+Gathered together, such data would be invaluable and represent an important resource for the community.
 
-Combining data from many different studies, each using it's own set of terms and units is no trivial task.
+We (D.
+Falster and R.
+Duursma, with the help of D Barneche, A Vårhammer and R FitzJohn) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
+The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.
+As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$dataset))` different published and unpublished studies.
+
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task.
 This compilation would not have been possible without several new technologies.
@@
 
-1. Our entire workflow is scripted, using the program `R`. The database can be rebuilt in an instant from the raw datafiles, using a single command.
 1.
-We wrote a new R package [dataMashR](https://github.com/dfalster/dataMashR) for merging data from different studies in a transparent and reproducible fashion.
-While it was written with BAAD in mind, the package could easily be deployed for new compilations.
+Our entire workflow is scripted, using the `R` computing environment.
+The database can be rebuilt in an instant from the raw datafiles, using a single command.
+When a single data contribution has been modified, the final master dataset is thus easily recreated.
+2.
+We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) for merging data from different studies in a transparent and reproducible fashion.
+While it was written with BAAD in mind, the package is designed to be easily  deployed for other compilations.
 3.
-We used version control and the code sharing website [github](www.github.com) to track the evolution of the database over time.
-Datasets inevitably contain errors and modifications are made.
-All of these changes have been tracked and can thus be traced to particular sources and modification dates.
-3.
-We used the package `knitr` within the program `R` to produce reports comparing the data provided from a single study to the rest of the database.
-These reports, which were also sent to data contributors, made it easy to identify errors within the dataset.
+We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data.
+Datasets as well as processing methods inevitably contain errors.
+With version control, we are able to trace these errors to the source.
+4.
+We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database.
+These reports were sent to data contributors for verification, and allowed easy identification of errors in the raw data or processing.
 
-Until recently, scientific publications in biology, by their nature have mostly been concerned with `small data`.
-Datasets arrive in small parcels, corresponding to the work of an individual study.
-It is exciting to be able to put these small packages together, to create something bigger.
+Until recently, scientific publications in biology have mostly been concerned with `small data`.
+Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
+It is exciting to be able to put these parcels together, to create something larger.
 We hope other people will use this dataset and build on it.
@@
 Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
-While a variety of models dealing with allocation exist, a synethtic understanding about plant allocation is lacking, in part due to the lack of suitable datasets for validating and parametrising models.
+While a variety of models dealing with allocation exist, a syntethic understanding about plant allocation is lacking.
+This is partly due to the lack of suitable datasets for validating and parametrising models.
 To that end, we present the Biomass And Allometry Database (BAAD) for woody plants.
@@
 
-D.S. Falster, A. Vårhammer and D. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to Duursma. R FitzJohn was employed on a XXXXXX.
+D.S. Falster, A. Vårhammer and D. R. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to Duursma. R FitzJohn was employed on a XXXXXX.
 

### 2014-05-09 ms/ms.Rmd
added first trial of citations using knitcitations package

@@
 
+```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
+library(bibtex)
+library(knitcitations)
+```
+
 ## INTRODUCTION
+
+```{r, eval=FALSE, echo=FALSE}
+#these commands below from package knitcitations work if you knit the file from within RStudio (being in the right ms directory), but they do not work when you generate the file using make.
+#not working
+#citep(c(Falster2011="10.1111/j.1365-2745.2010.01735.x"))
+#citep(citation("base"))
+#r citet("https://github.com/dfalster/dataMashR")
+```
 
@@
 1.
-Our entire workflow is scripted, using the `R` computing environment.
+Our entire workflow is scripted, using the R computing environment.
 The database can be rebuilt in an instant from the raw datafiles, using a single command.
@@
 See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
+
+## REFERENCES
+```{r, echo=FALSE, results='asis'}
+bibliography(ordering = c("authors", "year", "title", "journal", "volume", "number", "pages", "url"))
+```
 

### 2014-05-09 ms/ms.Rmd
added references in ms using knitcitations

--- 2014-05-08 ms/ms.Rmd
@@
 
+```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
+library(bibtex)
+library(knitcitations)
+```
+
 ## INTRODUCTION
 
-Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science.
+Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(c(Falster2011="10.1111/j.1365-2745.2010.01735.x"))`.
 The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants.
@@
 1.
-Our entire workflow is scripted, using the `R` computing environment.
+Our entire workflow is scripted, using the R computing environment `r citep(citation("base"))`.
 The database can be rebuilt in an instant from the raw datafiles, using a single command.
@@
 2.
-We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) for merging data from different studies in a transparent and reproducible fashion.
+We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) `r citet("https://github.com/dfalster/dataMashR")` for merging data from different studies in a transparent and reproducible fashion.
 While it was written with BAAD in mind, the package is designed to be easily  deployed for other compilations.

### 2014-05-13 reports/report.Rmd
Add reports. Closes #93

--- 2013-06-07 report/report-1.Rmd
@@
 ```{r, echo=FALSE}
-
-locLevelInfo  <-  function(data){
-
-  vars <- c("location","longitude","latitude","vegetation")
-
-  loc <- data[!duplicated(data[,vars]),vars]
-
-  loc[is.na(loc) | loc=="" | loc=="NA"]  <-  "missing"  
-  return(loc)
-}
-
-standLevelInfo  <-  function(data){
-
-  vars <- c("location","grouping","growingCondition","status")
-  sta <- data[!duplicated(data[,vars]),vars]
-
-  if(all(is.na(sta$grouping)))
-    sta  <-  sta[,-which(names(sta) == "grouping")]
-
-  sta[is.na(sta) | sta=="" | sta=="NA"]  <-  "missing"
-
-  return(sta)
-}
-
-spLevelInfo  <-  function(data){
-  spec         <-  data.frame(species=as.character(unique(data$species)), stringsAsFactors=FALSE)
-  for(z in c("family", "pft")){
-    spec[[z]]  <-  as.character(data[[z]][match(spec$species,data$species)])
-    i          <-  spec[[z]]=="" | is.na(spec[[z]]) 
-    if(any(i)){
-      spec[[z]][i]  <-  "missing"
-    }
-  }
-  j            <-  spec$species=="" | is.na(spec$species) 
-  if(any(j)){
-    spec$species[j]  <-  "missing"
-  }
-  spec
-}
-
-printMeta  <-  function(data){
-  dataset   <-  as.character(unique(data$dataset))
-  openMeta  <-  read.csv(paste0("../",dir.rawData,"/",dataset,"/studyMetadata.csv"), h=TRUE, stringsAsFactors=FALSE)
-  openMeta
-}
-
+source('report-fun.R')
 ```
@@
 Information about the primary reference for the data you provided.
-```{r echo=FALSE, comment=''}
-  bibfile<- paste0("../", as.character(.dat$ref$filename))
-```
-**Reference:** `r formatBib(bibfile)[[1]]`
 
-**DOI:** `r  bibGetElement(bibfile, element = "doi")`
+**Reference:** `r .dat$references$citation`
 
-**Abstract:** `r  bibGetElement(bibfile, element = "abstract")`
+**DOI:** `r  bibGetElement(.dat$bib, element = "doi")`
+
+**Abstract:** `r  bibGetElement(.dat$bib, element = "abstract")`
 
@@
 counts  <-  counts[counts>0]  #only include no zeros
-counts  <-  counts[names(counts)%in%var.def$Variable[var.def$Group=="tree"]]
+counts  <-  counts[names(counts)%in%.dat$dictionary$Variable[.dat$dictionary$Group=="tree"]]
 cbind(counts)
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-standLevelInfo(.dat$data)
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
+kable(standLevelInfo(.dat$data))
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-spLevelInfo(.dat$data)
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment='',results="asis"}
+kable(spLevelInfo(.dat$data))
 ```
@@
 ### Methods
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-metaTable  <-  printMeta(.dat$data)
-for(j in 1:nrow(metaTable)){
-  cat(paste0(metaTable[j,1], ": ", metaTable[j,2], "\n\n"))
-}
-
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
+metaTable  <-  getMeta(.study)
+writeLines(paste0(metaTable[,1], ": ", metaTable[,2], "\n"))
 ```
@@
 ```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
-makePlotPanel(alldata$data, .study, pdf=FALSE, quiet=TRUE)
+makePlotPanel(alldata, .dat)
 ```

### 2014-05-13 reports/report-fun.R

@@ -1,0 +1,427 @@
+library(knitr, quietly = TRUE)
+library(maps, quietly = TRUE)
+library(mapdata, quietly = TRUE)
+suppressPackageStartupMessages(library(gdata))
+
+printAllStudyReports <- function(data = readRDS("../output/baad.rds"), studynames = unique(data$data$dataset),
+    progressbar = TRUE, ...) {
+    message("Generating ", length(studynames), " markdown reports.")
+    if (progressbar) {
+        wp <- txtProgressBar(min = 0, max = length(studynames), initial = 0, width = 50,
+            style = 3)
+    }
+
+    output <- list()
+    for (i in seq_along(studynames)) {
+        output[[i]] <- printStudyReport(data, studynames[i], ...)
+        if (progressbar)
+            setTxtProgressBar(wp, i)
+    }
+
+    if (progressbar)
+        close(wp)
+}
+
+# creates html reports using knitr
+printStudyReport <- function(alldata, study, RmdFile = "report.Rmd", path = "output/report-by-study",
+    name = study, delete = TRUE, reprocess = FALSE) {
+
+    knitThis(RmdFile = RmdFile, path = path, name = name, delete = TRUE, reprocess = reprocess,
+        predefined = list(.study = study, alldata = alldata, .dat = extractStudy(alldata,
+            study)))
+}
+
+# creates html reports using knitr, copies output file to desired location and
+# renames as required
+knitThis <- function(RmdFile = "reportmd.Rmd", path = "output/report-per-study",
+    name = "study", delete = TRUE, reprocess = TRUE, ..., predefined = list(...)) {
+
+    outputfile <- paste0(path, "/", name, ".html")
+
+    if (reprocess || !file.exists(outputfile)) {
+
+        # create new environment with predfined variables
+        e <- new.env()
+        if (length(predefined) > 0) {
+            # avoid issues when variables unnamed
+            if (is.null(names(predefined)) || any(names(predefined) == ""))
+                stop("All extra variables must be named")
+
+            for (v in names(predefined)) assign(v, predefined[[v]], e)
+        }
+
+        # knit
+        suppressMessages(knit2html(RmdFile, quiet = TRUE, envir = e))
+
+        # copy html file to output dir
+        if (!file.exists(path))
[…]

### 2014-05-14 ms/ms.Rmd
imporve manuscript: add variables, reformat studies, add refs. Closes #80

@@
 knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")
+
+library(plyr)
+library(bibtex)
+library(knitcitations)
+source("MS-fun.R")
 ```
@@
 ```{r, results="hide", echo=FALSE, message=FALSE}
-source("MS-fun.R")
 
@@
                     stringsAsFactors=FALSE)
-vars <- var.def$Variable[var.def$Type == "numeric"]
-vars <- setdiff(vars, c("map","mat","lai","age"))
+vars <- setdiff(var.def$Variable[var.def$Type == "numeric"], c("map","mat","lai","age"))
 ```
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-library(plyr)
-lisOfAuthors(vardef=var.def)
-```
-
-```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
-library(bibtex)
-library(knitcitations)
+writeLines(paste("**List of Authors:**", lisOfAuthors(vardef=var.def)))
 ```
@@
 
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task.
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c(Michener1997="10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2", Kandel2011="10.1145/1978942.1979444"))`.
 This compilation would not have been possible without several new technologies.
@@
 4.
-We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database.
+We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`.
 These reports were sent to data contributors for verification, and allowed easy identification of errors in the raw data or processing.
 
-Until recently, scientific publications in biology have mostly been concerned with `small data`.
+Until recently, scientific publications in biology have mostly been concerned with "small data" `r citep(c(VanNoorden2013="10.1038/nj7461-243a"))`.
 Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
@@
 See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
-
-## REFERENCES
-```{r, echo=FALSE, results='asis'}
-bibliography(ordering = c("authors", "year", "title", "journal", "volume", "number", "pages", "url"))
-```
 
@@
 Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
-While a variety of models dealing with allocation exist, a syntethic understanding about plant allocation is lacking.
+While a variety of models dealing with allocation exist, a synthetic understanding about plant allocation is lacking.
 This is partly due to the lack of suitable datasets for validating and parametrising models.
@@
 
-Identity: allData.txt
+Identity:
 Size:
@@
 
-## CLASS V. SUPPLEMENTAL DESCRIPTORS
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
+x <-daply(var.def, 1, function(x)
+		paste0("* ", x$Variable,
+		"\n\t- Label: ", x$label,
+		"\n\t- Description: ", x$Description,
+		"\n\t- Type: ", x$Type,
+		"\n\t- Units: ", gsub(".","/", x$Units, fixed=TRUE)
+		))
 
+writeLines(x[var.def$Variable])
+```
 ## ACKNOWLEDGMENTS
@@
 
-## LITERATURE CITED
+## REFERENCES
+```{r, echo=FALSE, results='asis'}
+bibliography(ordering = c("authors", "year", "title", "journal", "volume", "number", "pages", "url"))
+```

### 2014-05-14 ms/MS-fun.R

@@ -3,12 +3,13 @@ data.path <- function(...) {
   file.path("../data", ...)
 }
 
-pasteAndCollapse  <-  function(values, binder) {
+#Paste and collapse
+pasteC  <-  function(values, binder="\n") {
 	paste0(unique(values), collapse=binder)
 }
 
 readStudyFile  <-  function(studyName, File) {
-	read.csv(data.path(studyName, File), header = TRUE, 
+	read.csv(data.path(studyName, File), header = TRUE,
         na.strings = c(NA,''), stringsAsFactors = FALSE, strip.white = TRUE)
 }
 
@@ -21,52 +22,79 @@ getMetadataMethods  <-  function(studyName) {
 	metadata  <-  readStudyFile(studyName, "studyMetadata.csv")
 	if(nrow(metadata) > 0) {
 		metadata  <-  metadata[!is.na(metadata$Description) & !(tolower(metadata$Topic) %in% c('sampling strategy', 'age', 'growth environment')),]
-		pasteAndCollapse(paste0(metadata$Topic, ': ', metadata$Description), '\n\t\t\t')
+		pasteC(paste0(metadata$Topic, ': ', metadata$Description), '\n\t- ')
 	}
 }
 
 getPersonell  <-  function(studyName) {
-	pasteAndCollapse(unique(readStudyFile(studyName, "studyContact.csv")$name), '; ')
+	pasteC(unique(readStudyFile(studyName, "studyContact.csv")$name), ', ')
 }
 
+getSiteHistory <-function(code){
+	llply(code, function(x)
+		switch(x,
+			FW = "field wild",
+			FE = "field experimental",
+			GH = "glasshouse",
+			PU = "plantation unmanaged",
+			PM = "plantation managed",
+			GC = "growth chamber",
+			CG = "common garden",
+			""
+			)
+		)
+}
+
+getVegType <-function(code){
+	llply(code, function(x)
+		switch(x,
+			Sav = "Savannah",
+			TropRF = "Tropical rainforest",
+			TempRF = "Temperate rainforest",
+			TropSF = "Tropical seasonal forest",
+			TempF = "Temperate forest",
+			BorF = "Boreal forest",
+			Wo = "Woodland",
+			Gr = "Grassland",
+			Sh = "Shrubland",
+			De = "Desert",
+			""
+			)
+		)
+}
 
 class2Bdetails  <-  function(data) {
 
 	template <-
     	c('## {{studyName}}',
-      	'\t1. Site Description',
-      	'\t\ta. Site(s) type(s)',
[…]

### 2014-05-14 ms/formatBib.R

@@ -1,88 +1,0 @@
-library(gdata, quietly=TRUE)
-library(bibtex, quietly=TRUE)
-
-bibGetElement <- function(bibfile, element = "doi", i=1){
-  bibs <- read.bib(bibfile)
-  unclass(bibs[[i]])[[1]][[element]]
-}
-
-formatBib <- function(bibfile,
-                      initialSep=".",
-                      fieldSep=". "
-                      ){
-
-  bibs <- read.bib(bibfile)
-  
-  
-  f <- lapply(bibs, function(b){
-    
-    b <- unclass(b)[[1]]
-    
-    # strip html tags in title (typically <i> etc.):
-    tit <- gsub("<.+?>","",b$title)
-    Encoding(tit) <- "UTF-8"  #allows for funny characters
-    
-    # author initials
-    a <- unclass(b$author)
-    
-    # family name
-    family <- unlist(sapply(a, "[[", "family"))
-    # extra unlist is necessary for NULL elements
-    
-    Encoding(family) <- "UTF-8"
-    
-    # Extract initials from given names
-    given <- lapply(a, "[[", "given")
-    initials <- lapply(given, function(x){
-      
-      # If initials only, they are in {}
-      if(any(grepl("}",x,fixed=TRUE))){
-        cl <- gsub("{","",x,fixed=TRUE)
-        cl <- gsub("}","",cl,fixed=TRUE)
-        cl <- gsub(".","",cl,fixed=TRUE)
-        
-        init <- strsplit(cl, "")[[1]]
-        return(init)
-      } else {  
-        substr(x,1,1)
-      }
-    })
-    
-    
-    initials <- unname(sapply(initials, function(x){
-      paste0(paste(x, collapse=initialSep),initialSep)
-      }))
-    
-    # author names
-    p <- paste(family, initials, collapse=", ")
-    
-    # field separator, unless it is identical to the initial separator
-    if(trim(fieldSep) != trim(initialSep))
-      p <- paste0(p, fieldSep)
-    else
-      p <- paste0(p, " ")
-    
-    # year
-    p <- paste0(p, b$year, fieldSep)
-    
-    # title
-    p <- paste0(p, tit, fieldSep)
-    
-    # journal name
-    jour <- b$journal
-    Encoding(jour) <- "UTF-8"
-    p <- paste0(p, jour, fieldSep)
-    
-    # not sure how often this happens ... double '-'
[…]

### 2014-05-26 ms/ms.Rmd
Add Ishihara as 3rd coauthor, details about contributions

@@
 Falster and R.
-Duursma, with the help of D Barneche, A Vårhammer and R FitzJohn) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
+Duursma, with the help of D.
+Barneche, R.
+FitzJohn and A.
+Vårhammer) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
 The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.
+Moreover, we were invited to incorporate another compilation led by M.
+Ishihara and focussing on Japanese literature.
 As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$dataset))` different published and unpublished studies.
@@
 1.
-Our entire workflow is scripted, using the R computing environment `r citep(citation("base"))`.
+Our entire work flow is scripted, using the R computing environment `r citep(citation("base"))`.
 The database can be rebuilt in an instant from the raw datafiles, using a single command.
@@
 Diego Barneche, Rich FitzJohn and Angelica Vårhammar assisted with literature searching, data compilation, and coding.
+A separate compilation focussing on natural forests in Japan, was collected by Masae Ishihara, with assistance from Tsutom Hiura, Hajime Utsugi and Hiroyuki Tanouchi.
 All other authors contributed data to the project.

### 2014-05-26 ms/MS-fun.R

@@ -151,17 +151,21 @@ getContributions  <-  function(data, ...) {
 	sort(tapply(correctedData$contribution, correctedData$contacts, sum), decreasing=TRUE)
 }
 
-orderByLastName  <-  function(authorNames) {
-	sapply(authorNames, function(x){a<-strsplit(x,' ')[[1]];a[length(a)]})
-}
-
 lisOfAuthors  <-  function(alphabetical=TRUE, ...) {
 	wanted         <-  contributionColumns(...)
 	contributions  <-  getContributions(data, wanted)
-	firstAuthors   <-  c('Daniel Falster', 'Remko Duursma', 'Diego R. Barneche', 'Rich Fitzjohn', 'Angelica Vårhammar')
+	firstAuthors   <-  c('Daniel Falster', 'Remko Duursma', 'Masae Ishihara', 'Diego R. Barneche', 'Rich FitzJohn', 'Angelica Vårhammar')
 	contributions  <-  contributions[!(names(contributions) %in% firstAuthors)]
+
+	getLastName  <-  function(authorNames) {
+		sapply(authorNames, function(x){
+									a <- strsplit(x,' ')[[1]]
+									a[length(a)]
+							})
+	}
+
 	if(alphabetical) {
-		pasteC(c(firstAuthors, names(contributions)[order(orderByLastName(names(contributions)))]), ', ')
+		pasteC(c(firstAuthors, names(contributions)[order(getLastName(names(contributions)))]), ', ')
 	} else {
 		pasteC(c(firstAuthors, names(contributions)), ', ')
 	}

### 2014-06-10 ms/ms.Rmd
Latitude, longitude formatting fixed (always one digit); citation to base removed : citep() crashed on citep(citation("base"))

@@
 1.
-Our entire work flow is scripted, using the R computing environment `r citep(citation("base"))`.
+Our entire work flow is scripted, using the R computing environment.
 The database can be rebuilt in an instant from the raw datafiles, using a single command.
@@
 ```
+
 ## ACKNOWLEDGMENTS
 
-D.S. Falster, A. Vårhammer and D. R. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to Duursma. R FitzJohn was employed on a XXXXXX.
+D.S. Falster, A. Vårhammer and D. R. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A. Duursma. R FitzJohn was employed on a XXXXXX.
 

### 2014-06-10 ms/ms.Rmd
variableDefinitions gains UnitsMD column, units formatted with <sup> markdown tag. Used in ms.Rmd to format units nicely.

@@
 		"\n\t- Type: ", x$Type,
-		"\n\t- Units: ", gsub(".","/", x$Units, fixed=TRUE)
+		"\n\t- Units: ", gsub(".","/", x$UnitsMD, fixed=TRUE)
 		))

### 2014-06-17 ms/ms.Rmd
Added table with nr of observations by variable in ms.Rmd. Closes #114

@@
 library(knitcitations)
+library(Hmisc)
+# library(pander)
+
 source("MS-fun.R")
@@
 
+### C. Number of observations by variable
+
+```{r echo=FALSE, message=FALSE}
+
+thesevars <- setdiff(var.def$Variable[var.def$Type=="numeric"], 
+                     c("map","mat","lai"))
+Labels <- capitalize(var.def$label[var.def$Variable %in% thesevars])
+
+Count <- function(x)length(data[[x]][!is.na(data[[x]])])
+N <- sapply(thesevars, Count)
+
+dfr <- data.frame(Variable=Labels, N=N)
+dfr <- dfr[dfr$N > 0,]
+rownames(dfr) <- NULL
+```
+
+```{r results='asis', echo=FALSE}
+kable(dfr, "html", align="l")
+```
+
 ## ACKNOWLEDGMENTS

### 2014-06-18 ms/MS-fun.R
simplify code for auhtor list

@@ -27,7 +27,7 @@ getMetadataMethods  <-  function(studyName) {
 }
 
 getPersonell  <-  function(studyName) {
-	pasteC(unique(readStudyFile(studyName, "studyContact.csv")$name), ', ')
+	readStudyFile(studyName, "studyContact.csv")[,"name", drop=TRUE]
 }
 
 getSiteHistory <-function(code){
@@ -92,82 +92,26 @@ class2Bdetails  <-  function(data) {
            	          metadataDesign=getMetadataDesign(unique(data$dataset)),
            	          collection='',
            	          lab=getMetadataMethods(unique(data$dataset)),
-           	          personell=getPersonell(unique(data$dataset)),
+           	          personell=pasteC(getPersonell(unique(data$dataset))),
            	          citation=baad$references$citation[baad$references$dataset==data$dataset[1]]
                  ))
 
 }
 
-numberOfPoints  <-  function(data, wanted) {
-	sum(apply(data[ , colnames(data) != 'dataset'], 2, function(x)length(x[!is.na(x)])))
-}
 
-expandData <-  function(data) {
-	if(data$duplicates > 1) {
-		newnames       <-  strsplit(data$contacts, ', ')[[1]]
-		data           <-  data[rep(1, data$duplicates), ]
-		data$contacts  <-  newnames
-		data
-	} else {
-		data
-	}
-}
-
-contributionColumns  <-  function(vardef) {
-	wanted  <-  unique(vardef$Variable[vardef$Type=='numeric'])
-	wanted[!(wanted %in% c('map', 'mat'))]
-}
-
-filterData4Contributions  <-  function(data, contributionCols) {
-	data[, c('dataset', contributionCols)]
-}
-
-listContacts  <-  function(data) {
-	daply(data,  .(dataset), function(x)getPersonell(unique(x$dataset)))
-}
-
-numberOfContacts  <-  function(data) {
-	sapply(data, function(x)length(strsplit(x, ',')[[1]]))
-}
-
-numberOfcontributions  <-  function(data) {
-	daply(data, .(dataset), numberOfPoints)
-}
-
-correctData  <-  function(data) {
-	ddply(data, .(study), expandData)
-}
-
-getContributions  <-  function(data, ...) {
-	sdata   <-  filterData4Contributions(data, ...)
-	conts   <-  listContacts(data)
[…]

### 2014-06-18 ms/ms.Rmd
simplify code for auhtor list
Remove ISSN records from bib files

@@
 # *Ecological Archives* Exxxxx
-# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants.*Ecology* xxxx
+# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants. *Ecology* xxxx
 
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-writeLines(paste("**List of Authors:**", lisOfAuthors(vardef=var.def)))
+writeLines(paste("**List of Authors:**", lisOfAuthors(baad$contact)))
 ```

### 2014-06-19 ms/ms.Rmd
In Ms change 'dataset' to 'studyName'

@@
 Ishihara and focussing on Japanese literature.
-As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$dataset))` different published and unpublished studies.
+As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies.
 
@@
 To that end, we present the Biomass And Allometry Database (BAAD) for woody plants.
-BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$dataset))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
+BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
 Most of this data is associated with an existing publication, however, mostly raw data files were not released at time of publication.
@@
 library(whisker)
-details2B  <-  daply(data, .(dataset), function(x)class2Bdetails(x))
+details2B  <-  daply(data, .(studyName), function(x)class2Bdetails(x))
 writeLines(details2B,sep="\n\n")

### 2014-06-19 ms/ms.Rmd
minor editing of ms.Rmd

@@
 library(Hmisc)
+library(whisker)
+
 # library(pander)
@@
 We (D.
-Falster and R.
+Falster and R.A.
 Duursma, with the help of D.
@@
 BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
-Most of this data is associated with an existing publication, however, mostly raw data files were not released at time of publication.
+Most of these data are associated with an existing publication, however, raw data files were rarely made public at time of publication.
 Thus BAAD contains individual level data from different studies, transformed into standard units and variable names.
 The transformations were achieved using a new R package dataMashR.
-Other features which distinguish BAAD are that we focussed on studies where: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
-not indirectly via allometric equations (with one exception for trees > 60m).
+Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
+not indirectly via allometric equations (with a few exceptions large trees which were sub-sampled and scaled up).
 We included both wild and artificially grown plants.
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-library(plyr)
-library(whisker)
 details2B  <-  daply(data, .(studyName), function(x)class2Bdetails(x))
@@
 Contact details for all authors are available in the attached table.
+For problems with the dataset, please submit an issue as described above.
 
@@
 As such, it may be freely used and redistributed.
-When using the dataset, we kindly request that you cite this article, recognising the hard work that went into collecting the data and the author's willingness to make it publicly available.
+When using the dataset, we kindly request that you cite this article, recognising the hard work that went into collecting the data and the authors' willingness to make it publicly available.
 

### 2014-06-19 ms/ms.Rmd
MS: imporve study descriptions: lat,lon, methods, citations

@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-details2B  <-  daply(data, .(studyName), function(x)class2Bdetails(x))
+library(plyr)
+library(whisker)
+details2B  <-  daply(data, .(studyName), function(x) studyDetails(x))
 writeLines(details2B,sep="\n\n")

### 2014-06-19 ms/ms.Rmd
Fix author initials. Closes #120

@@
 
-We (D.
+We (D.S.
 Falster and R.A.
-Duursma, with the help of D.
-Barneche, R.
+Duursma, with the help of D.R.
+Barneche, R.G.
 FitzJohn and A.
@@
 **1.
-Originators**: The study was initiated by Daniel Falster and Remko Duursma.
-Diego Barneche, Rich FitzJohn and Angelica Vårhammar assisted with literature searching, data compilation, and coding.
-A separate compilation focussing on natural forests in Japan, was collected by Masae Ishihara, with assistance from Tsutom Hiura, Hajime Utsugi and Hiroyuki Tanouchi.
+Originators**: The study was initiated by D.S.
+Falster and R.A.
+Duursma.
+D.R.
+Barneche, R.G.
+FitzJohn and A.
+Vårhammar assisted with literature searching, data compilation, and coding.
+A separate compilation focussing on natural forests in Japan, was collected by M.
+Ishihara, with assistance from T.
+Hiura, H.
+Utsugi and H.
+Tanouchi.
 All other authors contributed data to the project.
@@
 
-D.S. Falster, A. Vårhammer and D. R. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A. Duursma. R FitzJohn was employed on a XXXXXX.
+D.S.
+Falster, A.
+Vårhammer and D.R.
+Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A.
+Duursma.
+R.G.
+FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).
 

### 2014-06-19 ms/MS-fun.R

@@ -129,7 +129,7 @@ getLastName  <-  function(authorNames) {
 	}
 
 lisOfAuthors  <-  function(data) {
-	firstAuthors   <-  c('Daniel Falster', 'Remko A. Duursma', 'Masae Ishihara', 'Diego R. Barneche', 'Rich FitzJohn', 'Angelica Vårhammar')
+	firstAuthors   <-  c('Daniel S. Falster', 'Remko A. Duursma', 'Masae Ishihara', 'Diego R. Barneche', 'Rich G. FitzJohn', 'Angelica Vårhammar')
 	dataAuthors <- unique(do.call(c, lapply(unique(data$studyName), getPersonell)))
 
 	dataAuthors <- dataAuthors[!dataAuthors %in% firstAuthors]

### 2014-06-19 ms/ms.Rmd
Work on paper, citations. #119

@@
 
-# library(pander)
+bib <- read.bibtex("refs.bib")
 
@@
 
-Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(c(Falster2011="10.1111/j.1365-2745.2010.01735.x"))`.
+Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(c("10.1111/j.1365-2745.2010.01735.x"))`.
 The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants.
-While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications, or as totals for a forest plot.
-Until now, much of the raw data resides piecemeal on the hard drives of individual scientists spread around the world.
-
-Several studies have gathered together the fitted (allometric) equations for separate datasets, yet few have attempted to organize and share the raw individual plant data underpinning these equations.
-Gathered together, such data would be invaluable and represent an important resource for the community.
+While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications.
+Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
+Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(c("10.3832/ifor0901-006","10.1016/S0378-1127(97)00019-4"))`, `citep(bib["zianis_biomass_2005"])`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
 
@@
 
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c(Michener1997="10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2", Kandel2011="10.1145/1978942.1979444"))`.
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518"))`.
 This compilation would not have been possible without several new technologies.
@@
 
-Until recently, scientific publications in biology have mostly been concerned with "small data" `r citep(c(VanNoorden2013="10.1038/nj7461-243a"))`.
+Until recently, scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`.
 Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
@@
 
-**Latest update**: xx for all data files.
+**Latest update**: xxxx for all data files.
 
@@
 
-**Data verification**:
+**Data verification**: Data have passed through several rounds of review: i) initial analysis by data contributors, ii) Review by compilers of database, iii) Additional review by contributors using automated reports provided in XXXXX.
 
@@
 are denoted as 'NA'.
-Header information:
 
@@
 
-### C. Number of observations by variable
+**Number of observations by variable**
 

### 2014-06-19 ms/ms.Rmd
missing r in citation command

@@
 Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
-Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(c("10.3832/ifor0901-006","10.1016/S0378-1127(97)00019-4"))`, `citep(bib["zianis_biomass_2005"])`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
+Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(c("10.3832/ifor0901-006","10.1016/S0378-1127(97)00019-4"))`, `r citep(bib["zianis_biomass_2005"])`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
 Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.

### 2014-06-19 reports/report.Rmd
Update reports to use new tables names

@@
 
-**Data contributor:** `r .dat$contact$name`.
+**Data contributor:** `r .dat$contacts$name`.
 
-**Email:** `r .dat$contact$email`.
+**Email:** `r .dat$contacts$email`.
 
-**Address:** `r .dat$contact$address`.
+**Address:** `r .dat$contacts$address`.
 
@@
 
-Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.
+Data for `r length(.dat$data$studyName)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.
 

### 2014-06-19 ms/ms.Rmd
Added more info to data verification, and citations to Taxonstand and taxize. For #122

@@
 
-**Data verification**: Data have passed through several rounds of review: i) initial analysis by data contributors, ii) Review by compilers of database, iii) Additional review by contributors using automated reports provided in XXXXX.
+**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was approppriately matched to variables in BAAD, and to minimize outliers.
+
+i) Initial analysis by data contributors. The majority of studies have been published and as such have gone through quality assurance by the data contributor.
+ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
+iii) All species names were matched against online databases with the Taxonstand R package, `r citet("10.1111/j.2041-210X.2012.00232.x")`and the taxize R package, `r citet(bib["chamberlain2013taxize"])`.
+iv) Additional review by contributors using automated repots.
+For each study, we generated automated reports using the dataMashR R package (XXXX).
+In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, meta-data provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).
+Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.
 

### 2014-06-19 ms/ms.Rmd
added a citation

@@
 
-Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(c("10.1111/j.1365-2745.2010.01735.x"))`.
+Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`.
 The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants.

### 2014-06-20 ms/ms.Rmd
more citations in intro (ms.Rmd)

--- 2014-06-19 ms/ms.Rmd
@@
 
-Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(c("10.1111/j.1365-2745.2010.01735.x"))`.
-The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants.
+Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`.
+The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
 While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications.
 Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
-Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(c("10.3832/ifor0901-006","10.1016/S0378-1127(97)00019-4"))`, `r citep(bib["zianis_biomass_2005"])`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
+Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4","10.3832/ifor0901-006",bib["zianis_biomass_2005"]))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.
 
@@
 It is exciting to be able to put these parcels together, to create something larger.
-We hope other people will use this dataset and build on it.
+We hope others will use this dataset and build on it.
 See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.

### 2014-06-20 ms/ms.Rmd
add reference

@@
+---
+output: html_document
+---
 # *Ecological Archives* Exxxxx
@@
 library(whisker)
+library(scales)
+library(maps)
 
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-writeLines(paste("**List of Authors:**", lisOfAuthors(baad$contact)))
+writeLines(paste("**List of Authors:**", pasteC(listOfAuthors(baad$contact), ', ')))
 ```
@@
 Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`.
-The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
+The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
 While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications.
 Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
-Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4","10.3832/ifor0901-006",bib["zianis_biomass_2005"]))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.
+Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
 
@@
 
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518"))`.
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))`.
 This compilation would not have been possible without several new technologies.
@@
 See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
+
+```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6}
+drawWorldPlot  <-  function(data, horlines=TRUE, sizebyn=FALSE, add=FALSE,
+                            pchcol="red",
+                            legend=TRUE){
+
+  if(!add){
+    map('world',col="grey80",bg="white",lwd=0.5,fill=TRUE,resolution=0,wrap=TRUE, border="grey80")
+    map('world',col="black",boundary=TRUE,lwd=0.2,interior=FALSE,fill=FALSE,add=TRUE,resolution=0,wrap=TRUE)
+  }
+
+  if(horlines){
+    lines(c(-180,180),c(-100,-100), lty="dashed", xpd=TRUE, lwd=2)
+    lines(c(-180,180),c(100,100), lty="dashed", xpd=TRUE, lwd=2)
+  }
+
+  # Remove all duplicates (increases speed and minimizes file size)
+  latlon <- with(data, paste(latitude, longitude))
+  lat <- data$latitude[!duplicated(latlon)]
+  lon <- data$longitude[!duplicated(latlon)]
+
+  j  <-  !is.na(lat) & !is.na(lon)
+
+  # Location only sometimes missing - but lat/lon can still be in dataset anyway.
+  # & data$loc != "NA" | is.na(data$loc)
+
+  if(!any(j)){
+    polygon(c(-100,95,95,-100), c(-10,-10,15,15), col=rgb(0,0,0,240,maxColorValue=255))
+    text(-100, 0, expression(paste(bold("Missing coordinate/location"))), col="red", xpd=TRUE, pos=4, cex=0.8)
+  } else {
+
+    if(!sizebyn){
+      points(lon,lat, pch=19, col=alpha("blue",0.5), cex=1.6)
+    } else {
+      n <- table(latlon)
+      symbols(lon,lat, circles=log10(n), inches=0.1, fg="black", bg=pchcol, add=TRUE)
+
+      if(legend){
+        ns <- c(10,100,1000)
+        X <- rep(-170,3)
+        Y <- seq(-30,-10,by=10)
+        rect(xleft=-200, xright=-120, ybottom=-50, ytop=10, col="white", border=NA)
+        symbols(x=X, y=Y, circles=log10(ns), inches=0.1,
+                fg="black", bg=pchcol, add=TRUE)
+        text(x=X+5, y=Y, labels=as.character(ns),pos=4)
+        text(x=X+5, y=Y[3]+10, labels=expression(italic(n)), pos=4)
+      }
+
+    }
+
+  }
+
+}
+
+# Colored by vegetation type
+  cols <- alpha(c("blue","brown","forestgreen","darkorange","dodgerblue2",
+                    "navajowhite4","lightgoldenrod3","lavender"),0.67)
+  vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
+  vegs_labels <- c("Boreal forest","Savannah","Temperate forest",
+                   "Temperate rainforest","Tropical rainforest",
+                   "Tropical seasonal forest","Woodland","Glasshouse / Common garden")
+
+  drawWorldPlot(data[data$vegetation == vegs[1],], horlines=FALSE,
+                sizebyn=TRUE, pchcol=cols[1])
+  for(i in 2:length(vegs)){
+    drawWorldPlot(data[data$vegetation == vegs[i],], horlines=FALSE,
+                  sizebyn=TRUE, pchcol=cols[i], add=TRUE)
+  }
+  drawWorldPlot(data[data$growingCondition == "GH",], horlines=FALSE, sizebyn=TRUE,
+                pchcol=cols[8], add=TRUE)
+  par(xpd=NA)
+  legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
+  legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')
+
+```
+
+**Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.
 
@@
 
-allometry; trait; leaf area; height; size; architecture; vegetation
-model; carbon; growth; wood
+plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle
 
@@
 
-Count <- function(x)length(data[[x]][!is.na(data[[x]])])
-N <- sapply(thesevars, Count)
+Count_obs <- function(x)length(data[[x]][!is.na(data[[x]])])
+N <- sapply(thesevars, Count_obs)
+
+Count_studies <- function(x)length(unique(data[["studyName"]][!is.na(data[[x]])]))
+Nstud <- sapply(thesevars, Count_studies)
+
+df <- rbind.fill(apply(data[,thesevars], 2, function(x)data.frame(Min=min(x, na.rm=T),Max=max(x, na.rm=T),Median=median(x, na.rm=T))))
 
 dfr <- data.frame(Variable=Labels, N=N)
+dfr <- cbind(dfr,df)
+
 dfr <- dfr[dfr$N > 0,]
 rownames(dfr) <- NULL
+
+f <- function(x)formatC(x, digits=2, format="fg")
+dfr$Min <- f(dfr$Min)
+dfr$Max <- f(dfr$Max)
+dfr$Median <- f(dfr$Median)
+
 ```
@@
 ```{r results='asis', echo=FALSE}
-kable(dfr, "html", align="l")
+options(digits=9)
+kable(dfr, "markdown", align="l", padding=10)
 ```

### 2014-06-20 ms/ms.Rmd
more citations, closes #119

@@
 Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`.
-The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
+The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
 While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications.
 Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
-Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4","10.3832/ifor0901-006",bib["zianis_biomass_2005"]))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
+Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
 Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.

### 2014-06-20 reports/report.Rmd
report tables improved

@@
 
-```{r echo=FALSE, comment=''}
+```{r echo=FALSE, comment='',results='asis'}
 counts  <-  apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
@@
 counts  <-  counts[names(counts)%in%.dat$dictionary$Variable[.dat$dictionary$Group=="tree"]]
-cbind(counts)
+kable(cbind(counts))
 ```
@@
 **Location data**
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment='',results='asis'}
 options(width =160)
-locLevelInfo(.dat$data)
+kable(locLevelInfo(.dat$data))
 ```

### 2014-06-20 reports/report-fun.R

@@ -415,6 +415,8 @@ standLevelInfo <- function(data) {
         sta <- sta[, -which(names(sta) == "grouping")]
 
     sta[is.na(sta) | sta == "" | sta == "NA"] <- "missing"
+    rownames(sta) <- NULL
+    
   sta
 }
 

### 2014-06-21 ms/ms.Rmd
Changed keywords. Closes #121
Added colored world plot to Introduction; Nstudies added back to table at the end

--- 2014-06-20 ms/ms.Rmd
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-writeLines(paste("**List of Authors:**", pasteC(listOfAuthors(baad$contact), ', ')))
+writeLines(paste("**List of Authors:**", lisOfAuthors(baad$contact)))
 ```
@@
 Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.
 
@@
 
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))`.
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518"))`.
 This compilation would not have been possible without several new technologies.
@@
 
-dfr <- data.frame(Variable=Labels, N=N)
+dfr <- data.frame(Variable=Labels, N=N, Nstudies=Nstud)
 dfr <- cbind(dfr,df)

### 2014-06-25 reports/report.Rmd
Refine reports
report2

--- 2014-05-13 reports/report.Rmd
@@
 
-**Data contributor:** `r .dat$contact$name`.
+**Data contributor:** `r .dat$contacts$name`
 
-**Email:** `r .dat$contact$email`.
+**Email:** `r paste0(.dat$contacts$email[.dat$contacts$email!=""], collapse = ", ")`
 
-**Address:** `r .dat$contact$address`.
+**Address:**
 
-## Citation
-Information about the primary reference for the data you provided.
+`r paste0("- ", .dat$contacts$address, collapse = "\n")`
 
-**Reference:** `r .dat$references$citation`
+## Data source
 
-**DOI:** `r  bibGetElement(.dat$bib, element = "doi")`
+**Citation:** `r .dat$references$citation`
 
-**Abstract:** `r  bibGetElement(.dat$bib, element = "abstract")`
+**DOI:** `r  paste0('[', bibGetElement(.dat$bibtex, element = "doi"), '](http://doi.org/', bibGetElement(.dat$bibtex, element = "doi"),')')`
+
+**Abstract:** `r  bibGetElement(.dat$bibtex, element = "abstract")`
 
@@
 
-Data for `r length(.dat$data$dataset)` individuals from `r length(unique(.dat$data$location))` locations covering `r length(unique(.dat$data$species))` species.
+The dataset includes records for `r length(.dat$data$studyName)` individuals from `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` family(ies), presenting `r length(unique(.dat$data$pft))` functional type(s), growing in `r length(unique(.dat$data$growingCondition))` condition(s) within `r length(unique(.dat$data$vegetation))` major type(s) of habitat, with data included for the following variables:
 
-The dataset includes `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` family(ies), presenting `r length(unique(.dat$data$pft))` functional type(s), growing in `r length(unique(.dat$data$growingCondition))` condition(s) within `r length(unique(.dat$data$vegetation))` major type(s) of habitat:
-
-### Number of records for each variable
-
-```{r echo=FALSE, comment=''}
-counts  <-  apply(!is.na(.dat$data), MARGIN=2, FUN = sum)
-counts  <-  counts[counts>0]  #only include no zeros
-counts  <-  counts[names(counts)%in%.dat$dictionary$Variable[.dat$dictionary$Group=="tree"]]
-cbind(counts)
+```{r results='asis', echo=FALSE, warning=FALSE}
+options(digits=9)
+kable(
+	summaryTable(.dat$data, var.def= read.csv("../config/variableDefinitions.csv", stringsAsFactors=FALSE)),
+	 "markdown", align="l", padding=10)
 ```
 
-## Study details
+## Locations sampled
 
-### Location
+Data were sampled at the following locations:
 
-**Location data**
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment=''}
-options(width =160)
-locLevelInfo(.dat$data)
-```
-
-**Global map**
-
-```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
+```{r echo=FALSE, fig.width=12, fig.height=10, message=FALSE, warning=FALSE}
 prepdata  <-  prepMapInfo(alldata$data, .study)
@@
 
-**Country map**
+And locally within the country:
 
-```{r echo=FALSE, fig.width=8, fig.height=6, message=FALSE, warning=FALSE}
+```{r echo=FALSE, fig.width=12, fig.height=10, message=FALSE, warning=FALSE}
 repMissingInfo(prepdata)
@@
 
-**Stand description**
-Information about the growing condition of plants in your study, i.e. their local environment.
+The sites sampled are:
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, comment='',results='asis'}
+kable(locLevelInfo(.dat$data),"markdown", align="l", padding=2)
+```
+
+The growing conditions of sampled plants was:
 
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
-kable(standLevelInfo(.dat$data))
+kable(standLevelInfo(.dat$data),"markdown", align="l", padding=2)
 ```
 
-### Species data
+## Species sampled
 
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment='',results="asis"}
-kable(spLevelInfo(.dat$data))
+kable(spLevelInfo(.dat$data), "markdown", align="l", padding=2)
 ```
 
-### Methods
+## Methods used
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
 metaTable  <-  getMeta(.study)
-writeLines(paste0(metaTable[,1], ": ", metaTable[,2], "\n"))
+writeLines(paste0("**",metaTable[,1],"**", ": ", metaTable[,2], "\n"))
 ```
@@
 
-This is how the study `r .study` (each colour represents a species) fits in the entire dataset (grey).
+This is how the study `r .study` fits in the entire dataset (grey). each colour represents a species. A legend of species names with colours is included at the end for reports with 1 < n < 20 species.
 
-```{r fig.height=8, fig.width=8, echo=FALSE, warning=FALSE, message=FALSE}
+```{r fig.height=12, fig.width=12, echo=FALSE, warning=FALSE, message=FALSE}
 makePlotPanel(alldata, .dat)

### 2014-06-25 reports/report-fun.R

@@ -1,6 +1,8 @@
 library(knitr, quietly = TRUE)
 library(maps, quietly = TRUE)
 library(mapdata, quietly = TRUE)
+library(plyr, quietly = TRUE)
+
 suppressPackageStartupMessages(library(gdata))
 
 printAllStudyReports <- function(data = readRDS("../output/baad.rds"),
@@ -27,15 +29,8 @@ printAllStudyReports <- function(data = readRDS("../output/baad.rds"),
 printStudyReport <- function(alldata, study, RmdFile = "report.Rmd", path = "output/report-by-study",
     name = study, delete = TRUE, reprocess = FALSE) {
 
-    knitThis(RmdFile = RmdFile, path = path, name = name, delete = TRUE, reprocess = reprocess,
-        predefined = list(.study = study, alldata = alldata, .dat = extractStudy(alldata,
-            study)))
-}
-
-# creates html reports using knitr, copies output file to desired location and
-# renames as required
-knitThis <- function(RmdFile = "reportmd.Rmd", path = "output/report-per-study",
-    name = "study", delete = TRUE, reprocess = TRUE, ..., predefined = list(...)) {
+    predefined = list(.study = study, alldata = alldata, .dat = extractStudy(alldata,
+            study))
 
     outputfile <- paste0(path, "/", name, ".html")
 
@@ -52,7 +47,8 @@ knitThis <- function(RmdFile = "reportmd.Rmd", path = "output/report-per-study",
         }
 
         # knit
-        suppressMessages(knit2html(RmdFile, quiet = TRUE, envir = e))
+         knit(RmdFile, quiet = TRUE, envir = e)
+         rmarkdown::render(sub("Rmd", "md", RmdFile), "html_document")
 
         # copy html file to output dir
         if (!file.exists(path))
@@ -79,7 +75,7 @@ extractStudy <- function(alldata, study) {
     for (var in c("data", "contacts", "references"))
         alldata[[var]] <- alldata[[var]][alldata[[var]]$studyName ==study, ]
 
-    alldata[["bib"]] <- alldata[["bib"]][[study]]
+    alldata[["bibtex"]] <- alldata[["bibtex"]][[study]]
 
     alldata
 }
@@ -113,21 +109,21 @@ makePlotPanel <- function(data_all, data_study, dir = "report-per-study") {
             }
         }
     }
-  
+
[…]

### 2014-06-25 ms/ms.Rmd
Refine paper

--- 2014-06-20 ms/ms.Rmd
@@
 ```{r, results="hide", echo=FALSE, message=FALSE}
-library(knitr, quietly=TRUE)
+source("MS-fun.R")
+
 knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")
 
-library(plyr)
-library(bibtex)
-library(knitcitations)
-library(Hmisc)
-library(whisker)
-
 bib <- read.bibtex("refs.bib")
-
-source("MS-fun.R")
-```
-
-```{r, results="hide", echo=FALSE, message=FALSE}
-
 baad <- readRDS("../output/baad.rds")
 data <- baad$data
-
 var.def <- read.csv("../config/variableDefinitions.csv",
                     stringsAsFactors=FALSE)
+
 vars <- setdiff(var.def$Variable[var.def$Type == "numeric"], c("map","mat","lai","age"))
@@
 Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
 
@@
 See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
+
+```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height8}
+
+# World plot colored by vegetation type
+cols <- make.transparent(c(
+    "blue","brown","forestgreen","darkorange","dodgerblue2",
+    "navajowhite4","lightgoldenrod3","lavender"),0.67)
+vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
+vegs_labels <- c("Boreal forest","Savannah","Temperate forest",
+                 "Temperate rainforest","Tropical rainforest",
+                 "Tropical seasonal forest","Woodland","Glasshouse / Common garden")
+
+drawWorldPlot(data[data$vegetation == vegs[1],], sizebyn=TRUE, pchcol=cols[1])
+for(i in 2:length(vegs)){
+  drawWorldPlot(data[data$vegetation == vegs[i],], sizebyn=TRUE, pchcol=cols[i], add=TRUE)
+}
+drawWorldPlot(data[data$growingCondition == "GH",], sizebyn=TRUE,
+              pchcol=cols[8], add=TRUE)
+par(xpd=NA)
+legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
+legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')
+
+```
+
+**Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.
 
@@
 
-Identity:
-Size:
-Format and storage mode: ASCII text, comma delimited. Missing values
-are denoted as 'NA'.
+BAAD contains the following files:
 
-### B. Variable definitions
+- `dictionary.csv`: a table of variable definitions
+- `data.csv`: amalgamated dataset, with columns as defined in `dictionary.csv`
+- `references.bib`: bibtex entries containing the primary source for each study
+- `references.csv`: summary of bibtex file, giving DOI, URL and citation of  the primary source for each study
+- `contacts.csv`: list of contacts and affiliations for each study
+
+### B. Variable definitions for data.csv
 
@@
 		paste0("* ", x$Variable,
+    "\n\t- Type: ", x$Type,
 		"\n\t- Label: ", x$label,
 		"\n\t- Description: ", x$Description,
-		"\n\t- Type: ", x$Type,
 		"\n\t- Units: ", gsub(".","/", x$UnitsMD, fixed=TRUE)
@@
 
-```{r echo=FALSE, message=FALSE}
-
-thesevars <- setdiff(var.def$Variable[var.def$Type=="numeric"],
-                     c("map","mat","lai"))
-Labels <- capitalize(var.def$label[var.def$Variable %in% thesevars])
-
-Count_obs <- function(x)length(data[[x]][!is.na(data[[x]])])
-N <- sapply(thesevars, Count_obs)
-
-Count_studies <- function(x)length(unique(data[["studyName"]][!is.na(data[[x]])]))
-Nstud <- sapply(thesevars, Count_studies)
-
-df <- rbind.fill(apply(data[,thesevars], 2, function(x)data.frame(Min=min(x, na.rm=T),Max=max(x, na.rm=T),Median=median(x, na.rm=T))))
-
-dfr <- data.frame(Variable=Labels, N=N)
-dfr <- cbind(dfr,df)
-
-dfr <- dfr[dfr$N > 0,]
-rownames(dfr) <- NULL
-
-f <- function(x)formatC(x, digits=2, format="fg")
-dfr$Min <- f(dfr$Min)
-dfr$Max <- f(dfr$Max)
-dfr$Median <- f(dfr$Median)
-
-```
-
-```{r results='asis', echo=FALSE}
+```{r results='asis', echo=FALSE, warning=FALSE}
 options(digits=9)
-kable(dfr, "markdown", align="l", padding=10)
+kable(summaryTable(data, var.def), "markdown", align="l", padding=10)
 ```

### 2014-06-25 ms/MS-fun.R

@@ -1,3 +1,8 @@
+library(knitr, quietly = TRUE)
+library(maps, quietly = TRUE)
+library(bibtex, quietly = TRUE)
+library(knitcitations, quietly = TRUE)
+
 ## Mimic dataMashR function here, but relative to *this* directory.
 data.path <- function(...) {
   file.path("../data", ...)
@@ -137,3 +142,99 @@ listOfAuthors  <-  function(data) {
 }
 
 
+capitalize <- function (string) {
+    capped <- grep("^[^A-Z]*$", string, perl = TRUE)
+    substr(string[capped], 1, 1) <- toupper(substr(string[capped],
+        1, 1))
+    string
+}
+
+## Make colours semitransparent:
+make.transparent <- function(col, opacity = 0.5) {
+    if (length(opacity) > 1 && any(is.na(opacity))) {
+        n <- max(length(col), length(opacity))
+        opacity <- rep(opacity, length.out = n)
+        col <- rep(col, length.out = n)
+        ok <- !is.na(opacity)
+        ret <- rep(NA, length(col))
+        ret[ok] <- Recall(col[ok], opacity[ok])
+        ret
+    } else {
+        tmp <- col2rgb(col)/255
+        rgb(tmp[1, ], tmp[2, ], tmp[3, ], alpha = opacity)
+    }
+}
+
+drawWorldPlot  <-  function(data, sizebyn=FALSE, add=FALSE,
+                            pchcol="red",
+                            legend=TRUE){
+
+
+  require(maps)
+  if(!add){
+    map('world',col="grey80",bg="white",lwd=0.5,fill=TRUE,resolution=0,wrap=TRUE, border="grey80", ylim=c(-80, 80), mar=rep(0,4))
+    map('world',col="black",boundary=TRUE,lwd=0.2,interior=FALSE,fill=FALSE,add=TRUE,resolution=0,wrap=TRUE)
+  }
+
+
+  # Remove all duplicates (increases speed and minimizes file size)
+  latlon <- with(data, paste(latitude, longitude))
+  lat <- data$latitude[!duplicated(latlon)]
+  lon <- data$longitude[!duplicated(latlon)]
+
+  j  <-  !is.na(lat) & !is.na(lon)
+
+  # Location only sometimes missing - but lat/lon can still be in dataset anyway.
[…]

### 2014-06-25 email.R
Spelling, short link

@@ -19,7 +19,7 @@ You kindly contributed data from the publication: ", dat$references$citation[1],
 
 We are delighted to be (finally) sharing a draft of the paper with you for comment. Without a doubt, this compilation has also emerged as an extraordinary resource of which we can all be rightly very proud.
 
-To assist us in communicating with 85 co-authors, we are asking that you return your comments on the paper by filling out this form within the next 2 weeks: https://docs.google.com/forms/d/1TD_UV8Zj1mRBas14C_EH4F0wSxLyVCLpFMF5STptgE4/viewform?usp=send_form
+To assist us in communicating with 85 co-authors, we are asking that you return your comments on the paper by filling out this form within the next 2 weeks: http://bit.ly/BAAD-feedback
 If you cannot reply within this time please indicate when you will be able to respond.
 
 We look forwards to hearing from you shortly.

### 2014-07-07 ms/ms.Rmd
Spelling, short link
dataMashR update: Use lower case for names of data dictionary. Closes 128

@@
 
-vars <- setdiff(var.def$Variable[var.def$Type == "numeric"], c("map","mat","lai","age"))
+vars <- setdiff(var.def$variable[var.def$type == "numeric"], c("map","mat","lai","age"))
 ```
@@
 FitzJohn and A.
-Vårhammer) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
+Vårhammar) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
 The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.
@@
 x <-daply(var.def, 1, function(x)
-		paste0("* ", x$Variable,
-    "\n\t- Type: ", x$Type,
+		paste0("* ", x$variable,
+    "\n\t- Type: ", x$type,
 		"\n\t- Label: ", x$label,
-		"\n\t- Description: ", x$Description,
-		"\n\t- Units: ", gsub(".","/", x$UnitsMD, fixed=TRUE)
+		"\n\t- Description: ", x$description,
+		"\n\t- Units: ", gsub(".","/", x$unitsMD, fixed=TRUE)
 		))
 
-writeLines(x[var.def$Variable])
+writeLines(x[var.def$variable])
 ```

### 2014-07-07 ms/MS-fun.R

@@ -215,7 +215,7 @@ drawWorldPlot  <-  function(data, sizebyn=FALSE, add=FALSE,
 
 summaryTable <- function(data, var.def, digits = 2){
 
-	thesevars <- setdiff(var.def$Variable[var.def$Type=="numeric"], c("map","mat","lai"))
+	thesevars <- setdiff(var.def$variable[var.def$type=="numeric"], c("map","mat","lai"))
 
 	N <- sapply(thesevars, function(x)length(data[[x]][!is.na(data[[x]])]))
 
@@ -229,9 +229,9 @@ summaryTable <- function(data, var.def, digits = 2){
 	for(v in c("Min", "Max", "Median"))
 		dfr[[v]] <- formatC(dfr[[v]], digits=digits, format="fg")
 
-	dfr[["Units"]] <- var.def$UnitsMD[match(thesevars, var.def$Variable)]
-	dfr[["Variable"]] <- var.def$Variable[match(thesevars, var.def$Variable)]
-	dfr[["Label"]] <- capitalize(var.def$label[match(thesevars, var.def$Variable)])
+	dfr[["Units"]] <- var.def$unitsMD[match(thesevars, var.def$variable)]
+	dfr[["Variable"]] <- var.def$variable[match(thesevars, var.def$variable)]
+	dfr[["Label"]] <- capitalize(var.def$label[match(thesevars, var.def$variable)])
 
 	dfr <- dfr[dfr$N > 0,c("Variable",  "Label", "Units", "N", "Studies","Min","Median","Max")]
 	rownames(dfr) <- NULL

### 2014-07-07 reports/report-fun.R

@@ -85,11 +85,11 @@ makePlotPanel <- function(data_all, data_study, dir = "report-per-study") {
     dict <- data_all$dictionary
 
     labels <- as.list(paste0(dict$label, " (", dict$Units, ")"))
-    names(labels) <- dict$Variable
+    names(labels) <- dict$variable
 
     # determine variables to plot
-    plot.vars <- dict$Variable[dict$Group == "tree" & dict$Type == "numeric" & colSums(!is.na(data_study$data[,
-        dict$Variable])) > 0]
+    plot.vars <- dict$variable[dict$group == "tree" & dict$type == "numeric" & colSums(!is.na(data_study$data[,
+        dict$variable])) > 0]
 
     # set up a vector of colors, each species with different color
     species <- as.numeric(as.factor(data_study$data$species))
@@ -480,7 +480,7 @@ getMeta <- function(study) {
 
 summaryTable <- function(data, var.def, digits = 2){
 
-    thesevars <- setdiff(var.def$Variable[var.def$Type=="numeric"], c("map","mat","lai"))
+    thesevars <- setdiff(var.def$variable[var.def$type=="numeric"], c("map","mat","lai"))
 
     N <- sapply(thesevars, function(x)length(data[[x]][!is.na(data[[x]])]))
 
@@ -493,9 +493,9 @@ summaryTable <- function(data, var.def, digits = 2){
     for(v in c("Min", "Max", "Median"))
         dfr[[v]] <- formatC(dfr[[v]], digits=digits, format="fg")
 
-    dfr[["Units"]] <- var.def$Units[match(thesevars, var.def$Variable)]
-    dfr[["Variable"]] <- var.def$Variable[match(thesevars, var.def$Variable)]
-    dfr[["Label"]] <- capitalize(var.def$label[match(thesevars, var.def$Variable)])
+    dfr[["Units"]] <- var.def$Units[match(thesevars, var.def$variable)]
+    dfr[["Variable"]] <- var.def$variable[match(thesevars, var.def$variable)]
+    dfr[["Label"]] <- capitalize(var.def$label[match(thesevars, var.def$variable)])
 
     dfr <- dfr[dfr$N > 0,c("Variable",  "Label", "Units", "N", "Min","Median","Max")]
     rownames(dfr) <- NULL

### 2014-07-11 ms/ms.Rmd
few changes/suggestions to ms, changed language to US English

@@
 Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`.
-The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
-While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications.
+The allocation of reduced carbon into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
+While many scientific studies have quantified plant traits via destructive methods, most of these data have only been made available in the form of summary tables or figures included in publications.
 Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
 Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
 
@@
 FitzJohn and A.
-Vårhammar) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
+V??rhammar) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
 The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.
 Moreover, we were invited to incorporate another compilation led by M.
-Ishihara and focussing on Japanese literature.
+Ishihara and focusing on Japanese literature.
 As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies.
@@
 Our entire work flow is scripted, using the R computing environment.
-The database can be rebuilt in an instant from the raw datafiles, using a single command.
+The database can be rebuilt in an instant from the raw data files, using a single command.
 When a single data contribution has been modified, the final master dataset is thus easily recreated.
@@
 FitzJohn and A.
-Vårhammar assisted with literature searching, data compilation, and coding.
-A separate compilation focussing on natural forests in Japan, was collected by M.
+V??rhammar assisted with literature searching, data compilation, and coding.
+A separate compilation focusing on natural forests in Japan, was collected and/or compiled by M.
 Ishihara, with assistance from T.
@@
 **2.
-Abstract**: Understanding how different plants are plant constructed -- i.e.
-how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modelling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere.
+Abstract**: Understanding how different plants are constructed -- i.e.
+how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere.
 Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
-While a variety of models dealing with allocation exist, a synthetic understanding about plant allocation is lacking.
-This is partly due to the lack of suitable datasets for validating and parametrising models.
+While a variety of models dealing with carbon allocation exist **(Cite references?)**, we still lack a synthetic understanding about the underlying processes.
+This is partly due to the lack of suitable datasets for validating and parameterizing models.
 To that end, we present the Biomass And Allometry Database (BAAD) for woody plants.
@@
 The transformations were achieved using a new R package dataMashR.
-Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
-not indirectly via allometric equations (with a few exceptions large trees which were sub-sampled and scaled up).
+Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100 m in height; and iii) biomass was estimated directly, i.e.
+not indirectly via allometric equations, except in very large trees where biomass was first sub-sampled and then scaled up to the entire individual.
 We included both wild and artificially grown plants.
@@
 We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, status, light, experimental treatment, vegetation type) where available.
-It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the worlds vegetation.
+It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling.
 
@@
 
-**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was approppriately matched to variables in BAAD, and to minimize outliers.
+**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in BAAD, and to minimize outliers.
 
@@
 ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
-iii) All species names were matched against online databases with the Taxonstand R package, `r citet("10.1111/j.2041-210X.2012.00232.x")`and the taxize R package, `r citet(bib["chamberlain2013taxize"])`.
-iv) Additional review by contributors using automated repots.
+iii) All species names were matched against online databases with the Taxonstand and taxize R packages, `r citet(c("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
+iv) Additional review by contributors using automated reports.
 For each study, we generated automated reports using the dataMashR R package (XXXX).
-In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, meta-data provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).
+In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, metadata provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).
 Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.
@@
 As such, it may be freely used and redistributed.
-When using the dataset, we kindly request that you cite this article, recognising the hard work that went into collecting the data and the authors' willingness to make it publicly available.
+When using the dataset, we kindly request that you cite this article, recognizing the hard work that went into collecting the data and the authors' willingness to make it publicly available.
 
@@
 Falster, A.
-Vårhammer and D.R.
+V??rhammer and D.R.
 Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A.

### 2014-07-15 ms/ms.Rmd
small fixes

--- 2014-07-07 ms/ms.Rmd
@@
 ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
-iii) All species names were matched against online databases with the Taxonstand R package, `r citet("10.1111/j.2041-210X.2012.00232.x")`and the taxize R package, `r citet(bib["chamberlain2013taxize"])`.
-iv) Additional review by contributors using automated repots.
-For each study, we generated automated reports using the dataMashR R package (XXXX).
+iii) All species names were matched against online databases with the Taxonstand R package, `r citet("10.1111/j.2041-210X.2012.00232.x")` and the taxize R package, `r citet(bib["chamberlain2013taxize"])`.
+iv) Additional review by contributors using reports generated from the data.
 In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, meta-data provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).

### 2014-07-15 ms/MS-fun.R

@@ -118,7 +118,7 @@ studyDetails  <-  function(data) {
                       siteHistory=pasteC(getSiteHistory(unique(data$growingCondition)),'; '),
            	          metadataDesign=getMetadataDesign(unique(data$studyName)),
            	          lab=getMetadataMethods(unique(data$studyName)),
-           	          personell=pasteC(getPersonell(unique(data$studyName))),
+           	          personell=pasteC(getPersonell(unique(data$studyName)),", "),
            	          citation=getCitation(data$studyName[1]),
            	          vars=pasteC(names(removeNAcols(data,
            	          	exclude = c("studyName","species","speciesMatched","location", "latitude","longitude","vegetation","map","mat","family","pft","growingCondition", "grouping"))), ", ")

### 2014-07-15 ms/ms.Rmd
Add missing quotes

@@
 library(whisker)
+
+## for debugging
+# for(s in unique(data$studyName)){
+#  cat(s)
+#  studyDetails(data[data$studyName==s,])}
+
 details2B  <-  daply(data, .(studyName), function(x) studyDetails(x))

### 2014-09-16 ms/ms.Rmd
Add acknowledgments. Closes #136

@@
 FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).
+M.
+Ishihara was supported by the Environmental Research and Technology Development Fund (S-9-3) of the Ministry of the Environment, Japan.
+We thank Takahiro Inoue, Hayato Aoyama and Shin-ichiro Aiba for valuable assistance.
+In addition, we thank the authors thank the following people for contributions to the following studies:
+
+```{r,results="asis",echo=FALSE, message=FALSE}
+
+getAcknowledgments <- function(studyName) {
+  metadata  <-  readStudyFile(studyName, "studyMetadata.csv")
+  i <- match("Acknowledgements",metadata$Topic)
+  out <- ""
+  if(!is.na(i))
+    out <- sprintf("- %s: %s.\n", studyName, metadata$Description[i])
+  out
+}
+writeLines(paste(sapply(unique(data$studyName), getAcknowledgments), collapse= ""))
+```
 

### 2014-09-16 ms/ms.Rmd
Incorporate suggestions to MS from Diego. Closes #130

@@
 The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
-While many studies in have destructively harvested woody plants in the name of science, most of these data have only been made available in the form of summary tables or figures included in publications.
+While many scientific studies have destructively harvested woody plants, most of the data collected have only been made available in the form of summary tables or figures included in publications.
 Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
 Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognised need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
 
@@
 Moreover, we were invited to incorporate another compilation led by M.
-Ishihara and focussing on Japanese literature.
+Ishihara and focusing on Japanese literature.
 As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies.
@@
 Vårhammar assisted with literature searching, data compilation, and coding.
-A separate compilation focussing on natural forests in Japan, was collected by M.
+A separate compilation focusing on natural forests in Japan, was compiled by M.
 Ishihara, with assistance from T.
@@
 Abstract**: Understanding how different plants are plant constructed -- i.e.
-how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modelling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere.
+how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere.
 Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
-While a variety of models dealing with allocation exist, a synthetic understanding about plant allocation is lacking.
+While a variety of models dealing with carbon and biomass allocation exist, we still lack a synthetic understanding of the underlying processes.
 This is partly due to the lack of suitable datasets for validating and parametrising models.
@@
 Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
-not indirectly via allometric equations (with a few exceptions large trees which were sub-sampled and scaled up).
+not indirectly via allometric equations (except in very large trees where biomass was estimated from detailed sub-sampling).
 We included both wild and artificially grown plants.
@@
 We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, status, light, experimental treatment, vegetation type) where available.
-It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the worlds vegetation.
+It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the world's vegetation.
 
@@
 ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
-iii) All species names were matched against online databases with the Taxonstand R package, `r citet("10.1111/j.2041-210X.2012.00232.x")` and the taxize R package, `r citet(bib["chamberlain2013taxize"])`.
+iii) All species names were matched against online databases with the R packages `Taxonstand` and `taxize` R package citet(c("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
 iv) Additional review by contributors using reports generated from the data.
-In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, meta-data provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).
+In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, metadata provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).
 Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.

### 2014-09-17 ms/ms.Rmd
Add author affiliations. Closes #159

@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-writeLines(paste("**List of Authors:**", pasteC(listOfAuthors(baad$contact), ', ')))
+details <- authorDetails(baad$contact)
+
+writeLines(paste("**List of Authors:** ", paste0(details$authors, "^", details$address_code,"^",collapse = ", ") ))
+
+writeLines("\n\n")
+
+writeLines(paste0(details$address$code, ". ", details$address$address, collapse="\n"))
 ```

### 2014-09-17 ms/MS-fun.R

@@ -2,6 +2,7 @@ library(knitr, quietly = TRUE)
 library(maps, quietly = TRUE)
 library(bibtex, quietly = TRUE)
 library(knitcitations, quietly = TRUE)
+library(plyr, quietly = TRUE)
 
 ## Mimic dataMashR function here, but relative to *this* directory.
 data.path <- function(...) {
@@ -29,10 +30,6 @@ getMetadataMethods  <-  function(studyName) {
 		metadata  <-  metadata[!is.na(metadata$Description) & !(tolower(metadata$Topic) %in% c('sampling strategy', 'age', 'growth environment')),]
 		pasteC(paste0(metadata$Topic, ': ', metadata$Description), '\n\t- ')
 	}
-}
-
-getPersonell  <-  function(studyName) {
-	readStudyFile(studyName, "studyContact.csv")[,"name", drop=TRUE]
 }
 
 getSiteHistory <-function(code){
@@ -118,7 +115,8 @@ studyDetails  <-  function(data) {
                       siteHistory=pasteC(getSiteHistory(unique(data$growingCondition)),'; '),
            	          metadataDesign=getMetadataDesign(unique(data$studyName)),
            	          lab=getMetadataMethods(unique(data$studyName)),
-           	          personell=pasteC(getPersonell(unique(data$studyName)),", "),
+           	          personell=pasteC(
+           	          	readStudyFile(unique(data$studyName), "studyContact.csv")[,"name", drop=TRUE],", "),
            	          citation=getCitation(data$studyName[1]),
            	          vars=pasteC(names(removeNAcols(data,
            	          	exclude = c("studyName","species","speciesMatched","location", "latitude","longitude","vegetation","map","mat","family","pft","growingCondition", "grouping"))), ", ")
@@ -133,14 +131,21 @@ getLastName  <-  function(authorNames) {
 							})
 	}
 
-listOfAuthors  <-  function(data) {
-	firstAuthors   <-  c('Daniel S. Falster', 'Remko A. Duursma', 'Masae Ishihara', 'Diego R. Barneche', 'Rich G. FitzJohn', 'Angelica Vårhammar')
-	dataAuthors <- unique(do.call(c, lapply(unique(data$studyName), getPersonell)))
+authorDetails  <-  function(data) {
+	firstAuthors   <-  read.csv("../config/contact.csv", header = TRUE,
[…]

### 2014-09-18 ms/ms.Rmd
Bibliography rendering incorrectly

@@
 ---
-output: html_document
+output:
+  html_document
 ---
+
 # *Ecological Archives* Exxxxx
@@
 2.
-We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) `r citet("https://github.com/dfalster/dataMashR")` for merging data from different studies in a transparent and reproducible fashion.
+We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) for merging data from different studies in a transparent and reproducible fashion.
 While it was written with BAAD in mind, the package is designed to be easily  deployed for other compilations.
@@
 ## REFERENCES
+
 ```{r, echo=FALSE, results='asis'}
-bibliography(ordering = c("authors", "year", "title", "journal", "volume", "number", "pages", "url"))
+bibliography(sorting = 'ynt')
 ```

### 2014-09-21 ms/ms.Rmd
Standardise Ishihara's name across study. closes #163

@@
 The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.
-Moreover, we were invited to incorporate another compilation led by M.
+Moreover, we were invited to incorporate another compilation led by M.I.
 Ishihara and focusing on Japanese literature.
@@
 Vårhammar assisted with literature searching, data compilation, and coding.
-A separate compilation focusing on natural forests in Japan, was compiled by M.
+A separate compilation focusing on natural forests in Japan, was compiled by M.I.
 Ishihara, with assistance from T.
@@
 FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).
-M.
+M.I.
 Ishihara was supported by the Environmental Research and Technology Development Fund (S-9-3) of the Ministry of the Environment, Japan.

### 2014-09-22 ms/ms.Rmd
Fixed figure size in ms. Closes #166
Fixed taxize citation for #165

@@
 knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")
+options(citation_format = "text", max.names=2)
 
@@
 
-```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height8}
+```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
 
@@
 ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
-iii) All species names were matched against online databases with the R packages `Taxonstand` and `taxize` R package citet(c("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
+iii) All species names were matched against online databases with the R packages `Taxonstand` and `taxize` R package `r citep(list("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
 iv) Additional review by contributors using reports generated from the data.

### 2014-09-24 ms/ms.Rmd
minor glitch in acknowledgments
Kluge to avoid knitcitations bug. Closes #165

@@
 
+```{r precite, results='hide', message=FALSE, echo=FALSE}
+# Kluge to avoid all author names being cited. Bug in knitcitations, see:
+# https://github.com/cboettig/knitcitations/issues/65
+citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))
+citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))
+citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))
+citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))
+citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))
+citep("10.1038/nj7461-243a")
+```
+
 ## INTRODUCTION
@@
 Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))`.
+Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.
 
@@
 
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))`.
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.1016/j.tree.2011.11.016","10.12688/f1000research.4518"))`.
 This compilation would not have been possible without several new technologies.
@@
 We thank Takahiro Inoue, Hayato Aoyama and Shin-ichiro Aiba for valuable assistance.
-In addition, we thank the authors thank the following people for contributions to the following studies:
+In addition, the authors thank the following people for contributions to the following studies:
 

### 2014-09-26 ms/ms.Rmd
Incorporate feedback from Ishihara. Clsoes #174.

@@
 
-D.S.
+**Compilers:** D.S.
 Falster, A.
@@
 We thank Takahiro Inoue, Hayato Aoyama and Shin-ichiro Aiba for valuable assistance.
-In addition, the authors thank the following people for contributions to the following studies:
 
+In addition, we thank the following people for contributions to the following studies:
 ```{r,results="asis",echo=FALSE, message=FALSE}
-
 getAcknowledgments <- function(studyName) {
@@
   if(!is.na(i))
-    out <- sprintf("- %s: %s.\n", studyName, metadata$Description[i])
+    out <- sprintf(" **%s:** %s;", studyName, metadata$Description[i])
   out
@@
 ```
+.
 

### 2014-09-26 data/Mori1992/dataManipulate.R

@@ -1,12 +1,0 @@
-manipulate  <-  function(raw){
-
-# Adding branch mass to stem mass
-raw$m.st <- raw$m.st + raw$m.br
-
-	# Remove vine species
-	vines <- c("Trachelospermum asiaticum", "Akebia trifoliata", "Gardneria nutans", "Cocculus trilobus", "Wisteria floribunda", "Vitis saccharifera", "Hedera rhombea", "Ficus nipponica")
-
-	raw <- subset(raw, !(species %in% vines))
-
-	raw
-}

### 2014-10-03 ms/ms.Rmd
print author details to file
Shorten abstract to meet word limit

@@
 writeLines(paste0(details$address$code, ". ", details$address$address, collapse="\n"))
+
+write.csv(data.frame(name=details$authors, email=details$emails, address=details$address$address[match(details$address_code, details$address$code)]), "../output/authors.csv", row.names=FALSE)
 ```
@@
 **2.
-Abstract**: Understanding how different plants are plant constructed -- i.e.
-how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping the flow of energy in the terrestrial biosphere.
-Allocation patterns can differ widely through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
-While a variety of models dealing with carbon and biomass allocation exist, we still lack a synthetic understanding of the underlying processes.
+Abstract**: Understanding how plants are plant constructed -- i.e.
+how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere.
+Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
+While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes.
 This is partly due to the lack of suitable datasets for validating and parametrising models.
@@
 The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues.
-We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, status, light, experimental treatment, vegetation type) where available.
+We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, light, experimental treatment, vegetation type) where available.
 It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the world's vegetation.

### 2014-11-20 ms.Rmd
First version of manuscript support

--- 2014-10-03 ms/ms.Rmd
@@
 ```{r, results="hide", echo=FALSE, message=FALSE}
-source("MS-fun.R")
-
 knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")
@@
 bib <- read.bibtex("refs.bib")
-baad <- readRDS("../output/baad.rds")
 data <- baad$data
-var.def <- read.csv("../config/variableDefinitions.csv",
-                    stringsAsFactors=FALSE)
-
-vars <- setdiff(var.def$variable[var.def$type == "numeric"], c("map","mat","lai","age"))
+var_def <- baad$dictionary
+vars <- setdiff(var_def$variable[var_def$type == "numeric"], 
+                c("map", "mat", "lai", "age"))
 ```
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-details <- authorDetails(baad$contact)
-
-writeLines(paste("**List of Authors:** ", paste0(details$authors, "^", details$address_code,"^",collapse = ", ") ))
-
+writeLines(paste("**List of Authors:** ", 
+                 paste0(author_info$authors$name, 
+                        "^", author_info$authors$address_code, "^", 
+                        collapse = ", ")))
 writeLines("\n\n")
-
-writeLines(paste0(details$address$code, ". ", details$address$address, collapse="\n"))
-
-write.csv(data.frame(name=details$authors, email=details$emails, address=details$address$address[match(details$address_code, details$address$code)]), "../output/authors.csv", row.names=FALSE)
+writeLines(paste0(author_info$address_table$code, ". ", 
+                  author_info$address_table$address))
 ```
@@
 ```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
-
 # World plot colored by vegetation type
-cols <- make.transparent(c(
+cols <- make_transparent(c(
     "blue","brown","forestgreen","darkorange","dodgerblue2",
@@
 legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')
-
 ```
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-library(plyr)
-library(whisker)
-
-## for debugging
-# for(s in unique(data$studyName)){
-#  cat(s)
-#  studyDetails(data[data$studyName==s,])}
-
-details2B  <-  daply(data, .(studyName), function(x) studyDetails(x))
-writeLines(details2B,sep="\n\n")
+details2B <- sapply(split(data, data$studyName), study_details, baad)
+writeLines(details2B, sep="\n\n")
 ```
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
-x <-daply(var.def, 1, function(x)
-		paste0("* ", x$variable,
-    "\n\t- Type: ", x$type,
-		"\n\t- Label: ", x$label,
-		"\n\t- Description: ", x$description,
-		"\n\t- Units: ", gsub(".","/", x$unitsMD, fixed=TRUE)
-		))
-
-writeLines(x[var.def$variable])
+writeLines(variable_details(var_def))
 ```
@@
 options(digits=9)
-kable(summaryTable(data, var.def), "markdown", align="l", padding=10)
+kable(summary_table(data, var_def), "markdown", align="l", padding=10)
 ```
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-getAcknowledgments <- function(studyName) {
-  metadata  <-  readStudyFile(studyName, "studyMetadata.csv")
-  i <- match("Acknowledgements",metadata$Topic)
-  out <- ""
-  if(!is.na(i))
-    out <- sprintf(" **%s:** %s;", studyName, metadata$Description[i])
-  out
-}
-writeLines(paste(sapply(unique(data$studyName), getAcknowledgments), collapse= ""))
+ack <- sapply(unique(data$studyName), get_acknowledgements)
+writeLines(paste(ack, collapse=""))
 ```

### 2014-11-20 functions_ms.R

@@ -1,0 +1,267 @@
+## Author information is handled as a proper maker target:
+author_details <- function(filename_first_authors, baad) {
+  first_authors <- read.csv(filename_first_authors, na.strings=c(NA, ''),
+                          stringsAsFactors=FALSE, strip.white=TRUE)
+  data_authors <- baad$contact
+  data_authors <- data_authors[order(last_name(data_authors$name),
+                                     data_authors$name),]
+  data_authors$email[data_authors$email == ""] <- NA
+
+  cols <- c("name", "email", "address")
+  all_authors <- rbind(first_authors[cols], data_authors[,cols])
+  all_authors <- all_authors[!duplicated(all_authors$name),]
+
+  address <- unique(all_authors$address)
+  address_table <- data.frame(code=seq_along(address), address=address)
+
+  all_authors$address_code <- match(all_authors$address, address)
+  
+  list(authors=all_authors,
+       address_table=address_table)
+}
+
+last_name <- function(author) {
+  sapply(author, function(x)
+         last(strsplit(x, ' ', useBytes=TRUE)[[1]]))
+}
+
+save_author_details <- function(d, filename) {
+  write.csv(d$authors[c("name", "email", "address")], filename,
+            row.names=FALSE)
+}
+
+## Below here still needs work
+
+## TODO: This needs adding to the main baad object.
+get_metadata_methods  <-  function(metadata) {
+  if (nrow(metadata) > 0) {
+    metadata <- metadata[!is.na(metadata$Description) &
+                         !(tolower(metadata$Topic) %in% c('sampling strategy', 'age', 'growth environment', 'acknowledgements')),]
+    paste0(metadata$Topic, ': ', metadata$Description, collapse='\n\t- ')
+  }
+}
+
+get_metadata_design  <-  function(metadata) {
+  metadata$Description[metadata$Topic == 'Sampling strategy']
+}
+
+drawWorldPlot <- function(data, sizebyn=FALSE, add=FALSE,
+                          pchcol="red", legend=TRUE) {
+  if (!add){
+    map('world',col="grey80",bg="white",lwd=0.5,fill=TRUE,resolution=0,wrap=TRUE, border="grey80", ylim=c(-80, 80), mar=rep(0,4))
[…]

### 2014-11-24 ms/ms.Rmd

@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-getAcknowledgments <- function(studyName) {
-  metadata  <-  readStudyFile(studyName, "studyMetadata.csv")
-  i <- match("Acknowledgements",metadata$Topic)
-  out <- ""
-  if(!is.na(i))
-    out <- sprintf(" **%s:** %s;", studyName, metadata$Description[i])
-  out
-}
-writeLines(paste(sapply(unique(data$studyName), getAcknowledgments), collapse= ""))
+writeLines(get_acknowledgements(baad))
 ```

### 2014-11-24 reports/report.Rmd

@@
-```{r, echo=FALSE}
-source('report-fun.R')
+```{r, echo=FALSE, results="hide"}
+study_name <- data_study$key
 ```
 
-# Report for study: `r .study`
+# Report for study: `r study_name`
 
@@
 
-**Data contributor:** `r .dat$contacts$name`
+**Data contributor:** `r data_study$contacts$name`
 
-**Email:** `r paste0(.dat$contacts$email[.dat$contacts$email!=""], collapse = ", ")`
+**Email:** `r paste0(data_study$contacts$email[data_study$contacts$email!=""], collapse = ", ")`
 
@@
 
-`r paste0("- ", .dat$contacts$address, collapse = "\n")`
+`r paste0("- ", data_study$contacts$address, collapse = "\n")`
 
@@
 
-**Citation:** `r .dat$references$citation`
+**Citation:** `r data_study$references$citation`
 
-**DOI:** `r  paste0('[', bibGetElement(.dat$bibtex, element = "doi"), '](http://doi.org/', bibGetElement(.dat$bibtex, element = "doi"),')')`
+**DOI:** `r md_link_doi(bib_element(data_study$bibtex, "doi"))`
 
-**Abstract:** `r  bibGetElement(.dat$bibtex, element = "abstract")`
+**Abstract:** `r  bib_element(data_study$bibtex, "abstract")`
 
@@
 
-The dataset includes records for `r length(.dat$data$studyName)` individuals from `r length(unique(.dat$data$species))` species belonging to `r length(unique(.dat$data$family))` family(ies), presenting `r length(unique(.dat$data$pft))` functional type(s), growing in `r length(unique(.dat$data$growingCondition))` condition(s) within `r length(unique(.dat$data$vegetation))` major type(s) of habitat, with data included for the following variables:
+The dataset includes records for `r nrow(data_study$data)` individuals from `r length(unique(data_study$data$species))` species belonging to `r length(unique(data_study$data$family))` family(ies), presenting `r length(unique(data_study$data$pft))` functional type(s), growing in `r length(unique(data_study$data$growingCondition))` condition(s) within `r length(unique(data_study$data$vegetation))` major type(s) of habitat, with data included for the following variables:
 
-```{r results='asis', echo=FALSE, warning=FALSE}
-options(digits=9)
-kable(
-	summaryTable(.dat$data, var.def= read.csv("../config/variableDefinitions.csv", stringsAsFactors=FALSE)),
-	 "markdown", align="l", padding=10)
+```{r summary_table, results='asis', echo=FALSE}
+# From the ms functions:
+tab <- summary_table(data_study$data, baad$dictionary)
+kable(tab[names(tab) != "Studies"], "markdown", align="l", padding=10)
 ```
 
-## Locations sampled
-
-Data were sampled at the following locations:
-
-```{r echo=FALSE, fig.width=12, fig.height=10, message=FALSE, warning=FALSE}
-prepdata  <-  prepMapInfo(alldata$data, .study)
-drawWorldPlot (prepdata)
+```{r world_map, echo=FALSE, fig.width=12, fig.height=10}
+map_info <- prep_map_info(data_study$data, study_name)
+drawWorldPlot(map_info)
 ```
@@
 
-```{r echo=FALSE, fig.width=12, fig.height=10, message=FALSE, warning=FALSE}
-repMissingInfo(prepdata)
-drawCountryPlot(prepdata)
+```{r country_map, echo=FALSE, fig.width=12, fig.height=10}
+report_missing_info(map_info)
+drawCountryPlot(map_info)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment='',results='asis'}
-kable(locLevelInfo(.dat$data),"markdown", align="l", padding=2)
+```{r location_level_info, echo=FALSE, comment='', results='asis'}
+kable(location_level_info(data_study$data), "markdown", align="l", padding=2)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
-kable(standLevelInfo(.dat$data),"markdown", align="l", padding=2)
+```{r stand_level_info, echo=FALSE, comment='', results="asis"}
+kable(stand_level_info(data_study$data), "markdown", align="l", padding=2)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment='',results="asis"}
-kable(spLevelInfo(.dat$data), "markdown", align="l", padding=2)
+```{r species_level_info, echo=FALSE, comment='', results="asis"}
+kable(species_level_info(data_study$data), "markdown", align="l", padding=2)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
-metaTable  <-  getMeta(.study)
-writeLines(paste0("**",metaTable[,1],"**", ": ", metaTable[,2], "\n"))
+```{r metadata, echo=FALSE, comment='', results="asis"}
+metadata <- data_study$metadata
+writeLines(sprintf("**%s**: %s\n", metadata$Topic, metadata$Description))
 ```
@@
 
-This is how the study `r .study` fits in the entire dataset (grey). each colour represents a species. A legend of species names with colours is included at the end for reports with 1 < n < 20 species.
+This is how the study `r study_name` fits in the entire dataset (grey).
+each colour represents a species.
+A legend of species names with colours is included at the end for reports with 1 < n < 20 species.
 
-```{r fig.height=12, fig.width=12, echo=FALSE, warning=FALSE, message=FALSE}
-makePlotPanel(alldata, .dat)
+```{r variable_plots, fig.height=12, fig.width=12, echo=FALSE}
+report_variable_plots(data_study, baad)
 ```

### 2014-11-25 reports/report.Rmd
Split apart baad to avoid report rebuilds
Depends only on the baad data, and less on other studies metadata.

@@
 # From the ms functions:
-tab <- summary_table(data_study$data, baad$dictionary)
+tab <- summary_table(data_study$data, baad_dictionary)
 kable(tab[names(tab) != "Studies"], "markdown", align="l", padding=10)
@@
 ```{r variable_plots, fig.height=12, fig.width=12, echo=FALSE}
-report_variable_plots(data_study, baad)
+report_variable_plots(data_study, baad_data, baad_dictionary)
 ```

### 2014-11-25 R/reports.R

@@ -184,18 +184,15 @@ species_level_info <- function(data) {
 
 ######################################################################
 
-report_variable_plots <- function(data_study, baad) {
-  dict <- baad$dictionary
-  dict$units <- gsub(".", "/", dict$units, fixed=TRUE)
-  
-  labels <- sprintf("%s (%s)", dict$label, dict$units)
-  names(labels) <- dict$variable
+report_variable_plots <- function(data_study, data, dictionary) {
+  labels <- sprintf("%s (%s)", dictionary$label, dictionary$units)
+  names(labels) <- dictionary$variable
 
   ## determine variables to plot
-  var_present <- colSums(!is.na(data_study$data[dict$variable])) > 0
-  plot_vars <- dict$variable[dict$group == "tree" &
-                             dict$type == "numeric" &
-                             var_present]
+  var_present <- colSums(!is.na(data_study$data[dictionary$variable])) > 0
+  plot_vars <- dictionary$variable[dictionary$group == "tree" &
+                                   dictionary$type == "numeric" &
+                                   var_present]
 
   ## set up a vector of colours, each species with different color
   species <- as.factor(data_study$data$species)
@@ -207,7 +204,7 @@ report_variable_plots <- function(data_study, baad) {
     v1 <- plot_vars[[i]]
     for (v2 in plot_vars[seq_along(plot_vars) > i]) {
       report_variable_plot1(data_study$data[c(v1, v2)],
-                            baad$data[c(v1, v2)], studycol=cols,
+                            data[c(v1, v2)], studycol=cols,
                             xlab=labels[[v1]], ylab=labels[[v2]])
     }
   }
@@ -288,4 +285,13 @@ axis_log10 <- function(side=1, horiz=FALSE, labels=TRUE,
     axis(side, at=10^at, FALSE, col=if (horiz) fg else NA,
          col.ticks=fg, las=las)
   }
+}
+
+## Dancing around the need to rebuild everything if we update methods
+## only, not generally useful.
+extract_baad_dictionary <- function(baad) {
+  baad$dictionary
+}
+extract_baad_data <- function(baad) {
+  baad$data
 }

### 2014-12-02 reports/report.Rmd
Simplify country plotting code. #191

@@
 ```{r world_map, echo=FALSE, fig.width=12, fig.height=10}
-map_info <- prep_map_info(data_study$data, study_name)
-drawWorldPlot(map_info)
+drawWorldPlot(data_study$data)
 ```
@@
 ```{r country_map, echo=FALSE, fig.width=12, fig.height=10}
-report_missing_info(map_info)
-drawCountryPlot(map_info)
+map_sites_all_countries(data_study$data)
 ```

### 2014-12-02 R/reports.R

@@ -17,128 +17,40 @@ md_link_doi <- function(doi) {
   md_link(doi, paste0("http://doi.org/", doi))
 }
 
-prep_map_info <- function(data, study_name=NA) {
+map_sites_all_countries <- function(data) {
+
   cols <- c("latitude", "longitude", "location")
-  is_unique <- !duplicated(data[c("latitude", "longitude", "location")])
-  dsub <- data[is_unique, cols]
-
-  i <- !is.na(dsub$latitude) | !is.na(dsub$longitude)
-  if (any(i)) {
-    dsub$country <- NA_character_
-    dsub$country[i] <- map.where(x=as.numeric(dsub$longitude[i]),
-                                 y=as.numeric(dsub$latitude[i]))
-  }
-
-  dsub
+  data <- data[!duplicated(data[,cols]), cols]
+  data$country <- map.where(x=as.numeric(data$longitude),
+                                 y=as.numeric(data$latitude))
+  d_ply(data, "country", map_sites_in_country)
 }
 
-## TODO: Can someone look into this and see if this is still useful?
-## TODO: Why is "NA" a possibility?  I thought that NA strings would
-## have dealt with that?
-## TODO: 
-report_missing_info <- function(data) {
-  j <- (!is.na(data$latitude) & !is.na(data$longitude) &
-        data$location != "NA" | is.na(data$location))
-  sj <- data[!j, ]
+map_sites_in_country <- function(data){
 
-  # location Info
-  k <- !is.na(sj$location) & is.na(sj$longitude)
-  if (length(k[k == TRUE]) > 0) {
-    cat("Please notice that there are no coordinates for the following location(s):")
-    return(cbind(sj$location[k]))
-  }
-  # coordinate Info
-  k <- !is.na(sj$longitude) & is.na(sj$location) | sj$location == "NA"
-  if (any(k)) {
-    cat("Please notice that there is no location information for the following coordinate(s):")
-    return(cbind(lon=sj$lon[k], lat=sj$lat[k]))
-  }
-  # country Info
-  k <- !is.na(sj$location) & is.na(sj$country)
-  if (any(k)) {
-    cat("Please notice that there is no country information for the following location(s):")
-    return(cbind(sj$location[k]))
[…]

### 2014-12-10 ms/ms.Rmd
Revisions to ms; added brief description of post-processing; added Figure2; minor changes and typo fixes.

@@
 When a single data contribution has been modified, the final master dataset is thus easily recreated.
-2.
-We wrote a new `R` package [dataMashR](https://github.com/dfalster/dataMashR) for merging data from different studies in a transparent and reproducible fashion.
-While it was written with BAAD in mind, the package is designed to be easily  deployed for other compilations.
+2. We minimised the amount of code by requiring each study to conform to a common format (Figure 2).
 3.
+We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`.
+These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing.
+Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata.
+The updated reports are available on the database website.
+4.
 We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data.
@@
 With version control, we are able to trace these errors to the source.
-4.
-We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`.
-These reports were sent to data contributors for verification, and allowed easy identification of errors in the raw data or processing.
 
-Until recently, scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`.
+Until recently, most scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`.
 Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
@@
 **Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.
+
+![baadworkflow](Figure2.png)
+**Figure2** Workflow of BAAD. Each of the raw data files was processed in the same way, resulting in a single dataset ('BAAD') with a common format.
 
@@
 While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes.
-This is partly due to the lack of suitable datasets for validating and parametrising models.
+This is partly due to the lack of suitable datasets for validating and parameterising models.
 To that end, we present the Biomass And Allometry Database (BAAD) for woody plants.
@@
 Thus BAAD contains individual level data from different studies, transformed into standard units and variable names.
-The transformations were achieved using a new R package dataMashR.
+The transformations were achieved using a common workflow for all raw data files.
 Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
@@
 
-**Latest update**: xxxx for all data files.
+**Latest update**: The dataset made available through this publication was processed on 2014-12-10.
+Newer updates may be retrieved from the BAAD website ([https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
 
-**Metadata status**: Metadata are complete to last update and are stored
-with data.
+**Metadata status**: Metadata are complete to last update and are stored with data.
 
-**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was approppriately matched to variables in BAAD, and to minimize outliers.
+**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in BAAD, and to minimize outliers.
 
@@
 Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.
+v) Derived variables were calculated from the raw data where appropriate.
+For example, a dataset may contain crown length and plant height but not height to crown base, in which case the latter was calculated from the other two variables.
 
@@
 Storage location and medium**: The data published here is the first release from the Biomass and allometry database stored at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
-The version of the data made available for this publication is available here XXXXX.
+The version of the data made available for this publication is available here ((LINK TO ECOLOGICAL ARCHIVES DATA FILE)).
 For scientific analyses, we suggest you use this version.
@@
 
-If you notice any problem with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).
+If you notice any problems with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).
 
@@
 For queries about the entire dataset, please contact either Daniel Falster or Remko Duursma.
-Contact details for all authors are available in the attached table.
 For problems with the dataset, please submit an issue as described above.

### 2014-12-10 ms/ms.Rmd
Expanded description on search for data in ms Intro

@@
 FitzJohn and A.
-Vårhammar) set out to create such a resource, by asking authors directly whether they would be willing to make their raw data files freely available.
+Vårhammar) set out to create such a resource.
+We identified data sources from existing collaborations as well as a wide-ranging literature search, focussed on woody plants in field as well as controlled conditions.
+We constrained the search by including only data where a wide range in plant sizes was measured, as well as a set of minimum variables (plant height and leaf mass or area), and an emphasis on field-grown plants (nearly all data from plants grown in controlled conditions arose from existing collaborations).
+We do not claim to have done an exhaustive search of the literature, due to the very large number of potential studies that could be included.
+In each case we asked authors directly whether they would be willing to make their raw data files freely available.
 The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.

### 2014-12-10 ms/ms.Rmd
Update landing page. Closes #68

@@
 
-BAAD contains the following files:
+BAAD contains the following elements:
 
-- `dictionary.csv`: a table of variable definitions
-- `data.csv`: amalgamated dataset, with columns as defined in `dictionary.csv`
-- `references.bib`: bibtex entries containing the primary source for each study
-- `references.csv`: summary of bibtex file, giving DOI, URL and citation of  the primary source for each study
-- `contacts.csv`: list of contacts and affiliations for each study
+- `data`: amalgamated dataset, with columns as defined in `dictionary`
+- `dictionary`: a table of variable definitions
+- `metadata`: a table with columns "studyName","Topic","Description", containing written information about the methods used to collect the data
+- `methods`: a table with columns as in data, but containing a code for the methods used to collect the data. See [config/methodsDefinitions.csv](config/methodsDefinitions.csv) for codes.
+- `references`:  as both summary table (giving DOI, URL, citation) and bibtex entries containing the primary source for each study
+- `contacts`: table with contacts and affiliations for each study.
+
+These elements are available as both
+
+1) a list within the file `baad.rds` (for use in R)
+2) as a series of csv and text files.
 

### 2014-12-11 ms/ms.Rmd
Minor tweaks to user information, and manuscript

@@
 Ishihara and focusing on Japanese literature.
-As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies.
-
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.1016/j.tree.2011.11.016","10.12688/f1000research.4518"))`.
-This compilation would not have been possible without several new technologies.
-In particular,
-
-1.
-Our entire work flow is scripted, using the R computing environment.
-The database can be rebuilt in an instant from the raw data files, using a single command.
-When a single data contribution has been modified, the final master dataset is thus easily recreated.
-2. We minimised the amount of code by requiring each study to conform to a common format (Figure 2).
-3.
-We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`.
-These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing.
-Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata.
-The updated reports are available on the database website.
-4.
-We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data.
-Datasets as well as processing methods inevitably contain errors.
-With version control, we are able to trace these errors to the source.
-
-Until recently, most scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`.
-Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
-It is exciting to be able to put these parcels together, to create something larger.
-We hope others will use this dataset and build on it.
-See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
+As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies (Fig.
+1).
 
@@
 
-**Figure 1** Global distribution of studies included in BAAD. Symbols are sized relative to the number of individual plants in each dataset.
+**Figure 1** Global distribution of studies included in the BAAD. Symbols are sized relative to the number of individual plants in each dataset.
+
+Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.1016/j.tree.2011.11.016","10.12688/f1000research.4518"))`.
+This compilation would not have been possible without several new technologies.
+In particular,
+
+1.
+Our entire work flow is scripted, using the R computing environment.
+The database can be rebuilt in an instant from the raw data files, using a single command.
+When a single data contribution has been modified, the final master dataset is thus easily recreated.
+2.
+We minimised the amount of code by requiring each study to conform to a common format, with separate files for raw data, units of data, meta data, contact details, and citation (see examples in attached code).
+These files were then processed in a standardised way (Fig.
+2).
+3.
+We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`.
+These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing.
+Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata.
+The updated reports are available on the database website.
+4.
+We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data.
+Datasets as well as processing methods inevitably contain errors.
+With version control, we are able to trace these errors to the source, and review review the history of changes to any given dataset.
 
 ![baadworkflow](Figure2.png)
-**Figure2** Workflow of BAAD. Each of the raw data files was processed in the same way, resulting in a single dataset ('BAAD') with a common format.
+**Figure2** Workflow for building the BAAD. Data from each study is processed in the same way, using a standardised set of input files, resulting in a single dataset with a common format.
+
+Until recently, most scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`.
+Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
+It is exciting to be able to put these parcels together, to create something larger.
+We hope others will use this dataset and build on it.
+See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
 
@@
 ### A. DATA SET IDENTITY
-BAAD: a Biomass And Allometry Database for woody plants
+the BAAD: a Biomass And Allometry Database for woody plants
 
@@
 To that end, we present the Biomass And Allometry Database (BAAD) for woody plants.
-BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
+The BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
 Most of these data are associated with an existing publication, however, raw data files were rarely made public at time of publication.
-Thus BAAD contains individual level data from different studies, transformed into standard units and variable names.
+Thus the BAAD contains individual level data from different studies, transformed into standard units and variable names.
 The transformations were achieved using a common workflow for all raw data files.
-Other features which distinguish BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
+Other features which distinguish the BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e.
 not indirectly via allometric equations (except in very large trees where biomass was estimated from detailed sub-sampling).
@@
 
-**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in BAAD, and to minimize outliers.
+**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in the BAAD, and to minimize outliers.
 
@@
 
-BAAD contains the following elements:
+The BAAD contains the following elements:
 

### 2014-12-15 ms/ms.Rmd
Minor typos, fixes

@@
 4.
-We used version control and the code sharing website [github](www.github.com) to track the evolution of the database and the code to process the raw data.
+We used version control and the code sharing website [github](https://github.com/dfalster/baad/) to track the evolution of the database and the code to process the raw data.
 Datasets as well as processing methods inevitably contain errors.
@@
 
-**Latest update**: The dataset made available through this publication was processed on 2014-12-10.
+**Latest update**: The dataset made available through this publication was processed on 2014-12-15.
 Newer updates may be retrieved from the BAAD website ([https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).

### 2014-12-15 R/ms.R

@@ -137,7 +137,7 @@ study_details <- function(data, baad) {
     personell    = personell,
     citation     = get_citation_ms(study_name, baad),
     vars         = paste0(vars, collapse=", "),
-    species      = paste0(unique(data$species), collapse=", "))
+    species      = paste0(sort(unique(data$species)), collapse=", "))
 
   whisker.render(template, dat)
 }

### 2014-12-23 ms/ms.Rmd
Citation to Ishihara's accepted paper

@@
 citep("10.1038/nj7461-243a")
+citep(list(bib["Ishihara_Efficacy_2015"]))
 ```
@@
 Moreover, we were invited to incorporate another compilation led by M.I.
-Ishihara and focusing on Japanese literature.
+Ishihara and focusing on Japanese literature `r citep(list(bib["Ishihara_Efficacy_2015"]))`.
 As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies (Fig.

### 2014-12-23 ms/ms.Rmd
Reduce word count in abstract
Max of 350 for Ecology.

@@
 how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere.
-Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to or grown in different environments.
+Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to different environments.
 While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes.
@@
 The BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species.
-Most of these data are associated with an existing publication, however, raw data files were rarely made public at time of publication.
+Most of these data come from an existing publication, however, raw data were rarely made public at time of publication.
 Thus the BAAD contains individual level data from different studies, transformed into standard units and variable names.

### 2015-02-05 ms/ms.Rmd
Change link
Fix links, reword ambiguous text, fix license

@@
 We hope others will use this dataset and build on it.
-See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
+See [github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
 
@@
 **Latest update**: The dataset made available through this publication was processed on 2014-12-15.
-Newer updates may be retrieved from the BAAD website ([https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
+Newer updates may be retrieved from the BAAD website [github.com/dfalster/baad](https://github.com/dfalster/baad).
 
@@
 **1.
-Storage location and medium**: The data published here is the first release from the Biomass and allometry database stored at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
-The version of the data made available for this publication is available here ((LINK TO ECOLOGICAL ARCHIVES DATA FILE)).
-For scientific analyses, we suggest you use this version.
+Storage location and medium**: The data published on Ecological Archives is the first release from the Biomass and allometry database (v1.0.0) stored at [github.com/dfalster/baad](https://github.com/dfalster/baad/).
+The version of the data contained in this publication can also be downloaded from the releases page on gihtub: [github.com/dfalster/baad/releases](https://github.com/dfalster/baad/releases).
+Development versions of the dataset will continue to be made available at [github.com/dfalster/baad](https://github.com/dfalster/baad).
 
-Development versions of the dataset will continue to be available at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
-
-If you notice any problems with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).
+If you notice any problems with the dataset, please list an issue at [github.com/dfalster/baad/issues/new](https://github.com/dfalster/baad/issues/new).
 
@@
 
-**Copyright restrictions**: The dataset is released under the [Creative Commons CC0 license](http://creativecommons.org/choose/zero/?lang=en).
-As such, it may be freely used and redistributed.
+**Copyright restrictions**: The dataset is released under the the [Creative Commons Zero](https://creativecommons.org/publicdomain/zero/1.0/) public domain dedication.
 When using the dataset, we kindly request that you cite this article, recognizing the hard work that went into collecting the data and the authors' willingness to make it publicly available.

### 2015-02-10 ms/ms.Rmd
Ecology requires separate docs for authors and metadata, and in word

--- 2014-12-23 ms/ms.Rmd
@@
----
-output:
-  html_document
----
-
-# *Ecological Archives* Exxxxx
-# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants. *Ecology* xxxx
+# BAAD: a Biomass And Allometry Database for woody plants
 
@@
 
-bib <- read.bibtex("refs.bib")
 data <- baad$data
@@
 ```{r,results="asis",echo=FALSE, message=FALSE}
-writeLines(paste("**List of Authors:** ",
-                 paste0(author_info$authors$name,
+writeLines(paste0(author_info$authors$name,
                         "^", author_info$authors$address_code, "^",
-                        collapse = ", ")))
+                        collapse = ", "))
 writeLines("\n\n")
-writeLines(paste0(author_info$address_table$code, ". ",
-                  author_info$address_table$address))
+writeLines(paste0("^", author_info$authors$address_code, "^ ",
+                  author_info$address_table$address,"  "))
 ```
 
-```{r precite, results='hide', message=FALSE, echo=FALSE}
-# Kluge to avoid all author names being cited. Bug in knitcitations, see:
-# https://github.com/cboettig/knitcitations/issues/65
-citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))
-citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))
-citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))
-citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))
-citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))
-citep("10.1038/nj7461-243a")
-citep(list(bib["Ishihara_Efficacy_2015"]))
-```
-
-## INTRODUCTION
-
-Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`.
-The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`.
-While many scientific studies have destructively harvested woody plants, most of the data collected have only been made available in the form of summary tables or figures included in publications.
-Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world.
-Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale.
-Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.
-
-We (D.S.
-Falster and R.A.
-Duursma, with the help of D.R.
-Barneche, R.G.
-FitzJohn and A.
-Vårhammar) set out to create such a resource.
-We identified data sources from existing collaborations as well as a wide-ranging literature search, focussed on woody plants in field as well as controlled conditions.
-We constrained the search by including only data where a wide range in plant sizes was measured, as well as a set of minimum variables (plant height and leaf mass or area), and an emphasis on field-grown plants (nearly all data from plants grown in controlled conditions arose from existing collaborations).
-We do not claim to have done an exhaustive search of the literature, due to the very large number of potential studies that could be included.
-In each case we asked authors directly whether they would be willing to make their raw data files freely available.
-The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data.
-Moreover, we were invited to incorporate another compilation led by M.I.
-Ishihara and focusing on Japanese literature `r citep(list(bib["Ishihara_Efficacy_2015"]))`.
-As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies (Fig.
-1).
-
-```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
-# World plot colored by vegetation type
-cols <- make_transparent(c(
-    "blue","brown","forestgreen","darkorange","dodgerblue2",
-    "navajowhite4","lightgoldenrod3","lavender"),0.67)
-vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
-vegs_labels <- c("Boreal forest","Savannah","Temperate forest",
-                 "Temperate rainforest","Tropical rainforest",
-                 "Tropical seasonal forest","Woodland","Glasshouse / Common garden")
-
-drawWorldPlot(data[data$vegetation == vegs[1],], sizebyn=TRUE, pchcol=cols[1])
-for(i in 2:length(vegs)){
-  drawWorldPlot(data[data$vegetation == vegs[i],], sizebyn=TRUE, pchcol=cols[i], add=TRUE)
-}
-drawWorldPlot(data[data$growingCondition == "GH",], sizebyn=TRUE,
-              pchcol=cols[8], add=TRUE)
-par(xpd=NA)
-legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
-legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')
-```
-
-**Figure 1** Global distribution of studies included in the BAAD. Symbols are sized relative to the number of individual plants in each dataset.
-
-Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.1016/j.tree.2011.11.016","10.12688/f1000research.4518"))`.
-This compilation would not have been possible without several new technologies.
-In particular,
-
-1.
-Our entire work flow is scripted, using the R computing environment.
-The database can be rebuilt in an instant from the raw data files, using a single command.
-When a single data contribution has been modified, the final master dataset is thus easily recreated.
-2.
-We minimised the amount of code by requiring each study to conform to a common format, with separate files for raw data, units of data, meta data, contact details, and citation (see examples in attached code).
-These files were then processed in a standardised way (Fig.
-2).
-3.
-We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`.
-These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing.
-Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata.
-The updated reports are available on the database website.
-4.
-We used version control and the code sharing website [github](https://github.com/dfalster/baad/) to track the evolution of the database and the code to process the raw data.
-Datasets as well as processing methods inevitably contain errors.
-With version control, we are able to trace these errors to the source, and review review the history of changes to any given dataset.
-
-![baadworkflow](Figure2.png)
-**Figure2** Workflow for building the BAAD. Data from each study is processed in the same way, using a standardised set of input files, resulting in a single dataset with a common format.
-
-Until recently, most scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`.
-Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies.
-It is exciting to be able to put these parcels together, to create something larger.
-We hope others will use this dataset and build on it.
-See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.
-
-## METHODS
-
-## METADATA CLASS I. DATA SET DESCRIPTORS
-
-### A. DATA SET IDENTITY
-the BAAD: a Biomass And Allometry Database for woody plants
-
-### B. DATA SET IDENTIFICATION CODE:
-
-### C. DATA SET DESCRIPTORS:
-**1.
-Originators**: The study was initiated by D.S.
-Falster and R.A.
-Duursma.
-D.R.
-Barneche, R.G.
-FitzJohn and A.
-Vårhammar assisted with literature searching, data compilation, and coding.
-A separate compilation focusing on natural forests in Japan, was compiled by M.I.
-Ishihara, with assistance from T.
-Hiura, H.
-Utsugi and H.
-Tanouchi.
-All other authors contributed data to the project.
-
-**2.
-Abstract**: Understanding how plants are plant constructed -- i.e.
+Abstract.
+Understanding how plants are plant constructed -- i.e.
 how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere.
@@
 
-### D. Key words
-
-plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle
-
-## METADATA CLASS II. RESEARCH ORIGIN DESCRIPTORS
-
-### B. SPECIFIC SUBPROJECT DESCRIPTION
-
-```{r,results="asis",echo=FALSE, message=FALSE}
-details2B <- sapply(split(data, data$studyName), study_details, baad)
-writeLines(details2B, sep="\n\n")
-```
-
-## CLASS III. DATA SET STATUS AND ACCESSIBILITY
-
-### A. Status
-
-**Latest update**: The dataset made available through this publication was processed on 2014-12-15.
-Newer updates may be retrieved from the BAAD website ([https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
-
-**Metadata status**: Metadata are complete to last update and are stored with data.
-
-**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in the BAAD, and to minimize outliers.
-
-i) Initial analysis by data contributors. The majority of studies have been published and as such have gone through quality assurance by the data contributor.
-ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
-iii) All species names were matched against online databases with the R packages `Taxonstand` and `taxize` R package `r citep(list("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
-iv) Additional review by contributors using reports generated from the data.
-In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, metadata provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database).
-Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.
-v) Derived variables were calculated from the raw data where appropriate.
-For example, a dataset may contain crown length and plant height but not height to crown base, in which case the latter was calculated from the other two variables.
-
-### B. Accessibility
-
-**1.
-Storage location and medium**: The data published here is the first release from the Biomass and allometry database stored at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
-The version of the data made available for this publication is available here ((LINK TO ECOLOGICAL ARCHIVES DATA FILE)).
-For scientific analyses, we suggest you use this version.
-
-Development versions of the dataset will continue to be available at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).
-
-If you notice any problems with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).
-
-**Contact persons**: Queries about individual specific data points can be directed to the contributing author for that study.
-For queries about the entire dataset, please contact either Daniel Falster or Remko Duursma.
-For problems with the dataset, please submit an issue as described above.
-
-**Copyright restrictions**: The dataset is released under the [Creative Commons CC0 license](http://creativecommons.org/choose/zero/?lang=en).
-As such, it may be freely used and redistributed.
-When using the dataset, we kindly request that you cite this article, recognizing the hard work that went into collecting the data and the authors' willingness to make it publicly available.
-
-**Costs**: None.
-
-## CLASS IV. DATA STRUCTURAL DESCRIPTORS
-
-### A. Data Set Files
-
-The BAAD contains the following elements:
-
-- `data`: amalgamated dataset, with columns as defined in `dictionary`
-- `dictionary`: a table of variable definitions
-- `metadata`: a table with columns "studyName","Topic","Description", containing written information about the methods used to collect the data
-- `methods`: a table with columns as in data, but containing a code for the methods used to collect the data. See [config/methodsDefinitions.csv](config/methodsDefinitions.csv) for codes.
-- `references`:  as both summary table (giving DOI, URL, citation) and bibtex entries containing the primary source for each study
-- `contacts`: table with contacts and affiliations for each study.
-
-These elements are available as both
-
-1) a list within the file `baad.rds` (for use in R)
-2) as a series of csv and text files.
-
-### B. Variable definitions for data.csv
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
-writeLines(variable_details(var_def))
-```
-
-**Number of observations by variable**
-
-```{r results='asis', echo=FALSE, warning=FALSE}
-options(digits=9)
-kable(summary_table(data, var_def), "markdown", align="l", padding=10)
-```
-
-## ACKNOWLEDGMENTS
-
-**Compilers:** D.S.
-Falster, A.
-Vårhammer and D.R.
-Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A.
-Duursma.
-R.G.
-FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174).
-M.I.
-Ishihara was supported by the Environmental Research and Technology Development Fund (S-9-3) of the Ministry of the Environment, Japan.
-We thank Takahiro Inoue, Hayato Aoyama and Shin-ichiro Aiba for valuable assistance.
-
-In addition, we thank the following people for contributions to the following studies:
-```{r,results="asis",echo=FALSE, message=FALSE}
-writeLines(get_acknowledgements(baad))
-```
-.
-
-## REFERENCES
-
-```{r, echo=FALSE, results='asis'}
-bibliography(sorting = 'ynt')
-```
+Key words: plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle

### 2015-02-12 ms/ms.Rmd
Fix mangled address codes introduced in previous commit
Introduced in 916247c

@@
 writeLines("\n\n")
-writeLines(paste0("^", author_info$authors$address_code, "^ ",
+writeLines(paste0("^", author_info$address_table$code, "^ ",
                   author_info$address_table$address,"  "))

### 2015-02-13 ms/ms.Rmd
address in italics

@@
 writeLines("\n\n")
-writeLines(paste0("^", author_info$address_table$code, "^ ",
-                  author_info$address_table$address,"  "))
+writeLines(paste0("^", author_info$address_table$code, "^ *",
+                  author_info$address_table$address,"*  "))
 ```
